---
title: "1) Reading in All Data And Correcting IDs"
output: html_document
date: "2026-06-26"
---


```{r Universal code block setting}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)
knitr::opts_chunk$set(warning = FALSE)

```


```{r loading in the packages, warning = FALSE, echo = TRUE}
library(readxl)
library(tidyverse)
library(kableExtra)
library(psych)
library(gt)
```

```{r specifying inputs and outputs}
# Specify the pathway to where the MegaGrant data is
raw_data_input <- '\\\\files.times.uh.edu/labs/MIR_Lab/MEGAGRANT/STUDY 1/FINAL_DS'

# Specify the pathway to where the EEG data is
eegPath <- "\\\\files.times.uh.edu/labs/MIR_Lab/MEGAGRANT/STUDY 1/FINAL_DS/EEG/rsEEG"
  
# Specify the output to where we want our data to be saved
clean_data_output <- "~/GitHub/MegaGrant/Analyses/Final Masters Analysis/data"
```


### Part 1: Loading in Behavioral Data and Doing Some Minor Data Cleaning

Here we are loading demographic, WHOQOL_BRIEF, medical, CFIT, and ARFA Spelling information 

- demographic information (age, sex) - GOLD STANDARD FOR IDS
- Medical information (to remove neurodevelopmental disorders or head injuries)
- the item scores for the CFIT (non-verbal fluid intelligence test)
- the items for the ARFA spelling task

The read in data is all coming from the server: Specifically in this pathway:

- '\\files.times.uh.edu\labs\MIR_Lab\MEGAGRANT\STUDY 1'

There are incorrect/typo IDs in all of these datasets except for:

- 'MegaGrant_TBL_Database_Newest_MCh with duration.xlsx'

Thus, the IDs within this excel file, which will be saved as the object demo, will be the ones **chosen to represent the true IDs of MegaGrant**. This means that IDs in other datasets that differ from demo will be **instantly dropped**.


```{r load in the predictors covariates and outcome data, warning= FALSE, results='hide'}
# Set the working directory
Mega <- raw_data_input
setwd(Mega)

# load in the data 
demo <- read_excel("Demo/MegaGrant_TBL_Database_Newest_MCh with duration.xlsx")
CFIT <- read.csv("CFIT/CFIT.scores.csv")
ARFA <- read.csv("ARFA/ARFA.Spelling.Raw.Scores.csv")
M1_org <- read_excel("Medical_History/TBL_WHOQOL_BRIEF_Medical_S1_S3_DM_06.xlsx",  sheet = "med s1")
M3_org <- read_excel("Medical_History/TBL_WHOQOL_BRIEF_Medical_S1_S3_DM_06.xlsx",  sheet = "med s3")
W1_org <- read_excel("Medical_History/TBL_WHOQOL_BRIEF_Medical_S1_S3_DM_06.xlsx",  sheet = "whoqol s1")
W3_org <- read_excel("Medical_History/TBL_WHOQOL_BRIEF_Medical_S1_S3_DM_06.xlsx",  sheet = "whoqol s3")

# Combine datasets from S1 and S3 together
M1 <- select(M1_org, ID, HeadTrauma, Health2epilepsy, Health2autism, ADD, Dislexia)
M3 <- select(M3_org, ID, HeadTrauma, Health2epilepsy, Health2autism, ADD, Dislexia)
W1 <- select(W1_org, ID, sleep = WHOQOL16, neg_feel = WHOQOL26)
W3 <- select(W3_org, ID, sleep = WHOQOL16, neg_feel = WHOQOL26)

M_comb <- unique(rbind(M1, M3))
W_comb <- unique(rbind(W1, W3))

# Clean up CFIT data
CFIT2 <- select(CFIT, ID, CFITSub1Item1:CFITSub4Item8, Raw.Scores)
CFIT3 <- data.frame(sapply(CFIT2, function(x) as.numeric(x)))
CFIT3[is.na(CFIT3)] <- 0
CFIT4 <- select(CFIT3, - Raw.Scores)

# Rename the ARFA variables
names(ARFA) <- c("ID", paste0("SP_ER",1:22), "Total_Error")

# Have both ID variables be numeric
demo$ID <- as.numeric(demo$ID)
M_comb$ID <- as.numeric(M_comb$ID)
W_comb$ID <- as.numeric(W_comb$ID)
CFIT4$ID <- as.numeric(CFIT4$ID)
ARFA$ID <- as.numeric(ARFA$ID)

# Change additional variables to numeric
W_comb$sleep <- as.numeric(W_comb$sleep)
W_comb$neg_feel <- as.numeric(W_comb$neg_feel)

# If there are NA's for IDs drop them
demo <- drop_na(demo, ID)
M_comb <- drop_na(M_comb, ID)
W_comb <- drop_na(W_comb, ID)
CFIT4 <- drop_na(CFIT4, ID)
ARFA <- drop_na(ARFA, ID)

# Select variables of interest from demographics
demo <- demo %>%
  select(S1 = `S1 reg-list`,
         S3 = `S3 Reg-list`, 
         ID, 
         Sex, 
         Age, 
         Group) %>%
  mutate(ID = as.numeric(ID),
         Age = as.numeric(Age)) 

```

### Demographics + True IDS

```{r Showing the Demographics data head}
# Print the Demo Dataset
head(demo) %>%
  kbl() %>%
  kable_paper(full_width = F)

```

### WHOQOL_BRIEF (Kept only 2 items)

```{r Showing the WHOQOL_BRIEF data head}
# Print the WHOQOL_BRIEF Dataset
head(W_comb) %>%
  kbl() %>%
  kable_paper(full_width = F)

```

### Medical Questionnaire (Kept 5 items)

```{r Showing the Medical data head}
# Print the Medical Dataset
head(M_comb) %>%
  kbl() %>%
  kable_paper(full_width = F)

```

### CFIT Fluid Intelligence (46 items)

```{r Showing the CFIT data head}
# Print the CFIT Dataset
head(CFIT4) %>%
  kbl() %>%
  kable_paper(full_width = F)

```


### ARFA Spelling Subtest (22 Items)

```{r Showing the ARFA data head}
# Print the ARFA Dataset
head(ARFA) %>%
  kbl() %>%
  kable_paper(full_width = F)

```

### Part 2: Loading in the EEG FFT Data

FFT denotes the data files that include **power spectra** information of our resting-state EEG eyes-closed outcome. All we are doing is reading in the data that was originally saved in individual .csv files and combining them into one giant dataset plus a little bit of data cleaning.

**Important:** There are 3 EEG files that essentially have a duplicate with a different preprocessed file name. Meaning that when you extract the IDs from all the processed file names and use the unique() function (as we have done in the code), the duplicate data goes away. This is done for both the FFT and the QC section.

```{r Reading in the fft EEG files}
# Get all file names and create an empty list
all_fft_files <- list.files(path = file.path(eegPath, "fft"), pattern = "fft")
all_fft_files_full <- file.path(eegPath, "fft", all_fft_files)
all_files_list <- list()

# Read in each file one by one and save within a list
for(ii in 1:length(all_fft_files)){
  # Read in the file
  read_file <- read.csv(all_fft_files_full[ii])
  # Saves the file name within the file
  read_file$filename <- all_fft_files[ii]
  # Save the recording into a list
  all_files_list[[ii]] <- read_file
  names(all_files_list)[ii] <- all_fft_files[ii]
}

# Combine the FFT files into one dataset
FFT_dat <- do.call(rbind, all_files_list)

# Create a function that accurately extracts the ID of the file
clean_eeg_ids <- function(x) {
  x_gsub <- gsub("_preproc_fft.csv|EyesClosed|_RAW|_ RAW|RAW_|_", "", x)
  return(x_gsub)
}

# Recover the ID for each recording + specify condition
FFT_dat <- FFT_dat %>%
  mutate(ID = clean_eeg_ids(filename),
         Condition = ifelse(grepl("Open", filename), "Open", "Closed"))

# Drop the file name information and use unique()
FFT_dat <- select(FFT_dat, - filename) %>%
  unique()

# Remove row names
row.names(FFT_dat) <- NULL

# Force all IDs to be numeric
FFT_dat <- mutate(FFT_dat, ID = as.numeric(ID))

```


```{r Print information on the fft dataset}
# Show the information of the FFT data
head(FFT_dat, 8) %>%
  kbl(caption = paste0("First few rows of the FFT Dataset")) %>%
  kable_paper(full_width = F)
```


### Loading EEG Preprocessing QC

Here we are loading the QC performance for each EEG recording and only keeping a few variables of interest.

```{r Reading in the EEG Preprocessing QC Results}
# load in the EEG QC
EEG_QC <- read.csv(file.path(eegPath, "EEG_Preprocessing_Summary_Statistics_EC.csv"))

# Convert data to long format to get number of channels removed (80% threshold)
chan_rmvthresh <- .90

# Obtain the number of removed channels
EEG_QC$chan_rmv <- rowSums(select(EEG_QC, int_AF3:int_TP8) > chan_rmvthresh)
EEG_QC$chan_rmv2 <- rowSums(select(EEG_QC, int_AF3:int_TP8) > .99)

# Select variables of interest
EEG_QC2 <- EEG_QC %>% 
  select(subject,
         start_trial = s5_rjctbadtrials_initialtrials,
         start_globvar = s2_eegvarsum_globalvar,
         end_trial = endTrialnum,
         end_globvar = s10_eegvarsum_globalvar,
         int_total,
         #elec_pop = s4_rmvbadchan_pop_chan,
         #elec_flat = s4_rmvbadchan_flat_chan,
         #elec_nois = s4_rmvbadchan_noise_chan,
         chan_rmv,
         chan_rmv2) %>%
  mutate(trials_rmv = start_trial - end_trial)

# Use the same cleaning as above on this data
EEG_QC2 <- EEG_QC2 %>%
  mutate(ID = clean_eeg_ids(subject),
         Condition = ifelse(grepl("Open", subject), "Open", "Closed"))

# Drop the subject name information and use unique() - at the same time change order of variables
EEG_QC2 <- select(EEG_QC2, ID, Condition, start_trial,end_trial, trials_rmv, start_globvar, everything(), - subject) %>%
  unique()

# Force all IDs to be numeric
EEG_QC2 <- mutate(EEG_QC2, ID = as.numeric(ID))
```

```{r Show the head for the QC data}
# Show the information of the QC data
head(EEG_QC2, 8) %>%
  kbl(caption = paste0("First few rows of the QC Dataset")) %>%
  kable_paper(full_width = F)
```

### Part 3: Fixing a Minor ID Problem in the EEG Files

There is an excel file on the server that shows which IDs are typos and need to be renamed. The table below shows us the typo id (ID) and the correct ID name (RAW). Here we will be renaming the typos to the correct ID so that we do not need to throw out unecessary data.

Interestingly- it seems that the data that corresponds to the typo ID has already been corrected for in the data since the proper ID information is already present. Thus, we don't really need to do anything in this section since we already had the data for the proper ID name to begin with. 

```{r Removing problematic IDs}
# Read in the EEG mismatch dataset
megaGrant_dataInventory <- read_excel(file.path(eegPath, "MegaGrant_TBL.xlsx"))

# Check something
eeg_id_mistmatch <- megaGrant_dataInventory[megaGrant_dataInventory$ID != megaGrant_dataInventory$RAW,]

# Select the variables of interest and create a table
eeg_id_mistmatch_tab <- eeg_id_mistmatch %>%
  drop_na(ID) %>%
  select(ID, RAW)

# Print the table
cat("Below are IDs that have been incorrectly labeled (ID) and need to be renamed to the RAW values\n")
eeg_id_mistmatch_tab %>%
  kbl() %>%
  kable_paper(full_width = F)

# Visual inspection of the data
EEG_QC2 %>%
  filter(ID %in% c("34857", "34867"))
EEG_QC2 %>%
  filter(ID %in% c("97911", "97991"))

```


### Part 4: Removing All IDs From All Datasets That Do Not Match Demo

```{r Removing all IDs that do not match demo, echo = TRUE}
# Create a list that contains all datasets 
all_dat <- list(
  demo = demo,
  M_comb = M_comb,
  W_comb = W_comb,
  CFIT4 = CFIT4,
  ARFA = ARFA,
  FFT_dat = FFT_dat,
  EEG_QC2 = EEG_QC2
)

# Keep only the observations that had IDs that match demo
all_dat2 <- lapply(all_dat, function(x){filter(x, ID %in% demo$ID)})

# Compare the ID differences before and after the removal
compare_df <- cbind(
  do.call(c, lapply(all_dat, function(x) length(unique(x$ID)))),
  do.call(c, lapply(all_dat2, function(x) length(unique(x$ID))))
) %>%
  data.frame() %>%
  mutate(id_lost = X1-X2)

# Print the table
names(compare_df) <- c("Original", "Corrected", "Difference")
compare_df %>%
  kbl() %>%
  kable_paper(full_width = F)


# Paste a note on the number of unique IDs
cat("Based on the table above, our intital number of unique real IDs is", length(unique(demo$ID)), "although, not everyone will probably have demographic data!")
```


### MegaGrant Sample Demographics

```{r Get the demographics of the sample}
# Create a summary table for Total Errors Score 
describe(all_dat2$demo$Age) %>%
  round(2) %>%
  gt() %>%
  tab_header(title = "Descriptive Statistics of Demo Age") %>%
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_body(columns = everything()))

# Get the sex of the sample
table(all_dat2$demo$Sex)
```

### Save the Corrected ID Datasets Independently

Here we will be saving 7 datasets

- demo
- Medical Questionnaire
- WHOOQL
- CFIT
- ARFA
- FFT data
- EEG QC

```{r Saving the ID corrected datasets, echo = TRUE, warning = FALSE}
# Change the working directory
setwd(clean_data_output)

# Save all of the datasets
saveRDS(all_dat2$demo, "1_demo.rds")
saveRDS(all_dat2$M_comb, "1_Medical.rds")
saveRDS(all_dat2$W_comb, "1_WHOQL.rds")
saveRDS(all_dat2$CFIT4, "1_CFIT.rds")
saveRDS(all_dat2$ARFA, "1_ARFA.rds")
saveRDS(all_dat2$FFT_dat, "1_FFT.rds")
saveRDS(all_dat2$EEG_QC2, "1_EEGqc.rds")
```