---
title: "3) Exploring CFIT Using Item Response Theory Models"
output: html_document
date: "2026-06-29"
---

```{r Universal code block setting}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)
knitr::opts_chunk$set(warning = FALSE)

```


```{r loading in the packages, warning = FALSE, echo = TRUE}
library(tidyverse)
library(psych)
library(gt)
library(mirt) # mirt()
library(kableExtra)
library(caret) #findCorrelation() 
library(ggmirt)  #testInfoPlot()
library(patchwork)
```


```{r specifying inputs and outputs}
# Specify the pathway to where the MegaGrant data is
raw_data_input <- '~/GitHub/MegaGrant/Analyses/Final Masters Analysis/data'
  
# Specify the output to where we want our data to be saved
clean_data_output <- "~/GitHub/MegaGrant/Analyses/Final Masters Analysis/data"

# Specify the output for saving tables
save_table_output <- "~/GitHub/MegaGrant/Analyses/Final Masters Analysis/tables"
```

# Reading in the CFIT data

The CFIT Scale 2 Form B is a **timed** non-verbal fluid intelligence test. It contains 46 items and 4-subscales:

- Series (12 items)
- Classifications (14 items)
- Matrices (12 items)
- Conditions (8 items)



```{r Reading in the data, warning= FALSE}
# Reading in the CFIT data
setwd(raw_data_input)
CFIT <- readRDS("1_CFIT.rds")

# Cleaning the data a bit
CFIT_names <- names(CFIT)
CFIT_names <- gsub("CFIT", "", CFIT_names)
CFIT_names <- gsub("Item", "_", CFIT_names)
names(CFIT) <- CFIT_names

# Generate a total score
CFIT$raw <- rowSums(select(CFIT, Sub1_1:Sub4_8))
```

### Missing Data

We are quickly going to group items by subtest to assess for missing data. We will not remove anyone for having missing data since the IRT models are expected to be able to handle this.

```{r Missing data check}
# Create a list where each element corresponds to a subtest
subtest_list <- list(
  series = select(CFIT, Sub1_1:Sub1_12),
  classification = select(CFIT, Sub2_1:Sub2_14),
  matrices = select(CFIT, Sub3_1:Sub3_12),
  conditions = select(CFIT, Sub4_1:Sub4_8)
)

# Check for missing data
lapply(subtest_list, function(x) sum(is.na(x)))
```

```{r Exploring the item descriptives}
# Aesthetics (Leo version)
theme_clean <- function() {
  theme_minimal() +
    theme(legend.position = "bottom",
          panel.grid.minor = element_blank(),
          plot.title = element_text(hjust = 0.5)) 
}

# Create a summary table for Total Errors Score 
describe(CFIT$raw) %>%
  round(2) %>%
  gt() %>%
  tab_header(title = "Descriptive Statistics of CFIT Performance Score") %>%
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_body(columns = everything()))

# Number of unique IDs plus min and max scores
unique_Ids <- length(unique(CFIT$ID))
mean_score <- round(mean(CFIT$raw),1)
sd_score <- round(sd(CFIT$raw),1)
skew_score <- round(skew(CFIT$raw),2)
min_score <- min(CFIT$raw)
max_score <- max(CFIT$raw)

# Create a histogram of raw error sum score
CFIT %>%
  ggplot(aes(x = raw)) +
  geom_histogram(color = "black", fill = "white", bins = 20) +
  labs(title = paste0("CFIT Performance Score Distribution\n(n= ", unique_Ids,
                      "; mean =", mean_score,"; sd = ",sd_score, "; skew = ",skew_score ,"; min = ",min_score,"; max = ",max_score,")")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(x = "CFIT Raw Score", y = NULL) +
  theme_clean()

# Create a data frame with frequency of error by item
item_score_freq <- CFIT %>%
  select(- raw) %>%
  pivot_longer(cols = -ID,
               names_to = "Items",
               values_to = "Score") %>%
  group_by(Items, Score) %>%
  summarise(count = length(ID)) %>% 
  mutate(Score = factor(Score)) %>%
  ungroup()

# Create a factor with the levels of interest for plotting purposes
item_score_freq <- item_score_freq %>%
  mutate(Score = factor(Score, levels = c("0", "1")))

# Generate a plot
item_score_freq %>%
  mutate(
    subtest = str_split_i(Items, "_", 1),
    Item_num = str_split_i(Items, "_", 2),
    Item_num = factor(Item_num, 
                      levels = 14:1)) %>%
  ggplot(aes(x = Item_num, y = count, fill = Score)) +
  geom_col() +
  facet_wrap(~subtest, scales = "free") +
  theme_clean() +
  scale_fill_brewer(palette = "Paired") +
  labs(title = "Frequency of Number of Errors by Item") +
  coord_flip()

# Select only the items from CFIT
CFIT_items <- select(CFIT, Sub1_1:Sub4_8)

# Get the proportion of scores for each item
result <- lapply(CFIT_items, function(x) {
  tab <- table(x)
  prop.tab <- round(prop.table(tab), 2)
  data.frame(prop.tab) %>% rename(score = x, prob = Freq)
})

# Bind the rows and include item name information
CFIT_response_prob <- bind_rows(result, .id = "Item")

# Convert information into wide formar
CFIT_response_prob_wide <- CFIT_response_prob %>%
  pivot_wider(names_from = score, values_from = prob)

```

We have the code to show the proportion of incorrect and correct items for the CFIT- but it is a long table and it does not provide us any useful information that we can't already infer from the previous plots.

```{r currently hidden table of proportions of 0s and 1s across items, echo = TRUE}
# Create a gt() table for our CFIT response probabilities
Table1A <- CFIT_response_prob_wide %>%
  gt() %>%
  tab_options(
    table.font.size = 11,           # make tighter if needed
    table.width = pct(45)           # helps side-by-side fit
  ) %>%
  tab_header(
    title = "A: CFIT Subtest Probability of Response Category Endorsement by Item"   # Required: your main title
  )
# Table1A

```


Technically a parallel analysis is better but I don't really care tbh. The main takeaway is most of the variance is explained by one factor, but there is evidence in the data present for a few very weak factors, which likely correspond to question similarity, since questions correspond to one of four subtests named: series, classification, matrices, conditions.

```{r Running EFA on the CFIT Data}
# Create a matrix of polychoric correlations
CFIT_cor <- polychoric(CFIT_items)

# Run an exploratory factor analysis with five factors
efa2 <- fa(CFIT_cor$rho, nfactors = 5, fm = "minres") # Now, running EFA.
efa2$Vaccounted

```


```{r Scree plot of the EFA}
# Create a Scree Plot
data.frame(Eigenvalue = efa2$values,
           Factor = 1:length(efa2$values)) %>%
  ggplot(aes(x = Factor, y = Eigenvalue)) +
  geom_point(size = 3) +
  geom_line() +
  labs(title = "Scree Plot Using Eigenvalues from Exploratory Factor Analysis (CFIT)") +
  geom_hline(yintercept= 1, linetype="dashed", color = "red") +
  theme_clean()
```


### Running a 2PL, 4-factor 2PL, and Bi-Factor 2PL Model on the CFIT Data

We chose not to fit a hierarhical model because a) it failed to converge (producing an error) and b) theoretically it is not that good of a representation of IQ.

From the model comparison below, the Bi-Factor model produced the best fit (lowest AIC & BIC)

```{r running MIRT on the CFIT data First Iteration, echo=FALSE, results='hide'}
# Run a 2PL Model
CFIT_2pl <- mirt(CFIT_items, 1, itemtype = "2PL")

# Fitting a 4 factor correlated model
CFIT_4factor <- mirt(
  CFIT_items,
  model = "
  F1 = 1-12
  F2 = 13-26
  F3 = 27-38
  F4 = 39-46
  COV = F1*F2, F1*F3, F1*F4,
        F2*F3, F2*F4,
        F3*F4
  ",
  itemtype = "2PL"
)

# Create a vector indicating which specific factor each item belongs to
specific <- c(rep(1, 12),   # items 1-12   → F1
              rep(2, 14),   # items 13-26  → F2
              rep(3, 12),   # items 27-38  → F3
              rep(4, 8))    # items 39-46  → F4

# Fit a bi-factor model to the data
CFIT_bifactor <- bfactor(data = CFIT_items, 
                         model = specific,     # this defines the specific factors
                        itemtype = "2PL")


```

We don't care for the likelihood ratio tests below- what we are interested in is the AIC and BIC values, in which we see that the bi-factor model is the best fitting model for the data.

```{r First round of model comparison for the IRT models}
# Compare models to see if bi-factor is better than the 2PL
anova(CFIT_2pl, CFIT_4factor, CFIT_bifactor) %>%
  data.frame() %>%
  mutate_if(is.numeric, ~round(., 3)) %>%
  kbl(caption = "Model Fit Indices of the three CFIT IRT Models (First Round)") %>%
  kable_paper(full_width = F)
```


### Printing Bi-Factor Model Output Summary (Problematic)

We used the coef() function to produce discrimination (a) slopes for each item that correspond to the general factor and the specific-factor (question type). We do see some problems, in that some items have very very low slopes! (below 1). Thus, we should go ahead and remove these items. 

```{r Print bifactor model summary}
# Print the coefficients
#coef(CFIT_bifactor, IRTpars = TRUE, simplify = TRUE)

# Save the item slopes
item_slopes <- coef(CFIT_bifactor, IRTpars = TRUE, simplify = TRUE)$items %>%
  data.frame() %>%
  select(a1:a5) %>%
  round(2)

# Identify good and bad items
good_items <- row.names(item_slopes)[item_slopes$a1 >= 1]
bad_items <- row.names(item_slopes)[item_slopes$a1 < 1]


# Save a dataset that has the parameter information from the bi-factor model
item_pars <- coef(CFIT_bifactor, IRTpars = TRUE, simplify = TRUE)$items %>%
  data.frame() %>%
  select(a1:d)

# Introduce an item variable
item_pars$item <- row.names(item_pars)

# Extract rows that correspond to different subtests and save them as a list
subtest_list <- list(
  Sub1 = item_pars %>% filter(str_detect(item, "^Sub1")),
  Sub2 = item_pars %>% filter(str_detect(item, "^Sub2")),
  Sub3 = item_pars %>% filter(str_detect(item, "^Sub3")),
  Sub4 = item_pars %>% filter(str_detect(item, "^Sub4"))
)

# Change the item info to just numbers
subtest_list <- lapply(subtest_list, function(x) {
  x$item <- as.character(str_split_i(x$item, "_", 2))
  x
})

# making sure the variables are all in the same orientation for each list
subtest_tables <- imap(subtest_list, function(df, subtest_name) {
  df %>%
    mutate(item_num = row_number()) %>%
    select(item_num, item, a1, a2, a3, a4, a5, d) %>%
    rename_with(
      .fn = function(x) paste(subtest_name, x, sep = "_"),
      .cols = -item_num
    )
})

# Combine everything into one wide dataset
wide_table <- reduce(subtest_tables, full_join, by = "item_num") %>% data.frame()
wide_table_org <- wide_table %>% mutate_if(is.numeric, ~round(., 2)) 

# Some data cleaning
wide_table <- wide_table %>% mutate_if(is.numeric, ~round(., 2))
wide_table[is.na(wide_table)] <- ""
wide_table <- select(wide_table, - item_num)
names(wide_table) <- rep(c("Item\nNum.","a1", "a2", "a3", "a4", "a5", "b"), 4)

# Print the table with the item parameters
wide_table %>%
  kbl() %>%
  add_header_above(c("Series" = 7, "Classifications" = 7, "Matrices" = 7, "Conditions" = 7)) %>%
  kable_paper(full_width = F)

# Save the table
table_name = "Table_APPEND_B_1.csv"
full_table_name = file.path(save_table_output, table_name)
write.csv(wide_table_org, full_table_name, row.names = FALSE, quote = FALSE)
```


### Removing Problematic Items (Low Slopes) and Re-running a Unidimensional and Bi-Factor Model

From the results above, the coefficients for each item, we decided to remove the following items:
- subtest1: 3, 9, 10, 11, 12
- subtest2: 4, 5, 7, 9, 10, 11, 12, 13, 14
- subtest3: 12
- subtest4: 6, 7, 8

Looks like the bi-factor model failed to converge so we will go with the undimensional 2PL version


```{r Removing problematic CFIT Items, echo = FALSE, results='hide'}
# Remove these items from the data
CFIT_rmvItms <- CFIT_items %>%
  select(- bad_items)

# Number of remaining items per factor
substr(names(CFIT_rmvItms), 1, 4) %>% table()

# Run the unidimensional model
CFIT_2pl_rmvItms <- mirt(CFIT_rmvItms, 1, itemtype = "2PL")

# Fitting a 4 factor correlated model
CFIT_4factor_rmvItems <- mirt(
  CFIT_rmvItms,
  model = "
  F1 = 1-7
  F2 = 8-12
  F3 = 13-23
  F4 = 24-28
  COV = F1*F2, F1*F3, F1*F4,
        F2*F3, F2*F4,
        F3*F4
  ",
  itemtype = "2PL"
)

# Create a vector indicating which specific factor each item belongs to
specific <- c(rep(1, 7), 
              rep(2, 5),  
              rep(3, 11),   
              rep(4, 5))    

# Fit a bi-factor model to the data
CFIT_bifactor_rmvItems <- bfactor(data = CFIT_rmvItms, 
                                  model = specific,     # this defines the specific factors
                                  itemtype = "2PL")

```


While the bi-factor model **DOES PRODUCE BETTER FIT INDICIES**- THE MODEL **FAILED TO CONVERGE** UP to 2,500 ITERATIONS. It seems that the Max-Change value is just stuck in some local maxima.

```{r Show the model fit indices for the second round of IRT fitting}
# Model Comparison
anova(CFIT_2pl_rmvItms, CFIT_4factor_rmvItems, CFIT_bifactor_rmvItems)  %>%
  data.frame() %>%
  mutate_if(is.numeric, ~round(., 3)) %>%
  kbl(caption = "Model Fit Indices of the three CFIT IRT Models (Second Round)") %>%
  kable_paper(full_width = F)
```


### Local Dependence Check on the Unidimensional CFIT Model

There is local dependence for the undimensional model (not good)

```{r Local dependecy check on unidimensional CFIT model, echo = FALSE, results='hide'}
# Extract the residuals from the models
q3 <- residuals(CFIT_2pl_rmvItms, type = "Q3")

# Find the items that have higher residual correlations with other items than the threshold
LD_items <- findCorrelation(q3, cutoff = 0.20, verbose = F)
```

Below is the list of items that have local dependence with another item on the list

```{r Show the local dependence correlations}
# Print out the items
LD_items
```

### Removing Items with Local Dependence

To address the local dependence problem, we will be removing items based on the results above. Right now the problematic item pairs are: 

Problematic row pairs (correlations)

- row 25 & 26
- row 26 & 24
- row 28 & 6
- row 3 & 4
- row 4 & 8
- row 6 & 7
- row 8 & 10
- row 8 & 9
- row 14 and 13

We will remove the following rows/items:

- Remove: row 26: Sub4_3
- Remove: row 6: Sub1_7
- Remove: row 4: Sub1_5
- Remove: row 8: Sub2_1
- Remove: row 13: Sub3_1


Second round!
There was still local depedence after the initial removal of items above, so we also removed these items as well:

- Remove: Sub4_1
- Remove: Sub4_2
- Remove: Sub1_4


```{r dropping local dependent items in CFIT, results = 'hide'}
# Create a data frame to quickly index what item number correspond to what items
item_num <- data.frame(num = 1:length(CFIT_rmvItms), cbind(names(CFIT_rmvItms)))

# Drop redundant items
CFIT_rmvItmsv2 <- CFIT_rmvItms %>%
  select(-Sub4_3,- Sub1_7, -Sub1_5, -Sub2_1, - Sub3_1)

# Drop some more items (these items also showed local dependence after removal of above items)
CFIT_rmvItmsv3 <- CFIT_rmvItmsv2 %>%
  select(-Sub1_4 ,- Sub4_1, -Sub4_2)

# Fit the 2PL model on new dataset
CFIT_2pl_rmvMore <- mirt(CFIT_rmvItmsv3, 1, itemtype = "2PL")

```

```{r Check for local dependence in this 2PL model, results='hide'}
# Identify if residual correlations exceed threshold, suggesting local dependence (Bad)
q3 <- residuals(CFIT_2pl_rmvMore, type = "Q3")
LD_items2 <- findCorrelation(q3, cutoff = 0.2, verbose = T)
```

There are no LD item pairs in the 2PL model after having several items removed

```{r Print out the LD items in the 2PL Model}
LD_items2
```

#### Final 2PL CFIT Model Parameters

We lost quite a bit of items! But that should not be true problematic assuming that we still have good coverage for the information/conditional reliability curves.

```{r Print the 2PL Model Parameters of Final Unidimensional Model}
# Print the coefficients (slopes and thresholds) for each item
item_pars2 <- coef(CFIT_2pl_rmvMore, IRTpars = TRUE, simplify = TRUE)$items %>%
  data.frame() %>%
  mutate_if(is.numeric, ~round(., 2)) 

# Introduce an item variable
item_pars2$item <- row.names(item_pars2)

# Extract rows that correspond to different subtests and save them as a list
subtest_list2 <- list(
  Sub1 = item_pars2 %>% filter(str_detect(item, "^Sub1")),
  Sub2 = item_pars2 %>% filter(str_detect(item, "^Sub2")),
  Sub3 = item_pars2 %>% filter(str_detect(item, "^Sub3")),
  Sub4 = item_pars2 %>% filter(str_detect(item, "^Sub4"))
)

# Change the item info to just numbers
subtest_list2 <- lapply(subtest_list2, function(x) {
  x$item <- as.character(str_split_i(x$item, "_", 2))
  x
})

# making sure the variables are all in the same orientation for each list
subtest_tables2 <- imap(subtest_list2, function(df, subtest_name) {
  df %>%
    mutate(item_num = row_number()) %>%
    select(item_num, item, a, b) %>%
    rename_with(
      .fn = function(x) paste(subtest_name, x, sep = "_"),
      .cols = -item_num
    )
})

# Combine everything into one wide dataset
wide_table2 <- reduce(subtest_tables2, full_join, by = "item_num") %>% data.frame()

# Some data cleaning
wide_table2 <- wide_table2 %>% mutate_if(is.numeric, ~round(., 2))
wide_table2[is.na(wide_table2)] <- ""
wide_table2 <- select(wide_table2, - item_num)
names(wide_table2) <- rep(c("Item\nNum.","a", "b"), 4)

# Print the table with the item parameters
wide_table2 %>%
  kbl(caption = "Final Items of the CFIT 2PL Model") %>%
  add_header_above(c("Series" = 3, "Classifications" = 3, "Matrices" = 3, "Conditions" = 3)) %>%
  kable_paper(full_width = F)


```


```{r M2 Statistic of the Final 2PL Model }
# Print an omnibus test of the model
M2(CFIT_2pl_rmvMore, type = "C2")  %>%
  data.frame() %>%
  mutate_if(is.numeric, ~round(., 4)) %>%
  kbl(caption = "M2 Model Fit Plus Other Fit Indices for Final 2PL") %>%
  kable_paper(full_width = F)

```

Item level diagnostics all look good! No item has a p-value less than 0.001

```{r Showing the Item Level Diagnostics for the Final 2PL Model}
# Print out the item fit statistics
itemfit(CFIT_2pl_rmvMore) %>%
  data.frame() %>%
  mutate_if(is.numeric, ~round(., 4)) %>%
  kbl(caption = "Item Level Diagnostics for Final 2PL CFIT Model") %>%
  kable_paper(full_width = F)

```

### Information and Conditional Reliability Curve Shows Good Coverage for Lower Fluid Intelligence Ability

```{r Showing the Information and Conditional Reliability Curve of the Final 2PL Model}
# Use the correct plot functions to plot Item Information Function and Reliability
#testInfoPlot(CFIT_2pl_rmvMore, title = "B: Test Information Function")
#conRelPlot(CFIT_2pl_rmvMore, title = "Reliability of the CFIT Data")

# Creating the information curve by hand
info_reliability_curve <- data.frame(
  G = seq(-4, 4, by =.10)) %>%
  mutate(info = testinfo(CFIT_2pl_rmvMore, G),
         se = 1/sqrt(info),
         relia = info/(info + 1))

# Convert dataset to long- remove se associated with reliability
info_reliability_curve_long <- info_reliability_curve %>%
  pivot_longer(cols = c(info,relia))
info_reliability_curve_long$se[info_reliability_curve_long$name == "relia"] <-NA

# Make it prettier
info_reliability_curve_long <- info_reliability_curve_long %>% 
  mutate(name = ifelse(name == "info", "Information", "Reliability"))

# Plot the information curve
p2 <- info_reliability_curve_long %>%
  filter(name == "Information") %>%
  ggplot(aes(x= G, y = value)) +
  geom_line(aes(y = se), linetype = "dashed") +
  geom_line() +
  labs(y = NULL, x = "Fluid Intelligence", title = "A: Final CFIT 2PL\nInformation Curve") +
  theme_clean()

# Plot the conditional relaibility curve
p3 <- info_reliability_curve_long %>%
  filter(name == "Reliability") %>%
  ggplot(aes(x= G, y = value)) +
  geom_line(aes(y = se), linetype = "dashed") +
  geom_line() +
  geom_hline(yintercept=0.7, linetype="dashed", color = "red") + 
  labs(y = NULL, x = "Fluid Intelligence", title = "B: Final CFIT 2PL\nConditional Reliability Curve") +
  theme_clean()

# Show the information and reliability plots side by side
p2 + p3


# Paste a message about the theta levels that correspond between .7 and .8 reliability
reliability_num <- filter(info_reliability_curve_long, name == "Reliability")
reliability_70 <- reliability_num$G[reliability_num$value > .7]
reliability_80 <- reliability_num$G[reliability_num$value > .8]

cat("Reliability of .7 is present between",min(reliability_70),"and",max(reliability_70),"theta values")
cat("Reliability of .8 is present between",min(reliability_80),"and",max(reliability_80),"theta values")
  
```


### CFIT EAP Scores vs Raw Sum Scores Do like Quite Different

```{r Showing the relationship between raw sum scores and EAP scores}
# Plot the raw scores against the EAP scores
plot(CFIT_2pl_rmvMore, type = "EAPsum")

# Extract EAP scores (or use MAP/WLE)
eap_scores <- fscores(CFIT_2pl_rmvMore, method = "EAP")[, 1]

# Create the plot
EAP_df <- data.frame(ID = CFIT$ID, 
                     raw_cfit = CFIT$raw, # Original raw scores before item removal
                     eap_cfit = eap_scores)

EAP_df %>%
  ggplot(aes(x = eap_cfit, y = raw_cfit)) +
  geom_point(alpha = 0.4, color = "steelblue") +
  geom_smooth(method = "loess", color = "red", se = TRUE) +
  labs(title = "Relationship between Raw Sum Scores and IRT EAP Theta",
       x = "IRT Ability (EAP Theta)",
       y = "Raw Summed Score") +
  theme_clean()

# Print the correlation between raw scores and EAP scores
cat("The correlation between raw scores and EAP scores is:", round(cor(EAP_df$raw_cfit, EAP_df$eap_cfit),2))

```

```{r Saving the EAP scores, echo = TRUE, warning = FALSE}
# Change the working directory
setwd(clean_data_output)

# Save the ARFA data with the IRT scores
saveRDS(EAP_df, "3_CFIT_eap.rds")

```

