---
title: "8) QTPME Follow Up and Interpretations"
output: html_document
date: "2026-07-01"
---

```{r Universal code block setting}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)
knitr::opts_chunk$set(warning = FALSE)

```

```{r loading in the packages, warning = FALSE, echo = TRUE}
library(tidyverse)
library(ggplot2)
library(patchwork)
library(nlme) #fixef()
library(emmeans)
library(kableExtra)
library(multcomp) # glht()
library(rafalib) # imagemat()
library(nlraa) #SSquadp3xs() 
```


### Problem with multcomp- it uses MASS, which messes with the dplyr select() function

```{r Addressing Problem with multcomp, echo = TRUE}
library(conflicted)
conflict_prefer("select", "dplyr")
conflict_prefer("fixef", "nlme")
```

```{r specifying inputs and outputs}
# Specify the pathway to where the MegaGrant data is
raw_data_input <- "~/GitHub/MegaGrant/Analyses/Final Masters Analysis/mods"
  
# Specify the output to where we want our data to be saved
#clean_model_output <- "~/GitHub/MegaGrant/Analyses/Final Masters Analysis/mods"

# Specify the output for saving tables
save_table_output <- "~/GitHub/MegaGrant/Analyses/Final Masters Analysis/tables"
```


```{r reading in the QTPME models , warning = FALSE}
# Reading in the combined data
setwd(raw_data_input)
delta_mods <- readRDS("delta_mod.rds")
theta_mods  <- readRDS("theta_mod.rds")
alpha1_mods <- readRDS("alpha1_mod.rds")
```

```{r Extracting the best fitting models from each frequency band dataset}
# Extract the best fitting models
delta_mod <- delta_mods$mod2
theta_mod <- theta_mods$mod2
alpha1_mod <- alpha1_mods$mod2
```

## Omnibus Tests From Each of the Best Fitting Models

Fortunately, since each of the best fitting models is the three-way interaction one, we can display all of this information in one large table!

This is the output of using the anova() function with the type argument = 'marginal', which basically shows us the omnibust test using type III sums of squares. Basically, it is testing, does this term explain additional variation in the outcome after controlling for the effects of the other terms in the model. It is nice because this approach is not influenced by term order in the model.

Since the best fitting models did not include any interaction with spell group, everything we learned about generating a K matrix will not be helpful here lol. 

```{r Print the Omnibus Test for the Delta Mod}
# Create a custom function to fix up omnibus test results
clean_omnibus_results <- function(anova_tab) {
  anova_tab_df <- data.frame(anova_tab)
  anova_tab_df <- anova_tab_df %>%
    mutate(
      p.value = case_when(
        p.value < .001 ~ "<0.001***",
        p.value < .01 ~ paste0(round(p.value,3),"**"),
        p.value < .05 ~ paste0(round(p.value,3),"*"),
        TRUE ~ paste0(round(p.value,3))),
      F.value = round(F.value,3))
  anova_tab_df
}

# Create a nice table for each ANOVA test
delta_aov <- clean_omnibus_results(anova(delta_mod, type = "marginal"))
theta_aov <- clean_omnibus_results(anova(theta_mod, type = "marginal"))
alpha1_aov <- clean_omnibus_results(anova(alpha1_mod, type = "marginal"))

# Join the ANOVA tables together
joined_aov <- cbind(delta_aov, theta_aov, alpha1_aov)
rbind_aov <- rbind(delta_aov, theta_aov, alpha1_aov)

# Clean up the rbind dataset a bit
rbind_aov$Term <- row.names(rbind_aov); row.names(rbind_aov) <- NULL
rbind_aov <- select(rbind_aov, Term, everything())

# Modify the row names
row_names <- row.names(joined_aov)
row_names <- gsub("xs.", "", gsub("b.", "", gsub("a.", "", row_names)))
row_names <- gsub("\\(Intercept)", "Intercept", row_names)
row_names <- gsub("region", "Region", row_names)
row_names <- gsub("hemisphere", "Hemisphere", row_names)
row_names <- gsub("Spell_group", "Spell Group", row_names)
row_names <- gsub(":", " x ", row_names)

# Introduce the cleaned row names
row.names(joined_aov) <- NULL
joined_aov <- cbind(`.` = row_names, joined_aov)

# Print the joined table
joined_aov %>%
  kbl(
    caption = "Table 1: Omnibus Test of Fixed Effects for the Final QTPME Models",
    col.names = c("Term", 
                  "df num", "df den","F", "p",
                  "df num", "df den","F", "p",
                  "df num", "df den","F", "p"),
    align = c("l", 
              "c", "c", "c", "c", 
              "c", "c", "c", "c", 
              "c", "c", "c", "c"),
    format = "html" ) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman") %>%
  add_header_above(c(" " = 1 ,"Delta" = 4, "Theta" = 4, "Alpha1" = 4)) %>%
  pack_rows("Intercept Parameter (a)", 1, 6) %>%
  pack_rows("Instantaneous Slope Parameter (b)", 7, 8) %>%
  pack_rows("Join Point Parameter Parameter (xs)", 9, 10)


# Save the table
table_name1 = "Table3_Omnibus_Results.csv"
full_table_name1 = file.path(save_table_output, table_name1)
write.csv(rbind_aov, full_table_name1, row.names = FALSE, quote = FALSE)

```

## Spelling Beta Estimates Family With P-value Adjustments

```{r Create a custom function for a p value rounding function}
# Create a custom function to clean up p-values
clean_p <- function(p.value) {
  ifelse(p.value < .001, "<.001***",
         ifelse(p.value < .01, paste0(round(p.value,3), "**"),
                ifelse(p.value < .05, paste0(round(p.value,3), "*"),
                       paste0(round(p.value,3)))))
  
}
```

```{r Extract the spelling beta estimatess families and adjust p-values, echo = FALSE}
# Create a custom function to return all the spelling-related tests
extract_spellTests <- function(mod) {
  sum_tab <- data.frame(summary(mod)$tTable)
  sum_tab <- cbind(Term = row.names(sum_tab), sum_tab)
  row.names(sum_tab) <- NULL
  sum_tab[grepl("Spell|Inter", sum_tab$Term),]
}

# Extract the spelling tests for each of the frequency band datasets
delta_spllTest <- extract_spellTests(delta_mod)
theta_spllTest <- extract_spellTests(theta_mod)
alpha1_spllTest <- extract_spellTests(alpha1_mod)

# Add an adjusted p-value column
delta_spllTest$adj.p <- p.adjust(delta_spllTest$p.value, method = "BH")
theta_spllTest$adj.p <- p.adjust(theta_spllTest$p.value, method = "BH")
alpha1_spllTest$adj.p <- p.adjust(alpha1_spllTest$p.value, method = "BH")

# Clean the pvalues and round the other values
delta_spllTest <- mutate(delta_spllTest, p.value = clean_p(p.value), adj.p = clean_p(adj.p)) %>% mutate_if(is.numeric, ~round(., 2))
theta_spllTest <- mutate(theta_spllTest, p.value = clean_p(p.value), adj.p = clean_p(adj.p)) %>% mutate_if(is.numeric, ~round(., 2))
alpha1_spllTest <- mutate(alpha1_spllTest, p.value = clean_p(p.value), adj.p = clean_p(adj.p)) %>% mutate_if(is.numeric, ~round(., 2))

# Show the beta estimates with the adjusted p-values
comb_spllTest <- delta_spllTest %>%
  bind_rows(theta_spllTest) %>%
  bind_rows(alpha1_spllTest)

# Identify the spell rows
spell_rows <- which(grepl("Spell_group", comb_spllTest$Term))

# Print out the spelling tests
comb_spllTest %>%
  kbl(
    caption = "Table 2: Parameter Estimats for Spelling Group Effects",
    col.names = c("Term", 
                  "B", "SE", "DF","t", "p", "p (adjusted)"),
    align = c("l", 
              "c", "c", "c", "c", "c", "c"),
    format = "html" ) %>%
  kable_classic(full_width = FALSE, html_font = "Times New Roman") %>%
  #row_spec(spell_rows, bold = TRUE) %>%
  pack_rows("Delta Band", 1, 6) %>%
  pack_rows("Theta Band", 7, 12) %>%
  pack_rows("Alpha Band", 13, 18)

# Save the table
table_name2 = "Table4_Spelling_Beta_Adjusted.csv"
full_table_name2 = file.path(save_table_output, table_name2)
write.csv(comb_spllTest, full_table_name2, row.names = FALSE, quote = FALSE)

```

## Creating a Figure of the Predicted Values for each of the Spelling Groups

```{r Visualzing the Predicted Values for Each of the Spelling Groups, fig.height= 5, fig.width= 8}
# Aesthetics (Leo version)
theme_clean <- function() {
  theme_minimal() +
    theme(legend.position = "bottom",
          panel.grid.minor = element_blank(),
          plot.title = element_text(hjust = 0.5,
                                    size = 10),
          plot.caption = element_text(hjust = 0.5),
          strip.text = element_text(size = 12)
          ) 
}

# Create a custom function that returns an expanded grid of model factors
# We only care for spelling group- and since there were no significant spellin group interacton, we can assume its relation with power will be the same across all levels of topography and sex
expand_grid_fun <- function(dat) {
    expand.grid(
    Age_min_c = seq(min(dat$Age_min_c), max(dat$Age_min_c), length.out = 100),
    Sex = "F",
    hemisphere = "left",
    region = "central",
    Spell_group = levels(dat$Spell_group)
  )
}

# Extract the original data that the models were fitted to
delta_dat <- delta_mod$data
theta_dat <- theta_mod$data
alpha1_dat <- alpha1_mod$data

# Generate expanded grids for each of the three frequency bands
delta_expand <- expand_grid_fun(delta_dat)
theta_expand <- expand_grid_fun(theta_dat)
alpha1_expand <- expand_grid_fun(alpha1_dat)

# Add the predicted values back into the expanded grid
delta_expand$pred <- predict(delta_mod, newdata = delta_expand, level = 0)
theta_expand$pred <- predict(theta_mod, newdata = theta_expand, level = 0)
alpha1_expand$pred <- predict(alpha1_mod, newdata = alpha1_expand, level = 0)

# Add some important information to the expanded datasets
delta_expand <- mutate(delta_expand, Age = Age_min_c + 15, FB = "delta")
theta_expand <- mutate(theta_expand, Age = Age_min_c + 15, FB = "theta")
alpha1_expand <- mutate(alpha1_expand, Age = Age_min_c + 15, FB = "alpha1")

# Combine them together and same with the datasets
comb_expand <- rbind(delta_expand, theta_expand, alpha1_expand) %>%
  mutate(Spell_group = ifelse(Spell_group == "low", "Poor", "Good")) %>%
  mutate(FB = factor(FB, levels = c("delta", "theta", "alpha1"))) %>%
  mutate(Spell_group = factor(Spell_group, levels = c("Poor", "Good")))
comb_dat <- rbind(delta_dat, theta_dat, alpha1_dat) %>%
  mutate(FB = factor(FB, levels = c("delta", "theta", "alpha1")))  %>%
  mutate(Spell_group = ifelse(Spell_group == "low", "Poor", "Good"))
  
# Create the prediction plots
comb_dat %>%
  ggplot(aes(x = Age, y = log_absolute_power)) +
  geom_point(size = .2, alpha = .6) +
  facet_wrap(~FB, nrow = 1, scales = "free") +
  geom_line(data = comb_expand, aes(x = Age, y = pred, color = Spell_group), size = 1) +
  labs(y = "Log Absolute Power", x = "Age", color = "Spelling Group") +
  theme_clean()


```


## Full Fixed and Random Effects for All Models

```{r Creating several custom functions to extract random and fixed effects from the nlme models}
# Create a custom function to extract the fixed effects of the model in a tidy way
extract_fixed <- function(nlme){
  # Extract the model summary table
  coef_tab <- summary(nlme)$tTable
  
  # Save the information as a nice table frame
  coef_df <- data.frame(
    Term = rownames(coef_tab),
    coef_tab,
    row.names = NULL
  ) %>% tibble()
  
  # Clean up the p values
  coef_df <- mutate(coef_df, p.value = clean_p(p.value))
  
  # Extract the 95% CI
  CI <- nlme::intervals(nlme)$fixed %>% data.frame()
  
  # Combined the upper and lower into one var
  CI <- CI %>% transmute(est.,
                  CI = paste0("[",round(lower,2),",", round(upper,2),"]"))
  
  # Introduce the CI into the dataframe
  joined_df <- left_join(coef_df, rename(CI, Value = est.), by = "Value")
  
  # Change values to three decimal places
  joined_df <- joined_df %>% mutate_if(is.numeric, ~round(., 2))
  
  # Change the variable order
  joined_df %>% select(Term, Value ,Std.Error, CI , DF, t.value, p.value)
}

# Create a custom function to extract the random effects of the model in a tidy way
extract_random <- function(nlme){
  # Get the variance/covariance matrix
  vc <- nlme::VarCorr(nlme)
  
  # Extract the intercept and residual standard deviation
  var_intercept <- as.numeric(vc["a.(Intercept)", "Variance"])
  var_residual  <- as.numeric(vc["Residual", "Variance"])
  
  # Calculate the conditional ICC
  cond_ICC <- var_intercept / (var_residual + var_intercept)
  
  
  # Save the information as a data frame
  rand_df <- data.frame(
  Term = c("τ₀₀ (ID Intercept for 'a')", "σ² (Residual Variance)", "Conditional ICC"),
  Value = c(var_intercept, var_residual, cond_ICC) # Variances
  ) %>% tibble()
  
  # Change values to three decimal places
  rand_df %>% mutate_if(is.numeric, ~round(., 2))
}

# Create a custom function that returns a tidy table of fixed and random effects
tidy_fixed_random <- function(nlme) {
  bind_rows(extract_fixed(nlme), 
            extract_random(nlme))
}

```


```{r Extracting model fixed and random effects and generating a table}
# Get the fixed and random effects for each model
tidy_delta_mod <- tidy_fixed_random(delta_mod)
tidy_theta_mod <- tidy_fixed_random(theta_mod)
tidy_alpha1_mod <- tidy_fixed_random(alpha1_mod)

# Combine the datasets into one and round values to three decimal points
tidy_combined <- cbind(
  tidy_delta_mod,
  select(tidy_theta_mod, - Term),
  select(tidy_alpha1_mod, - Term)
)

# Replace NA's with just an empty character
tidy_combined[is.na(tidy_combined)] <- ""

# Create the table
full_param_table <- tidy_combined %>%
  kbl(
    caption = "Table 3: Fixed and Random Effect Parameter Estimates for each Frequency Band QTPME Model",
    col.names = c("Term", 
                  "B", "SE","95% CI", "DF","t", "p",
                  "B", "SE","95% CI", "DF","t", "p",
                  "B", "SE","95% CI", "DF","t", "p"),
    align = c("l", 
              "c", "c", "c", "c", "c", "c",
              "c", "c", "c", "c", "c", "c",
              "c", "c", "c", "c", "c", "c"),
    format = "html" #
  )  %>%
  # Apply strict APA styling
  kable_classic(full_width = FALSE, html_font = "Times New Roman") %>%
  add_header_above(c(" " = 1 ,"Delta" = 6, "Theta" = 6, "Alpha1" = 6)) %>%
  pack_rows("Intercept Parameter (a)", 1, 12, label_row_css = "font-weight: bold; border-bottom: 1px solid #eee;") %>%
  pack_rows("Instantaneous Slope Parameter (b)", 13, 14, label_row_css = "font-weight: bold; border-bottom: 1px solid #eee;") %>%
  pack_rows("Join Point Parameter (xs)", 15, 16, label_row_css = "font-weight: bold; border-bottom: 1px solid #eee;") %>%
  pack_rows("Random Effects", 17, 19, label_row_css = "font-weight: bold; border-bottom: 1px solid #eee;") %>%
  footnote(
    general = "95% CI represent 95% Wald confidence intervals. Parameter-level p-values are unadjusted and a direct result from the model output. τ₀₀ denotes the random intercept variance, σ² denotes the residual variance, and the conditional ICC represents the proportion of residual variance attributable to between-participant differences after accounting for the fixed effects.",
    general_title = "Note.",
    footnote_as_chunk = TRUE
  )

# Print the table
full_param_table
```





