---
title: "Analyzing resting-state EEG and Spelling Performance in Young Adults (Quadratic Plateau) VERSION3"
author: "Leandro Ledesma"
date: "2026-04-01"
output: html_document
---

### Universal block code settings

```{r setup}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)
knitr::opts_chunk$set(warning = FALSE)

```

### Loading in packages

```{r loading in the packages, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(readxl)
library(kableExtra)
library(broom) # Converts regression outputs into dataframes using the tidy() function
library(psych)
library(MASS, exclude = "select") # This package is loaded with QuantPsyc, must exclude "select" or you wont be able to use it. 
library(QuantPsyc) # Can use the lm.beta function to calculate the standardized betas
library(car) # To calculate VIF
library(performance) # ICC 
library(lme4) #glmer
library(interactions) # interact_plot
library(emmeans)
library(effects) #plot(allEffects(model))
library(sjPlot) # plot_model
library(MuMIn)
library(lmtest)
library(patchwork)
library(nlme) #gnls()
library(soiltestcorr) # quadratic_plateau()
library(mirt)
library(caret)
library(ggmirt)
library(skimr)
library(marginaleffects)
```


### Loading in our data and doing some data cleaning

Here we are loading demographic and medical information 

- demographic information (age, sex)
- Medical information (to remove DD or head injuries)

After some minor data cleaning we combine the information into one dataset. 

If there are NAs in the ID variable then those observations (rows) get dropped and any IDs that do not correspond to the first study (there are three of them) will also be removed.


```{r load in the predictors covariates and outcome data, warning= FALSE}
# Set the working directory
Mega <- '\\\\files.times.uh.edu/labs/MIR_Lab/MEGAGRANT/STUDY 1/FINAL_DS'
setwd(Mega)

# load in the data 
demo <- read_excel("Demo/MegaGrant_TBL_Database_Newest_MCh with duration.xlsx")
CFIT <- read.csv("CFIT/CFIT.scores.csv")
M1_org <- read_excel("Medical_History/TBL_WHOQOL_BRIEF_Medical_S1_S3_DM_06.xlsx",  sheet = "med s1")
M3_org <- read_excel("Medical_History/TBL_WHOQOL_BRIEF_Medical_S1_S3_DM_06.xlsx",  sheet = "med s3")
W1_org <- read_excel("Medical_History/TBL_WHOQOL_BRIEF_Medical_S1_S3_DM_06.xlsx",  sheet = "whoqol s1")
W3_org <- read_excel("Medical_History/TBL_WHOQOL_BRIEF_Medical_S1_S3_DM_06.xlsx",  sheet = "whoqol s3")

# Combine datasets from S1 and S3 together
M1 <- select(M1_org, ID, HeadTrauma, Health2epilepsy, Health2autism, ADD, Dislexia)
M3 <- select(M3_org, ID, HeadTrauma, Health2epilepsy, Health2autism, ADD, Dislexia)
W1 <- select(W1_org, ID, sleep = WHOQOL16, neg_feel = WHOQOL26)
W3 <- select(W3_org, ID, sleep = WHOQOL16, neg_feel = WHOQOL26)

M_comb <- unique(rbind(M1, M3))
W_comb <- unique(rbind(W1, W3))

# Duplication check
duplicated(M_comb$ID) %>% sum()
duplicated(W_comb$ID) %>% sum()

# Clean up CFIT data
CFIT2 <- select(CFIT, ID, CFITSub1Item1:CFITSub4Item8, Raw.Scores)
CFIT3 <- data.frame(sapply(CFIT2, function(x) as.numeric(x)))
CFIT3[is.na(CFIT3)] <- 0
CFIT4 <- select(CFIT3, - Raw.Scores)

# Duplication check
duplicated(CFIT4$ID) %>% sum()

# Have both ID variables be numeric
demo$ID <- as.numeric(demo$ID)
M_comb$ID <- as.numeric(M_comb$ID)
W_comb$ID <- as.numeric(W_comb$ID)
CFIT4$ID <- as.numeric(CFIT4$ID)

# Change additional variables to numeric
W_comb$sleep <- as.numeric(W_comb$sleep)
W_comb$neg_feel <- as.numeric(W_comb$neg_feel)

# If there are NA's for IDs drop them
demo <- drop_na(demo, ID)
M_comb <- drop_na(M_comb, ID)
W_comb <- drop_na(W_comb, ID)
CFIT4 <- drop_na(CFIT4, ID)

# Select variables of interest from demographics
demo <- demo %>%
  select(S1 = `S1 reg-list`,
         S3 = `S3 Reg-list`, 
         ID, 
         Sex, 
         Age, 
         Group) %>%
  mutate(ID = as.numeric(ID),
         Age = as.numeric(Age)) 

### Combine the dataset into one
data <- demo %>%
  full_join(CFIT4, by = "ID") %>%
  full_join(M_comb, by = "ID") %>%
  full_join(W_comb, by = "ID") 
  
# Drop any NA's from the following variables
data <- drop_na(data, ID)

# Check correlations
datt <- select(data, Age, sleep, neg_feel) %>%
  filter(complete.cases(.))

# Print out a correlation matrix
cor(datt) %>% round(2)

# Create a contingency table for extreme negative emotions and sleep problems
addmargins(xtabs(~sleep + neg_feel, datt))

# Counting missing data by row
data$rowMiss = rowSums(is.na(data))
```


### Loading in the ARFA Spelling data

- Here we will be loading ARFA spelling error data. This data has already been processed from an earlier script where each item had the potential to have several different types of errors. For example, each item had 7 variables where scores were given to indicate errors or even Russian words were used. No errors were rows where each of the 7 variables had all 0's, which was most of the dataset. What we are loading in below are **sum of errors for each item** capped at two (0, 1, 2). We capped it at two because that was recommended to us.

The task while in Russian, we can translate the 22 items down below:

1. Pardon
2. Compromise
3. Disciplined
4. Gouache
5. Fairly
6. Greenhouse
7. Supernatural
8. Uninspired
9. Away
10. Trajectory
11. Rely
12. Puddle
13. The flower couldn't grow long without being watered
14. At this bachelorette party, all the girls were inmarried
15. The birdcatcher waiter in the forest until midnight
16. Not all Zurich residents speak French
17. They hate sunbathing
18. She had nothing to thank him for
19. The train arrived at the northeast station
20. At ninety-five years old, her heart was still working well
21. What time did he sent out to meet the guests
22. The warrior is skilled with the sword


```{r loading and scoring ARFA spelling}
# Set the working directory
Mega <- '\\\\files.times.uh.edu/labs/MIR_Lab/MEGAGRANT/STUDY 1/FINAL_DS'
setwd(Mega)

# Load in errors for each item
ARFA <- read.csv("ARFA/ARFA.Spelling.Raw.Scores.csv")

# Rename the variables
names(ARFA) <- c("ID", paste0("SP_ER",1:22), "Total_Error")
```


### Number of Recruited Subjects

Our original excel file that contains either all or the great majority of IDs in MegaGrant. Basically MegaGrant is three studies saved in one directory, where participants between the studies for the most part are the same, however there are participants that are in one study and not another. This matters because I don't think the spelling data was collected just in STUDY1, but also collected in STUDY3 as well. 

We also need to identify and remove an IDs that have ARFA data but are not a part of the Final ID's. They either a) could have been forgotten to be added to the main demographic dataset or b) are a typo ID. To side with caution, we will consider them as typos and remove them. 

Actually, ignore below, it does not make to write an ID number for people who do not have data that we need for either the IRT scoring or the EEG as a function of age- so we will report an ID number that represents the total number of people with ARFA data.


```{r Number of recruited subjects}
# Print out the number of IDs
cat("There are",length(unique(data$ID)),"unique IDs in this dataset that are from S1 or S3\n\n")
cat("When it comes to ARFA data, we have", length(ARFA$ID),"unique IDS\n\n")

# Keep subjects only from study 1
data_S1 <- filter(data, S1 == "+")

# Print out IDs conditional on S1
cat("When only counting IDs from S1, we have", length(data_S1$ID), "unique IDs. This is concerning because this number is less than the ARFA dataset, which means that a few of the spelling scores are from IDs in S3. This confirmns that the ARFA spelling task is not unique to just S1 but also slightly to S3\n\n")

# Remove IDs that do not belong
cat("Below are the IDs in ARFA but not in the demo- they will be removed")
(rmdIDs <- setdiff(ARFA$ID, data$ID)) # IDs present in ARFA but no in demo excel
ARFA <- filter(ARFA, !ID %in% rmdIDs)

# Therefore the total number of participants we have is 
cat("\n\nThe number of IDs we have for both S1 and S3, which we are using for at least one analysis is", length(unique(data$ID))," but it does not matter we should use the ID number from available ARFA data which is", length(unique(ARFA$ID)))

# Quick demographic information
cat("\nMean age:"); mean(data$Age, na.rm = T)
cat("\nSD age:"); sd(data$Age, na.rm = T)
cat("\nProp F:"); prop.table(table(data$Sex))
```

### Visualizing the ARFA Spelling Error Data

Below we can visualize the distribution of total item error score (not the same as total number of incorrect items). Additionally, we can see the frequency of the number of errors by item.

```{r Visualize the ARFA spelling data}
# Aesthetics (Leo version)
theme_clean <- function() {
  theme_minimal() +
    theme(legend.position = "bottom",
          panel.grid.minor = element_blank(),
          plot.title = element_text(hjust = 0.5)) 
}

# Create a summary table for Total Errors Score 
library(gt)

describe(ARFA$Total_Error) %>%
  round(2) %>%
  gt() %>%
  tab_header(title = "Descriptive Statistics of Spelling Total Error Score") %>%
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_body(columns = everything()))

# Number of unique IDs plus min and max scores
unique_Ids <- length(unique(ARFA$ID))
mean_score <- round(mean(ARFA$Total_Error),1)
sd_score <- round(sd(ARFA$Total_Error),1)
skew_score <- round(skew(ARFA$Total_Error),2)
min_score <- min(ARFA$Total_Error)
max_score <- max(ARFA$Total_Error)

# Create a histogram of raw error sum score
ARFA %>%
  ggplot(aes(x = Total_Error)) +
  geom_histogram(color = "black", fill = "white", bins = 20) +
  labs(title = paste0("Total Spelling Item Error Score Distribution\n(n= ", unique_Ids,
                      "; mean =", mean_score,"; sd = ",sd_score, "; skew = ",skew_score ,"; min = ",min_score,"; max = ",max_score,")")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(x = "Total Spelling Error", y = NULL) +
  theme_clean()

# Create a data frame with frequency of error by item
item_err_freq <- ARFA %>%
  select(- Total_Error) %>%
  pivot_longer(cols = -ID,
               names_to = "Items",
               values_to = "Error") %>%
  group_by(Items, Error) %>%
  summarise(count = length(ID)) %>% 
  mutate(Error = factor(Error)) %>%
  ungroup()

# Create a factor with the levels of interest for plotting purposes
item_err_freq <- item_err_freq %>%
  mutate(Items = factor(Items, levels = paste0("SP_ER",22:1)),
         Error = factor(Error, levels = c("2", "1", "0")))

# Generate a plot
item_err_freq %>%
  ggplot(aes(x = Items, y = count, fill = Error)) +
  geom_col() +
  theme_clean() +
  scale_fill_brewer(palette = "Paired") +
  labs(title = "Frequency of Number of Errors by Item") +
  coord_flip()


```

### Switching Spelling Errors to Spelling Scores

In this section we have decided to take each item and subtract its error score from the value 2. Essentially this changes the direction of the data, such that if no error mistakes were made on an item, then they would get a score of 2. Additionally, if someone made 2 or more error mistakes for an item, then their spelling score will be a 0. 

Below we have a table essentially ordering items by their difficulty, with item means closer to 0 representing harder items while items closer to 2 being easier. Based on this we can see that items 1, 9 and 2 were considered predominalty easy while items 21 and 16 were harder. 

Anyway the following next sections are heavily inspired by: 

- "Getting started with the graded response model: An introduction and tutorial in R" by Rizqy Amelia Zein and Hanif Akhtar

```{r switching errors to scores}
# Extract just the ARFA items
ARFA_items <- select(ARFA, SP_ER1:SP_ER22) 

# Rename the items so that it looks cleaner in plots and tables
names(ARFA_items) <- paste0("Item ",1:22)

# Change the items to represent scores rather than errors
ARFA_items <- 2 - ARFA_items

# Use the skim function to look at descriptives of our data
ARFA_items_df <- skim(ARFA_items) %>% data.frame()

# Clean up the table
ARFA_items_df <- ARFA_items_df %>% mutate_if(is.numeric, ~round(., 3)) # Round numbers
ARFA_items_df2 <- select(ARFA_items_df, skim_variable:numeric.hist) # Keep variables of interest
ARFA_items_df3 <- rename(ARFA_items_df2, item = skim_variable,
                         p0 = numeric.p0, p25 = numeric.p25, p50 = numeric.p50, p75 = numeric.p75, p100 = numeric.p100,) # Rename variables

# View the item descriptives
ARFA_items_df3 %>%
  arrange(desc(numeric.mean)) %>%
  kbl() %>%
  kable_paper(full_width = F)
```


### Testing Data Unidimensionallity

Before we can specify an item response theory (IRT) model on our spelling data, we need to investigate that our data is **unidimensional**. This means that the items within our data have a similar general correlation with each other and that there are not subset of items that correlate highly with each other, which would indicate that they may be measuring something kinda different or very specific to spelling errors. We need unidimensionality because that is what our graded response model will assume. 

Anyway there are a couple of ways to assess this in our data, and we will be using both an **exploratory factor analysis** and a **parallel analysis approach**. They both are investigating unidimensionality but in slightly different ways. Regardless similar results between the two will give us certainty that our data are unidimensional. 

Below we are showing the output from the EFA, by output we mean the magnitude of the Eigenvalues produced. The parallel analysis literally says the same thing but takes a little longer to run, overall this shows support for a unidimensional dataset. 

In addition to the Scree Plot, in a 2 factor EFA, we see that the first factor explains 47% of the data while an additional factor increases this to 50%, showing further evidence that the data is well explained with just one factor.

```{r testing the data unidimensionality}
# Create a matrix of polychoric correlations
cor <- polychoric(ARFA_items)

# Run an exploratory factor analysis with two factors
efa2 <- fa(cor$rho, nfactors = 2, fm = "minres") # Now, running EFA.
print(efa2)

# Create a Scree Plot
data.frame(Eigenvalue = efa2$values,
           Factor = 1:length(efa2$values)) %>%
  ggplot(aes(x = Factor, y = Eigenvalue)) +
  geom_point(size = 3) +
  geom_line() +
  labs(title = "Scree Plot Using Eigenvalues from Exploratory Factor Analysis") +
  geom_hline(yintercept= 1, linetype="dashed", color = "red") +
  theme_clean()

# Parallel analysis
# fa.parallel(ARFA_items, fm="minres", fa="pc",  cor = "poly") # to remove or not remove?
# irt.fa(ARFA_items, nfactors = 1, fm = "minres")
```


### Running a Polytomous Graded Response Model (Item Reponse Theory Model)

We technically already did this in another R script. However, we will be redoing it again here to explain exactly how the process works better. Essentially, from the graphs above, we can see that each item contains 3 different types of categories (no mistakes = 0, one mistake = 1, at least two mistakes = 2) and that this information is in an ordinal scale (2 > 1 > 0). Thus, we can use a **polytomous graded response model** to basically estimate difficulty and discrimination parameters for each item. For the difficulty parameters, the model will be estimating two, which represent the level of ability where the item has a 50% change of having endorsement between category 0 and category 1 and between category 1 and category 2.


In the code below we essentially write mod = 'theta = 1-22', which basically means we are specifying a undimensional dataset with 22 items. 

```{r Running an IRT Model}
# Number of participants (rows)
nrow(ARFA_items)

# Specify the model
mod <- paste0("theta = 1-", length(ARFA_items))

# Run the mirt model and produce the theta scores
fit_mirt <- mirt(ARFA_items, 
                 1, 
                 model = mod, 
                 itemtype = "graded", 
                 SE = TRUE, 
                 technical = list(NCYCLES = 20000))
```


### Investigating Model Parameters

Below we are displaying a table that shows different parameters for each of the items. The first is the discrimination parameter (a), which basically represents how well the item is capturing the latent variable. In other words, how well the item can discriminate or distinguish between those that do and do not respond to an item at that level of ability. The next parameters are essentially **difficulty**, which represent where on the latent variable that item is being endorsed. However, we have two parameters of difficulty for each item, and that is because our original data had three possible responses for each item (0, 1, 2), thus the difficulty parameter is providing an estimate for ability level when there is a 50% chance that someone at that ability level will endorse either 0 or 1 and 1 or 2. We are using the word endorse because it is general, but since this is an ability measure, we can really think of this as the level in ability needed where there is a 50% change of getting either score in spelling ability. 

The other columns in the table represent factor loadings (lambda) and communality. Essentially, lambda (standardized factor loading) is a correlation that represents how strongly correlated that item is with the latent construct and communality is that value squared, representing how much of the variance in the item is explained by the latent variable. The factor loadings come from mirt also running an EFA on the data! It does it using a quasi-polychoric correlation matrix.

From the information below it looks like item 13 might be problematic since its discrimination is pretty low. We can confirm this with how low the standardized factor loading is. 


```{r Investigating GRM model parameters}
# Obtain the parameter estimates for each item
coefs <- coef(fit_mirt, IRTpars = TRUE, simplify = TRUE) # Saving model parameters (item discrimination (a) and threshold (b)) in a data frame.

# Get the model summary (factor loadings and communality)
mirt_sum <- summary(fit_mirt) # Displaying factor loadings and commonality.

# Create a dataframe from both objects that have item information
item_info <- cbind( coefs$items, mirt_sum$rotF, mirt_sum$h2) %>% data.frame()

# Some data cleaning
rownames(item_info) <- NULL
item_info <- item_info %>% mutate_if(is.numeric, ~round(., 3)) # round
item_info <- cbind(items = names(ARFA_items), item_info) %>% data.frame()
names(item_info) <- c("Item", "a (discrim.)", "b1", "b2", "λ (loading)", "h² (communality)")

# Print the table
Table1A <- item_info %>%
  gt() %>%
  tab_options(
    table.font.size = 11,           # make tighter if needed
    table.width = pct(45)           # helps side-by-side fit
  ) %>%
  tab_header(
    title = "A: Item Parameters"   # Required: your main title
  )


Table1A
```

### Item level statistics for Initital Model

Below is a table showing the item fit statistics. 

```{r Model item level statistics}
# Printing out item level statistics
item.fit <- itemfit(fit_mirt) %>% data.frame() # Estimating item fit statistics and saving them in a data frame.

# Data cleaning
item.fit$item <- names(ARFA_items)
item.fit <- item.fit %>% mutate_if(is.numeric, ~round(., 4))
names(item.fit) <- c("Item", "S-χ²", "df", "RMSEA", "p")

# Print the table 
Table1B <- item.fit %>%
  gt() %>%
  tab_options(
    table.font.size = 11,           # make tighter if needed
    table.width = pct(45)           # helps side-by-side fit
  ) %>%
  tab_header(
    title = "B: Item Fit Statistics"   # Required: your main title
  )


Table1B
```


```{r Removing bad items from ARFA Part 1}
# Bad thresholds
bad_thresh_slope <- 1
bad_thresh_p <- .001

# Identify the good and bad items based of slope
good_items_slope <- item_info$Item[item_info$`a (discrim.)` > bad_thresh]
bad_items_slope <- item_info$Item[!item_info$Item %in% good_items_slope]

# Identify the good and bad items based on item level statistic p-values
good_items_p <- item.fit$Item[item.fit$p > bad_thresh_p]
bad_items_p <- item.fit$Item[!item.fit$Item %in% good_items_p]

# Combine all bad items to get all good items
all_bad_items <- unique(c(bad_items_slope, bad_items_p))
all_good_items <- item_info$Item[!item_info$Item %in% all_bad_items]

# Remove them from the dataset
ARFA_items2 <- ARFA_items[,good_items]

# Rerun the model with the new dataset
mod <- paste0("theta = 1-", length(ARFA_items2))

# Run the mirt model and produce the theta scores
fit_mirt <- mirt(ARFA_items2, 
                 1, 
                 model = mod, 
                 itemtype = "graded", 
                 SE = TRUE, 
                 technical = list(NCYCLES = 20000))
```


```{r Investigating GRM model parameters for good items}
# Obtain the parameter estimates for each item
coefs <- coef(fit_mirt, IRTpars = TRUE, simplify = TRUE) # Saving model parameters (item discrimination (a) and threshold (b)) in a data frame.

# Get the model summary (factor loadings and communality)
mirt_sum <- summary(fit_mirt) # Displaying factor loadings and commonality.

# Create a dataframe from both objects that have item information
item_info <- cbind( coefs$items, mirt_sum$rotF, mirt_sum$h2) %>% data.frame()

# Some data cleaning
rownames(item_info) <- NULL
item_info <- item_info %>% mutate_if(is.numeric, ~round(., 3)) # round
item_info <- cbind(items = names(ARFA_items2), item_info) %>% data.frame()
names(item_info) <- c("Item", "a (discrim.)", "b1", "b2", "λ (loading)", "h² (communality)")

# Print the table
Table1A <- item_info %>%
  gt() %>%
  tab_options(
    table.font.size = 11,           # make tighter if needed
    table.width = pct(45)           # helps side-by-side fit
  ) %>%
  tab_header(
    title = "A: Item Parameters"   # Required: your main title
  )


Table1A
```

### Item level statistics for Initital Model

Below is a table showing the item fit statistics. 

```{r Model item level statistics}
# Printing out item level statistics
item.fit <- itemfit(fit_mirt) %>% data.frame() # Estimating item fit statistics and saving them in a data frame.

# Data cleaning
item.fit$item <- names(ARFA_items2)
item.fit <- item.fit %>% mutate_if(is.numeric, ~round(., 4))
names(item.fit) <- c("Item", "S-χ²", "df", "RMSEA", "p")

# Print the table 
Table1B <- item.fit %>%
  gt() %>%
  tab_options(
    table.font.size = 11,           # make tighter if needed
    table.width = pct(45)           # helps side-by-side fit
  ) %>%
  tab_header(
    title = "B: Item Fit Statistics"   # Required: your main title
  )


Table1B
```


```{r Removing bad items from ARFA Part 2}
# Bad thresholds
bad_thresh_slope <- 1
bad_thresh_p <- .001

# Identify the good and bad items based of slope
good_items_slope <- item_info$Item[item_info$`a (discrim.)` > bad_thresh]
bad_items_slope <- item_info$Item[!item_info$Item %in% good_items_slope]

# Identify the good and bad items based on item level statistic p-values
good_items_p <- item.fit$Item[item.fit$p > bad_thresh_p]
bad_items_p <- item.fit$Item[!item.fit$Item %in% good_items_p]

# Combine all bad items to get all good items
all_bad_items <- unique(c(bad_items_slope, bad_items_p))
all_good_items <- item_info$Item[!item_info$Item %in% all_bad_items]

# Remove them from the dataset
ARFA_items3 <- ARFA_items2[,good_items]

# Rerun the model with the new dataset
mod <- paste0("theta = 1-", length(ARFA_items3))

# Run the mirt model and produce the theta scores
fit_mirt <- mirt(ARFA_items3, 
                 1, 
                 model = mod, 
                 itemtype = "graded", 
                 SE = TRUE, 
                 technical = list(NCYCLES = 20000))
```


We can also get some descriptives for each of the paramters and see something interesting. Mainly for the dispersion of b1 and b2 parameters, which represent difficulty. Again, the lower the number the lower the ability in spelling and vice versa. So based on that information, when looking at b1, we see that it does a pretty good job at capturing low performers in our data. We see that the range is between -5.0450 and 0.1690 with a mean of -1.4943. Thus, items in general are able to capture low performers in the data pretty well. When looking at b2, we see the range is between -1.9290 and 0.8330, with a mean of -0.1300. This should raise some red flags because what this is showing is that none of the items are able to capture above average performers, not even individuals above one standard deviations above the mean! Aka less sensitive

```{r summary of parameter descriptives}
# Get a summary of item info
summary(item_info)
```


### Identifying Model Fit

Next we can use the `M2()` function to get a sense of model fit. We set the type to be equal to C2 because that provides the best results for polytomous models. The M2 statistics is conceptually similar to the chi-square statistics (X2) in SEM, such that the null hypothesis means the model is fitting the data well. Our results are significant, which means it does not- but it is commonly known that a lot of times this test will be significant if the sample size is pretty large, which ours is.

Technically if we were to go off thresholds (even though we are advised not to), the model fit indices produced below show that the model does a good job at capturing the data.

```{r Graded response model fit}
# Show the overall goodness of fit statistic for the model (M statistic, similar to G and Chi-square)
M2(fit_mirt, type = "C2") %>% round(3)
```

### Item level statistics 

Below is a table showing the item fit statistics. 

```{r Model item level statistics}
# Printing out item level statistics
item.fit <- itemfit(fit_mirt) %>% data.frame() # Estimating item fit statistics and saving them in a data frame.

# Data cleaning
item.fit$item <- names(ARFA_items2)
item.fit <- item.fit %>% mutate_if(is.numeric, ~round(., 4))
names(item.fit) <- c("Item", "S-χ²", "df", "RMSEA", "p")

# Print the table 
Table1B <- item.fit %>%
  gt() %>%
  tab_options(
    table.font.size = 11,           # make tighter if needed
    table.width = pct(45)           # helps side-by-side fit
  ) %>%
  tab_header(
    title = "B: Item Fit Statistics"   # Required: your main title
  )


Table1B
```



### Print out TABLE 1

```{r Print out Table1AB}
# Create a Table1B without item var
Table1B_v2 <- item.fit %>%
  select(-Item) %>%
  gt() %>%
  tab_options(
    table.font.size = 11,           # make tighter if needed
    table.width = pct(45)           # helps side-by-side fit
  ) %>%
  tab_header(
    title = "B: Item Fit Statistics"   # Required: your main title
  )

# Print out Table1
wrap_table(Table1A) + wrap_table(Table1B_v2) +
  plot_layout(ncol = 2) +
  plot_annotation(title = "")

```
### Examining the GRM Residuals 

No correlation of residuals exceed 0.2, which means we are good! (No local dependence)

```{r Examining the GRM residuals}
q3 <- residuals(fit_mirt, type = "Q3")

findCorrelation(q3, cutoff = 0.2, verbose = T)
```

### IRT Plots

- Plot1: Item probability functions of ARFA Spelling Subtest
- Plot2: Item information curves of ARFA Spelling Subtest
- Plot3: Test information curve of ARFA Spelling Subtest

```{r Displaying IRT plots}
# Load in stringr package
library(stringr)

# Create a trace plot and save it as p
p <- tracePlot(fit_mirt, title = "A: Item Probability Functions")

# reorder facet variable numerically
item_levels <- unique(p$data$var)
item_levels <- item_levels[order(as.numeric(str_extract(item_levels, "\\d+")))]
p$data$var <- factor(p$data$var, levels = item_levels)

# Print out the figure
p1 <- p +
  labs(color = "Response Categories") +
  scale_color_manual(values = c("#6A0DAD",   # dark/medium purple
                                "#9F7AEA",   # medium-bright purple
                                "#D6BCFA")) +  # light lavender-purple
  theme_clean()




itemInfoPlot(fit_mirt, facet = TRUE, title = "Item Information Functions of the ARFA Spelling Task")
p2 <- testInfoPlot(fit_mirt, title = "B: Test Information Function") + theme_clean()

```

### Print out Figure 1

```{r Print our Figure 1}
# Print out Figure 1
p1 + p2

```

### Showing Relationship Between Raw Sum Scores and EAP Scores (Bayesian IRT Scores)

```{r Showing the relaitonship between raw sum scores and EAP scores}
plot(fit_mirt, type = "EAPsum")

# Extract EAP scores (or use MAP/WLE)
eap_scores <- fscores(fit_mirt, method = "EAP")[, 1]

# Raw summed scores from your response data
raw_scores <- rowSums(ARFA_items, na.rm = TRUE)

# Create the plot
data.frame(raw = eap_scores, theta = raw_scores) %>%
  ggplot(aes(x = raw, y = theta)) +
  geom_point(alpha = 0.4, color = "steelblue") +
  geom_smooth(method = "loess", color = "red", se = TRUE) +
  labs(title = "Relationship between Raw Sum Scores and IRT EAP Theta",
        x = "IRT Ability (EAP Theta)",
       y = "Raw Summed Score") +
  theme_clean()

```



### Running a Partial Credit Model and a 2PL Model for Ability Comparisons

Here we are specifying two additional models, extracting their EAP scores and then correlating them. If the EAP correlations among the three analyses are similar then that means we should not worry too much about how well our items are functioning, overall it is fine.

```{r running models for ability comparison}
# Change the items to be dichotomous
ARFA_items_dich <- data.frame(ifelse(ARFA_items > 1, 1, 0))

# Run a 2PL model
twopl_mod <- mirt(ARFA_items_dich, model = mod, itemtype = '2PL', SE = TRUE, verbose = TRUE)       # See fitting progress

# Run a partial credit model
pcm_mod <- mirt(ARFA_items, model = mod, itemtype = 'Rasch', SE = TRUE, verbose = TRUE)


# model comparison
anova(pcm_mod, fit_mirt)

# Extracting theta values for the models
ARFA$theta_good <- c(fscores(fit_mirt, method = "EAP"))
theta_twopl <- c(fscores(twopl_mod, method = "EAP"))
theta_pcm <- c(fscores(pcm_mod, method = "EAP"))

# Create a dataframe of theta values
theta_df <- data.frame(
  Raw = raw_scores,
  GRM = ARFA$theta_good,
  PCM =theta_pcm,
  `2PL` = theta_twopl
)

# Rename the variables
names(theta_df) <- c("Raw", "GRM", "PCM", "2PL")

# Print a correlation matrix for the theta values
round(cor(theta_df),2) %>%
  kbl() %>%
  kable_paper(full_width = F)
```



# Visualizing the CFIT Data 


### Visualizing the ARFA Spelling Error Data

```{r Visualize the CFIT  data}
# Extract the CFIT data from the full dataset
CFIT_ID <- CFIT4

# Check each item has a correct and incorrect response
# sapply(CFIT, function(x) table(x)) #- all good
# sapply(CFIT, function(x) sum(is.na(x)))

# Rename the dataset (shorten names)
names(CFIT_ID) <- gsub("Item","_",gsub("CFIT", "", names(CFIT_ID)))

# Identify the IDs with no missing data
CFIT_ID_no_na <- CFIT_ID %>% filter(complete.cases(.))

# Calculate the raw scores and within subtest scores
CFIT_df <- data.frame(
  series = rowMeans(select(CFIT_ID_no_na, starts_with("Sub1")), na.rm = T),
  classification = rowMeans(select(CFIT_ID_no_na, starts_with("Sub2")), na.rm = T),
  matrices = rowMeans(select(CFIT_ID_no_na, starts_with("Sub3")), na.rm = T),
  conditions = rowMeans(select(CFIT_ID_no_na, starts_with("Sub4")), na.rm = T),
  RawScores = rowSums(select(CFIT_ID_no_na, - ID))
)

# Generate descriptives of CFIT data
describe(CFIT_df) %>%
  cbind(Tests = names(CFIT_df), .) %>%
  data.frame() %>%
  select(-vars) %>%
  mutate_if(is.numeric, ~round(., 2)) %>%
  gt() %>%
  tab_header(title = "Descriptive Statistics of the CFIT Subtests and Raw Score") %>%
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_body(columns = everything()))


# Number of unique IDs plus min and max scores
unique_Ids <- length(unique(CFIT_ID_no_na$ID))
mean_score <- round(mean(CFIT_df$RawScores),1)
sd_score <- round(sd(CFIT_df$RawScores),1)
skew_score <- round(skew(CFIT_df$RawScores),2)
min_score <- min(CFIT_df$RawScores)
max_score <- max(CFIT_df$RawScores)

# Create a histogram of raw error sum score
CFIT_df %>%
  ggplot(aes(x = RawScores)) +
  geom_histogram(color = "black", fill = "white", bins = 20) +
  labs(title = paste0("CFIT Raw Score Distribution\n(n= ", unique_Ids,
                      "; mean =", mean_score,"; sd = ",sd_score, "; skew = ",skew_score ,"; min = ",min_score,"; max = ",max_score,")")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(x = "CFIT Raw Scores", y = NULL) +
  theme_clean()

# Create a data frame with frequency of error by item
CFIT_item_err_freq <- CFIT_ID_no_na %>%
  pivot_longer(cols = -ID,
               names_to = "Items",
               values_to = "Score") %>%
  group_by(Items, Score) %>%
  summarise(count = length(ID)) %>% 
  mutate(Score = factor(Score)) %>%
  ungroup()

# Create a factor with the levels of interest for plotting purposes
CFIT_item_err_freq <- CFIT_item_err_freq %>%
  mutate(Subtest = substr(Items, 1, 4),
         Items = factor(Items, levels = names(select(CFIT_ID_no_na, - ID))),
         Score = factor(Score, levels = c("1", "0")),
         Subtest = case_when(
           Subtest == "Sub1" ~ "Series",
           Subtest == "Sub2" ~ "Classification",
           Subtest == "Sub3" ~ "Matrices",
           Subtest == "Sub4" ~ "Conditions"))

# Generate a plot
CFIT_item_err_freq %>%
  ggplot(aes(x = Items, y = count, fill = Score)) +
  geom_col() +
  theme_clean() +
  facet_wrap(~Subtest, scales = "free") +
  scale_fill_brewer(palette = "Paired") +
  labs(title = "Frequency of Correct Responses by Items Within Subtests") +
  coord_flip()

```



# Doing Exploratory Factor Analysis on the CFIT Data

Technically a parallel analysis is better but I don't really care tbh. The main takeaway is most of the variance is explained by one factor, but there is evidence in the data present for a few very weak factors, which likely correspond to question similarity, since questions correspond to one of four subtests named: series, classification, matrices, conditions.

```{r Running EFA on the CFIT Data}
# Create a dataset of CFIT data that does not have the ID variable
CFIT_no_na <- select(CFIT_ID_no_na, -ID)

# Create a matrix of polychoric correlations
CFIT_cor <- polychoric(CFIT_no_na)

# Run an exploratory factor analysis with two factors
CFIT_efa2 <- fa(CFIT_cor$rho, nfactors = 2, fm = "minres") # Now, running EFA.
print(CFIT_efa2)

# Create a Scree Plot
data.frame(Eigenvalue = CFIT_efa2$values,
           Factor = 1:length(CFIT_efa2$values)) %>%
  ggplot(aes(x = Factor, y = Eigenvalue)) +
  geom_point(size = 3) +
  geom_line() +
  labs(title = "Scree Plot Using Eigenvalues from Exploratory Factor Analysis (CFIT)") +
  geom_hline(yintercept= 1, linetype="dashed", color = "red") +
  theme_clean()
```

# Running a 2PL, 4-factor 2PL, and Bi-Factor 2PL Model on the CFIT Data

We chose not to fit a hierarhical model because a) it failed to converge (producing an error) and b) theoretically it is not that good of a representation of IQ.

From the model comparison below, the Bi-Factor model produced the best fit (lowest AIC & BIC)

```{r running MIRT on the CFIT data}
# Run a 2PL Model
CFIT_2pl <- mirt(CFIT_no_na, 1, itemtype = "2PL")

# Fitting a 4 factor correlated model
CFIT_4factor <- mirt(
  CFIT_no_na,
  model = "
  F1 = 1-12
  F2 = 13-26
  F3 = 27-38
  F4 = 39-46
  COV = F1*F2, F1*F3, F1*F4,
        F2*F3, F2*F4,
        F3*F4
  ",
  itemtype = "2PL"
)

# Create a vector indicating which specific factor each item belongs to
specific <- c(rep(1, 12),   # items 1-12   → F1
              rep(2, 14),   # items 13-26  → F2
              rep(3, 12),   # items 27-38  → F3
              rep(4, 8))    # items 39-46  → F4

# Fit a bi-factor model to the data
CFIT_bifactor <- bfactor(data = CFIT_no_na, 
                         model = specific,     # this defines the specific factors
                        itemtype = "2PL")


# Compare models to see if bi-factor is better than the 2PL
anova(CFIT_2pl, CFIT_4factor, CFIT_bifactor)
```

# Pring Bi-Factor Model Output Summary (Problematic)

We used the coef() function to produce discrimination (a) slopes for each item that correspond to the general factor and the specific-factor (question type). We do see some problems, in that some items have very very low slopes! (below 1). Thus, we should go ahead and remove these items. 

```{r Print bifactor model summary}
# Print the coefficients
coef(CFIT_bifactor, IRTpars = TRUE, simplify = TRUE)

# Save the item slopes
item_slopes <- coef(CFIT_bifactor, IRTpars = TRUE, simplify = TRUE)$items %>%
  data.frame() %>%
  select(a1:a5) %>%
  round(2)

# Identify good and bad items
good_items <- row.names(item_slopes)[item_slopes$a1 >= 1]
bad_items <- row.names(item_slopes)[item_slopes$a1 < 1]


# Save a dataset that has the parameter information from the bi-factor model
item_pars <- coef(CFIT_bifactor, IRTpars = TRUE, simplify = TRUE)$items %>%
  data.frame() %>%
  select(a1:d)

# Introduce an item variable
item_pars$item <- row.names(item_pars)

# Extract rows that correspond to different subtests and save them as a list
subtest_list <- list(
  Sub1 = item_pars %>% filter(str_detect(item, "^Sub1")),
  Sub2 = item_pars %>% filter(str_detect(item, "^Sub2")),
  Sub3 = item_pars %>% filter(str_detect(item, "^Sub3")),
  Sub4 = item_pars %>% filter(str_detect(item, "^Sub4"))
)

# Change the item info to just numbers
subtest_list <- lapply(subtest_list, function(x) {
  x$item <- as.character(str_split_i(x$item, "_", 2))
  x
})

# making sure the variables are all in the same orientation for each list
subtest_tables <- imap(subtest_list, function(df, subtest_name) {
  df %>%
    mutate(item_num = row_number()) %>%
    select(item_num, item, a1, a2, a3, a4, a5, d) %>%
    rename_with(
      .fn = function(x) paste(subtest_name, x, sep = "_"),
      .cols = -item_num
    )
})

# Combine everything into one wide dataset
wide_table <- reduce(subtest_tables, full_join, by = "item_num") %>% data.frame()

# Some data cleaning
wide_table <- wide_table %>% mutate_if(is.numeric, ~round(., 2))
wide_table[is.na(wide_table)] <- ""
wide_table <- select(wide_table, - item_num)
names(wide_table) <- rep(c("Item\nNum.","a1", "a2", "a3", "a4", "a5", "b"), 4)

# Print the table with the item parameters
wide_table %>%
  kbl() %>%
  add_header_above(c("Series" = 7, "Classifications" = 7, "Matrices" = 7, "Conditions" = 7)) %>%
  kable_paper(full_width = F)

```

### Removing Problematic Items (Low Slopes) and Rerun a Unidimensional and Bi-Factor Model

From the results above, the coefficients for each item, we decided to remove the following items:
- subtest1: 3, 9, 10, 11, 12
- subtest2: 4, 5, 7, 9, 10, 11, 12, 13, 14
- subtest3: 12
- subtest4: 6, 7, 8

Looks like the bi-factor model failed to converge so we will go with the undimensional 2PL version


```{r Removing problematic CFIT Items}
# Remove these items from the data
CFIT_ID_no_na_rmvItms <- CFIT_ID_no_na %>%
  select(- ID, - bad_items)

# Number of remaining items per factor
substr(names(CFIT_ID_no_na_rmvItms), 1, 4) %>% table()

# Run the unidimensional model
CFIT_2pl_rmvItms <- mirt(CFIT_ID_no_na_rmvItms, 1, itemtype = "2PL")

# Fitting a 4 factor correlated model
CFIT_4factor_rmvItems <- mirt(
  CFIT_no_na,
  model = "
  F1 = 1-7
  F2 = 8-12
  F3 = 13-23
  F4 = 24-28
  COV = F1*F2, F1*F3, F1*F4,
        F2*F3, F2*F4,
        F3*F4
  ",
  itemtype = "2PL"
)

# Create a vector indicating which specific factor each item belongs to
specific <- c(rep(1, 7), 
              rep(2, 5),  
              rep(3, 11),   
              rep(4, 5))    

# Fit a bi-factor model to the data
CFIT_bifactor_rmvItems <- bfactor(data = CFIT_ID_no_na_rmvItms, 
                                  model = specific,     # this defines the specific factors
                                  itemtype = "2PL")

# Model Comparison
anova(CFIT_2pl_rmvItms, CFIT_4factor_rmvItems, CFIT_bifactor_rmvItems)
```

### Local Dependence Check on the Unidimensional CFIT Model

There is local dependence for the undimensional model (not good)

```{r Local dependecy check on unidimensional CFIT model}
# Extract the residuals from the models
q3 <- residuals(CFIT_2pl_rmvItms, type = "Q3")

# Identify if residual correlations exceed threshold, suggesting local dependence (Bad)
findCorrelation(q3, cutoff = 0.20, verbose = T)
```


### Further Removing Items with Local Dependence

To address the local dependence problem, we will be removing items based on the results above. Right now the problematic item pairs are: 

Problematic row pairs (correlations)
- row 25 & 26
- row 26 & 24
- row 28 & 6
- row 3 & 4
- row 4 & 8
- row 6 & 7
- row 8 & 10
- row 8 & 9
- row 14 and 13

We will remove the following rows/items:
- Remove: row 26: Sub4_3
- Remove: row 6: Sub1_7
- Remove: row 4: Sub1_5
- Remove: row 8: Sub2_1
- Remove: row 13: Sub3_1


Second round!
There was still local depedence after the initial removal of items above, so we also removed these items as well:
- Remove: Sub4_1
- Remove: Sub4_2
- Remove: Sub1_4


```{r dropping local dependent items in CFIT}
# Create a data frame to quickly index what item number correspond to what items
item_num <- data.frame(num = 1:length(CFIT_ID_no_na_rmvItms), cbind(names(CFIT_ID_no_na_rmvItms)))


# Drop redundant items
CFIT_ID_no_na_rmvItmsv2 <- CFIT_ID_no_na_rmvItms %>%
  select(-Sub4_3,- Sub1_7, -Sub1_5, -Sub2_1, - Sub3_1)

# Drop some more items (these items also showed local dependence after removal of above items)
CFIT_ID_no_na_rmvItmsv3 <- CFIT_ID_no_na_rmvItmsv2 %>%
  select(-Sub1_4 ,- Sub4_1, -Sub4_2)

# Fit the 2PL model on new dataset
CFIT_2pl_rmvMore <- mirt(CFIT_ID_no_na_rmvItmsv3, 1, itemtype = "2PL")

# Identify if residual correlations exceed threshold, suggesting local dependence (Bad)
q3 <- residuals(CFIT_2pl_rmvMore, type = "Q3")
findCorrelation(q3, cutoff = 0.2, verbose = T)

```

### Print Unidimensional 2PL Model Parameters (For Publication)

```{r Print the 2PL Model Parameters of Final Unidimensional Model}
# Print the coefficients (slopes and thresholds) for each item
item_pars2 <- coef(CFIT_2pl_rmvMore, IRTpars = TRUE, simplify = TRUE)$items %>%
  data.frame() %>%
  mutate_if(is.numeric, ~round(., 2)) 

# Introduce an item variable
item_pars2$item <- row.names(item_pars2)

# Extract rows that correspond to different subtests and save them as a list
subtest_list2 <- list(
  Sub1 = item_pars2 %>% filter(str_detect(item, "^Sub1")),
  Sub2 = item_pars2 %>% filter(str_detect(item, "^Sub2")),
  Sub3 = item_pars2 %>% filter(str_detect(item, "^Sub3")),
  Sub4 = item_pars2 %>% filter(str_detect(item, "^Sub4"))
)

# Change the item info to just numbers
subtest_list2 <- lapply(subtest_list2, function(x) {
  x$item <- as.character(str_split_i(x$item, "_", 2))
  x
})

# making sure the variables are all in the same orientation for each list
subtest_tables2 <- imap(subtest_list2, function(df, subtest_name) {
  df %>%
    mutate(item_num = row_number()) %>%
    select(item_num, item, a, b) %>%
    rename_with(
      .fn = function(x) paste(subtest_name, x, sep = "_"),
      .cols = -item_num
    )
})

# Combine everything into one wide dataset
wide_table2 <- reduce(subtest_tables2, full_join, by = "item_num") %>% data.frame()

# Some data cleaning
wide_table2 <- wide_table2 %>% mutate_if(is.numeric, ~round(., 2))
wide_table2[is.na(wide_table2)] <- ""
wide_table2 <- select(wide_table2, - item_num)
names(wide_table2) <- rep(c("Item\nNum.","a", "b"), 4)

# Print the table with the item parameters
wide_table2 %>%
  kbl() %>%
  add_header_above(c("Series" = 3, "Classifications" = 3, "Matrices" = 3, "Conditions" = 3)) %>%
  kable_paper(full_width = F)


```


### Additional CFIT Unidimesional Model Checks

```{r Additional CFIT Unidimensional Model Checks}
# Print an omnibus test of the model
M2(CFIT_2pl_rmvMore, type = "C2")

# Print out the item fit statistics
itemfit(CFIT_2pl_rmvMore) %>%
  data.frame() %>%
  mutate_if(is.numeric, ~round(., 4)) 

# Use the correct plot functions to plot Item Information Function and Reliability
testInfoPlot(CFIT_2pl_rmvMore, title = "B: Test Information Function")
conRelPlot(CFIT_2pl_rmvMore, title = "Reliability of the CFIT Data")

# Creating the information curve by hand
info_curve <- data.frame(
  G = seq(-4, 4, by =.10)) %>%
  mutate(info = testinfo(CFIT_2pl_rmvMore, G),
         se = 1/sqrt(info),
         relia = info/(info + 1))

# Convert dataset to long- remove se associated with reliability
info_curve_long <- info_curve %>%
  pivot_longer(cols = c(info,relia))
info_curve_long$se[info_curve_long$name == "relia"] <-NA

# Make it prettier
info_curve_long <- mutate(info_curve_long, name = ifelse(name == "info", "Information", "Reliability"))

# Plot these curves side by side
info_curve_long %>%
  ggplot(aes(x= G, y = value)) +
  facet_wrap(~name, scales = "free") +
  geom_line(aes(y = se), linetype = "dashed") +
  geom_line() +
  labs(y = NULL, x = "Fluid Intelligence") +
  theme_clean()
  
```


  
# Extracing the Bayesian IRT Scores from the 2PL Model


```{r Extracting CFIT Scores from 2PL Model}
# Extract the Bayesian EAP scores from the model
CFIT_scores <- fscores(CFIT_2pl_rmvMore, method = "EAP")

# Convert into a data.frame and reintroduce IDs to data
CFIT_scores_df <- data.frame(cbind(ID = CFIT_ID_no_na$ID, CFIT_scores))

# Introduce the scores only from the G factor back into the data
full_data <- data %>%
  left_join(CFIT_scores_df, by = "ID")

# Save only variables of interest
full_data <- full_data %>%
  select(S1:neg_feel, G = F1)

# Get descriptives of the outcome
describe(full_data$G)

# Generate a histogram for cognitive latent variable EAP scores
full_data %>%
  ggplot(aes(x = G)) +
  geom_histogram(color = "black", fill = "white", bins = 20) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(x = "Fluid Intelligence (θ)", y = NULL) +
  theme_clean()

# Low IQ peeps
sum(full_data$G <= -1.5, na.rm = T)
```






### Loading EEG Information 

Here we are loading two types of EEG information:

- Individual power spectra (FFT) of rsEEG recordings
- An EEG quality control (QC) CSV 

The **power spectra** holds information of our resting-state EEG outcome. The analysis in subsequent sections will be attempting to explain the variance of this for the several frequency bands. The QC will be used to remove recordings that were problematic during the preprocessing stage- this occurs in the next section.

During this data cleaning code section, we are:

- Identifying two rsEEG recordings that need to be renamed
- Removing data from 5 recordings, one which was redundant, four because they were not real IDs (not sure what they are)
- Use interpolated channel x trial segments to identify channels as 'bad' (e.g., if 90% of the channel was interpolated then we should mark that as a bad channel)

The information in this section should **not** count toward reporting 'x' number of IDs removed

```{r Load and preprocess EEG data}
# Get the pathways to where the EEG data is saved
eegPath <- "\\\\files.times.uh.edu/labs/MIR_Lab/MEGAGRANT/STUDY 1/FINAL_DS/EEG/rsEEG"

# Get all file names and create an empty list
all_fft_files <- list.files(path = file.path(eegPath, "fft"), pattern = "fft")
all_fft_files_full <- file.path(eegPath, "fft", all_fft_files)
all_files_list <- list()

# Read in each file one by one and save within a list
for(ii in 1:length(all_fft_files)){
  # Read in the file
  read_file <- read.csv(all_fft_files_full[ii])
  # Saves the file name within the file
  read_file$filename <- all_fft_files[ii]
  # Save the recording into a list
  all_files_list[[ii]] <- read_file
  names(all_files_list)[ii] <- all_fft_files[ii]
}

# Combine the FFT files into one dataset
FFT_dat <- do.call(rbind, all_files_list)

# Create a function that accurately extracts the ID of the file
clean_eeg_ids <- function(x) {
  x_gsub <- gsub("_preproc_fft.csv|EyesClosed|_RAW|_ RAW|RAW_|_", "", x)
  return(x_gsub)
}

# Recover the ID for each recording + specify condition
FFT_dat <- FFT_dat %>%
  mutate(ID = clean_eeg_ids(filename),
         Condition = ifelse(grepl("Open", filename), "Open", "Closed"))

# load in the EEG QC
EEG_QC <- read.csv(file.path(eegPath, "EEG_Preprocessing_Summary_Statistics_EC.csv"))

# Delete weird column that has duplicate from one file (idk how that happened maybe parallel processing bug)
#EEG_QC <- select(EEG_QC, - s14_eegvarsum_chanvar, - s14_eegvarsum_globalvar)

# Use the same cleaning as above on this data
EEG_QC <- EEG_QC %>%
  mutate(ID = clean_eeg_ids(subject),
         Condition = ifelse(grepl("Open", subject), "Open", "Closed"))

# Join the datasets together
EEG_dat <- FFT_dat %>%
  full_join(EEG_QC, by = c("ID", "Condition"))


# Read in the EEG mismatch dataset
eeg_mismatch <- read_excel(file.path(eegPath, "MegaGrant_TBL.xlsx"))

# Data cleaning
eeg_mismatch <- select(eeg_mismatch, ID, RAW)
eeg_mismatch <- drop_na(eeg_mismatch, RAW)

# Identify which rsEEG recordings need to have their ID renamed
newEEGIDs <- setdiff(eeg_mismatch$RAW, eeg_mismatch$ID)
newEEGIDs_df <- eeg_mismatch[eeg_mismatch$RAW %in% newEEGIDs,]
cat("rsEEG IDs that need to be renamed (RAW -> ID)\n\n")
newEEGIDs_df

# Change the RAW eeg IDs to their correct names
mapping <- setNames(newEEGIDs_df$ID, newEEGIDs_df$RAW)
EEG_dat <- EEG_dat %>%
  mutate(ID = if_else(ID %in% names(mapping), 
                      mapping[ID], 
                      ID))

# Checking to see if it worked
cat("It it should say 0 below if it worked and IDs were renamed\n\n")
sum(newEEGIDs_df$RAW %in% unique(EEG_dat$ID))

# Create new variables for the EEG dataset
EEG_dat$trialsrmv <- EEG_dat$s5_rjctbadtrials_initialtrials - EEG_dat$endTrialnum # This is accurate
EEG_dat$trialpropremain <- round(EEG_dat$endTrialnum/EEG_dat$s5_rjctbadtrials_initialtrials,3)

# EEG Cleaning- keeping only the variables that we care about 
EEG_dat2 <- EEG_dat %>% 
  select(ID, 
         Condition, 
         start_trial = s5_rjctbadtrials_initialtrials,
         start_globvar = s2_eegvarsum_globalvar,
         trialsrmv,
         end_trial = endTrialnum,
         trialpropremain,
         end_globvar = s10_eegvarsum_globalvar,
         int_total,
         channel,
         frequency,
         power,
         elec_pop = s4_rmvbadchan_pop_chan,
         elec_flat = s4_rmvbadchan_flat_chan,
         elec_nois = s4_rmvbadchan_noise_chan,
         int_AF3:int_TP8)

# Convert data to long format to get number of channels removed (80% threshold)
chan_rmvthresh <- .90

# Obtain the number of removed channels
EEG_dat2$chan_rmv <- rowSums(select(EEG_dat2, int_AF3:int_TP8) > chan_rmvthresh)
EEG_dat2$chan_rmv2 <- rowSums(select(EEG_dat2, int_AF3:int_TP8) > .99)

# Dropping weird EEG recordings (64622 (2) is a duplicate of 64622- same data)
EEG_dat2 <- EEG_dat2 %>% 
  filter(!ID %in% c("64622 (2)",  # Redundant recording
                    "9754341", # Not a real ID
                    "Рюмина Е.Е.",  # Not a real ID
                    "3973",  # Not a real ID
                    "8222"  # Not a real ID
                    ))

# Convert the ID variable into numeric
EEG_dat2$ID  <- as.numeric(EEG_dat2$ID)
```


### Making the Second Dataset Dependent on ARFA Subject Number

Before loading in the EEG data, all of those previous sections were just following through with our first analysis, modeling spelling ability for each subject. We are now using the information from the model to produce spelling ability IRT scores, and thus our dataframe will begin there, with those same IDs used. In other words, even though we have a dataset will all IDs in S1 and S3, we will not be using that information in its entirely, instead we will report values conditional on having data for ARFA.

Okay so below we will be combining three datasets

- ARFA (contains spelling data)
- data (contains demographics and medical questionnaires)
- EEG_dat2 (contains EEG data and EEG QC)

```{r Making the second dataset dependent on ARFA}
# Combining datasets together
dataL <- select(ARFA, ID, theta_good) %>% # Choose few vars of interest
  left_join(full_data, by = "ID") %>%
  left_join(EEG_dat2, by = "ID")

# Combine this dataset to our other dataset with behavioral information
cat("The number of unique IDs in this now combined dataset with ARFA, Demo, Medical, and EEG data is", length(unique(dataL$ID)))

```


### Remove subjects that do not meet criteria (Behaviorally)

Here we will be removing subjects that do not have the data for the final analysis and that do not meet the behavior criteria. There is another section that will follow where we further exclude subjects who do not have adequate EEG data.

Data needed:
- Must have age data
- Must have spelling data
- Must have CFIT data
- Must have EEG data

This is the exclusion criteria:
- Have a CFIT IRT Score less than 1.5 SD 
- Self-report head trauma
- Epilepsy
- Autism
- ADD
- Reported 'Very dissatisfied (1)' to sleep quality in the last two weeks
- Reported 'Always (1)' to experiencing negative feelings in the last two weeks

*Update- modification occurred, we will now keep dyslexics
*Update - modification occurred, decided to keep in participants that are too old (3SD from the mean)*

We use the `slice_head()` function since this is information from level-2 predictors- this will speed up the code. 


```{r keep non rejects subjects}
# Drop rows that are missing data age, spelling, or EEG
dataL <- dataL %>%
  drop_na(theta_good, Age, G, power)

# Slice the dataset so there is one row per subject
sliced_data <- dataL %>%
  select(ID:ADD, sleep:G) %>%
  group_by(ID) %>%
  slice_head(n=1) %>%
  ungroup()

# Identify those at the bottom 15% percentile
sliced_data$G_low <- ifelse(sliced_data$G <= 1.5, "Y", "N")

# Identify those with sleep problems
sliced_data$sleep_prob <- ifelse(sliced_data$sleep == 1, "Y", "N")

# Identify those with anxiety
sliced_data$anx_prob <- ifelse(sliced_data$neg_feel == 1, "Y", "N")

# Scale some variables to later remove extremes
#sliced_data$Agez <- as.vector(scale(sliced_data$Age))

# Create categorical variations of the variables above
#sliced_data$Too_old <- ifelse(sliced_data$Agez >=3 , "Y", "N")

# Create a variable indicating if behavioral data is missing or not
sliced_data$MissingSpell <- ifelse(is.na(sliced_data$theta_good), "Y", "N")

# Create a variable indicating if behavioral data is missing or not
sliced_data$MissingAge <- ifelse(is.na(sliced_data$Age), "Y", "N")

# Create a variable indicating if CFIT data is missing
sliced_data$MissingCFIT <- ifelse(is.na(sliced_data$G), "Y", "N")

# Create an exclusion criteria for the behavioral data (Before accounting for noisy EEG data)
sliced_data <- sliced_data %>%
  mutate(
    # Exclusion1 = head injury or developmental disorder
    Exclusion1 = case_when(
      HeadTrauma == "Y" |
      Health2epilepsy == "Y" |
      Health2autism == "Y" |
      ADD == "Y" ~ "Y",
      TRUE ~ "N"
    ),
    
    # Exclusion2 = Missing between-subject data
    Exclusion2 = case_when(
      MissingSpell == "Y" |
      MissingAge == "Y" |
      MissingCFIT == "Y" ~ "Y",      
      # Too_old == "Y" ~ "Y",        
      TRUE ~ "N"
    ),
    
    # Exclusion3 = Reported negative feelings or bad sleep
    Exclusion3 = case_when(
      sleep_prob == "Y" |
      anx_prob == "Y" ~ "Y",
      TRUE ~ "N"
    ),
    
    # Exclusion4 = IQ scores in the bottom 15th percentile
    Exclusion4 = case_when(
      G_low == "Y" ~ "Y",
      TRUE ~ "N"
    ),
    
    # Final Exclusion variable
    Exclusion = case_when(
      Exclusion1 == "Y" | 
      Exclusion2 == "Y" | 
      Exclusion3 == "Y" | 
      Exclusion4 == "Y" ~ "Y",
      TRUE ~ "N"
    )
  )


# Indicate our starting sample size
cat("After removing participants that were missing data for ARFA, Age CFIT, or EEG, we have",length(unique(sliced_data$ID)),"unique IDs\n")

# Express how many subjects are being removed for injury, epilepsy, or autism
cat("We are removing",sum(sliced_data$Exclusion1 == "Y", na.rm = T),"participants for having head trauma, epilepsy, autism or ADD\n")

# Express how many subjects are being removed for having sleep or anxiety problems
cat("We are removing",sum(sliced_data$Exclusion3 == "Y", na.rm = T),"participants for having consistently bad sleep or anxiety problems in the past 2 werks\n")

# Create a table specifically for excluded subjects to clean and then report numbers
Excluded <- sliced_data %>% filter(Exclusion == "Y") %>% 
  select(HeadTrauma:ADD, G_low:anx_prob) # We included dislexia here to report the numbers
Included <- sliced_data %>% filter(Exclusion != "Y") 
  
# Clean the table a bit (convert NAs into N)
Excluded[is.na(Excluded)] <- "N"

# Quickly look at the frequency of responses for this dataset
sapply(Excluded, function(x) table(x))

# Report the number of unique IDs we have left
cat("This leaves us with data from",nrow(Included),"participants who may or may not have an EEG recording\n")

# Create a dataset that contains only participants that were not excluded
data2L <- dataL %>%
  filter(ID %in% Included$ID)

# Everything checks out
setdiff(Included$ID, data2L$ID)
setdiff(data2L$ID, Included$ID)
```

### Early Demographics

The demographics before we removed bad EEG recordings
- The sample size for Age is three less than it should be because of NAs

```{r initial demographics info}
# slice the data again
sliced_data_demo <- data2L %>%
  group_by(ID) %>%
  slice_head(n=1) %>%
  ungroup()

describe(sliced_data_demo$Age)
round(prop.table(table(sliced_data_demo$Sex)),2)
table(sliced_data_demo$Group)
paste0(length(unique(sliced_data_demo$ID)), " unique IDs")

```


### Descriptives of QC of our EEG data (Eyes Closed Only)

Below we can visualize different EEG QC measures to get a feel for how problematic EEG recordings were and in which aspect. These QC metrics were derived from MATLAB preprocessing code where custom functions were used to identify number of bad channels, trials, and overall data interpolation. With the exception of bad trials that used a threshold of SD =4, every other metric used a threshold of SD = 5. Specifically, this was using robust zscores since they are less likely to be inflated by skewed data.

```{r descriptives of our EEG QC}
# Aesthetics (Leo version)
theme_clean <- function() {
  theme_minimal() +
    theme(legend.position = "bottom",
          panel.grid.minor = element_blank(),
          plot.title = element_text(hjust = 0.5))
}

plot_hist <- function(data, x){
data %>%
  ggplot(aes(x={{x}})) +
  geom_histogram(fill = "grey", color = "black") +
  theme_clean()
 
}

plot_bar <- function(data, x){
EEG_QC_rec %>%
  group_by({{x}}) %>%
  summarise(Freq = length({{x}})) %>%
  drop_na() %>%
  ggplot(aes(x = {{x}}, y = Freq)) +
  geom_col(fill = "grey", color = "black") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) +
  theme_clean()
 
}

# Create a dataset where each row represents data from one condition recording
EEG_QC_rec <- data2L %>%
  group_by(ID) %>%
  slice_head(n = 1) %>%
  data.frame()

# Create histograms plots of the different QC measures
plot_hist(EEG_QC_rec, start_globvar) + labs(x = "Before Preprocessing Global Variance")
plot_bar(EEG_QC_rec, chan_rmv) + labs(x = "Number of Noisy Channels Interpolated")
plot_hist(EEG_QC_rec, trialpropremain) + labs(x = "Proportion of Trials Remaning")
plot_hist(EEG_QC_rec, end_globvar) + labs(x = "After Preprocessing Global Variance")
plot_hist(EEG_QC_rec, int_total) + labs(x = "Proportion of Recording Interpolated")
plot_hist(EEG_QC_rec, end_trial) + labs(x = "Number of Trials After Preprocessing\n(4 Second Trials)")
```

#### Identifying how many EEG recordings that are good 

Below we created the exclusion criteria for EEG recordings that will not be analyzed.

Below we specified what makes a good recording. The recordings need to:

- Have less than 18% of channels be interpolated or marked as bad. Since there are 62 channels in each recording, 62 * .18 = 11.16. Thus, if a recording has 12 or more bad channels they will not be used for analysis.
- Have at least 30 trials, each being 4 seconds long, which will result in 2 minutes worth of EEG recordings.
- Have missing EEG data, they will be fully removed.
- And if their ending global variance is larger than 4 robust z-scores then they will be removed. This means that we calculate the MAD score instead of the standard deviation, which is more robust to outliers, and multiply this value by 4 and add it to the mean of ending global variance. Any recordings with an ending global variance that exceeds this threshold will be removed.

```{r Identifying Good rsEEG Recordings}
# Create to identify included vs excluded recordings
apply_exclusion_criteria <- function(df, chan_num, end_trial, chan_rmv_prop) {
  df$Not_Engh_EEG <- ifelse(df$end_trial < end_trial, "Y", "N")
  df$Many_Chan_Intp <- ifelse(df$chan_rmv > round(chan_rmv_prop * chan_num), "Y", "N")
  df$Missing_EEG <- ifelse(is.na(df$power), "Y", "N")
  df$HigEndVar <- ifelse(df$end_globvar > mean(df$end_globva, na.rm  = T) + 4 * mad(df$end_globva, na.rm = T), "Y", "N")
  
  df %>% mutate(Exclusion = case_when(
    Not_Engh_EEG == "Y"  | Many_Chan_Intp == "Y"  | Missing_EEG == "Y" | HigEndVar == "Y" ~ "Y",
    TRUE ~ "N"
  ))
}

# Set parameters for exclusion criteria
channels <- 62
trials_needed <- 30
chann_int_prop <- .18

# Apply our exclusion criteria to our data
EEG_QC_rec2 <- apply_exclusion_criteria(EEG_QC_rec,
                                        channels,
                                        trials_needed,
                                        chann_int_prop)


# This seems reasonable
cat("Recordings split by whether they have enough trials (min =",trials_needed,")\n")
xtabs(~Not_Engh_EEG, EEG_QC_rec2)
cat("\n\nRecordings split by whether they exceed acceptable channel interpolation proportion from ",channels," channels (max =",chann_int_prop,")\n")
xtabs(~Many_Chan_Intp, EEG_QC_rec2)
cat("\n\nHad an ending global variance that exceeded 4 MAD above the mean\n")
xtabs(~HigEndVar, EEG_QC_rec2)
cat("\n\nIDs that did not have EEG data at all\n")
xtabs(~Missing_EEG, EEG_QC_rec2)
cat("\n\nNumber and percentage of recordings that need to be excluded\n")
xtabs(~Exclusion, EEG_QC_rec2)
round(prop.table(xtabs(~Exclusion, EEG_QC_rec2)),2)

# Remove the files that are marked for exclusion
EEG_QC_rec3 <- filter(EEG_QC_rec2, Exclusion == "N")

# Check for missing data
cat("\n\nMissing data of the observations (IDs) that will be used in the final analysis\n")
sapply(EEG_QC_rec3, function(x) sum(is.na(x)))

```


# Introducing Topography and Frequency Band Variables

We will start with the information present in the data frame that was created in the previous section. We will use the IDs that basically 'survived' for having adequate behavioral and EEG data and use that as an index for the dataset that contains all of the within-individual (group) variables (e.g. channels, frequency).

Here we will remove variables (columns) that are not much of interest to us and introduce frequency band (FB) and topography as factors. These factors are designed to simply our data, since right now each recording has power values across individual frequencies (this is also called the **Power Spectra Density**) for each of the 62 electrodes. Eventually we will use these factors to reduce the dimensions of the data, but for now we used the `summary()` function in combination with the `group_by()` function to keep the same dimension. This is so that in subsequent sections we can plot the power spectra without issues.  

Also, all recordings below are **eyes closed** only!

The frequency bands chosen were based on an extensive review of frequency bands looked at across several papers (check the rsEEG power changes with development excel sheet).

- Delta waves: 1 - 4 Hz
- Theta waves: 4 - 7.5 Hz
- Alpha1 (slow) waves: 8 - 10 Hz
- Alpha2 (fast) waves: 10 - 12.5 Hz
- Beta1 waves: 12.5 - 18 Hz
- Beta2 waves: 18 - 34.5 Hz

Why use these different frequency bands? What do they represent in terms of informing us about the brain?

Delta: 

Delta is the clearest developmental marker in your table. Neurophysiologically, high delta power reflects large-amplitude, highly synchronous slow activity across neuronal populations. In development, elevated delta is interpreted as a marker of neural immaturity, greater synaptic density, and less refined cortical organization. Whitford et al. specifically frame slow-wave power as arising from spatially coherent synaptic activity, especially local synchrony among pyramidal neurons; as synaptic pruning reduces redundant dendrites, spines, and axon terminals, scalp EEG amplitude decreases. This makes delta especially sensitive to adolescent cortical maturation. See rows 8 and 9, and also rows 3, 4, and 6.
Expected change: delta absolute power should decrease from adolescence into adulthood, with the largest and most robust declines compared with faster bands. This decrease should be nonlinear/decelerating, with stronger change during adolescence and a plateau/asymptote by the mid-20s. This is especially supported by rows 8 and 9. Topographically, the developmental decrease is often widespread, but several rows suggest posterior regions mature earlier and frontal/association cortices show later maturation. See rows 2, 3, 4, 8, and 9.

Theta:

Theta also reflects slow-wave developmental immaturity, but it is often discussed as slightly more cognitively/network-related than delta. In your rows, theta is tied to maturation of thalamo-cortical and cortico-cortical systems, reduction of slow-wave dominance, and the shift toward faster dominant frequencies. Cragg et al. describe delta/theta decreases as part of standard adolescent maturation and connect them to synaptic pruning and white matter development. See rows 4, 6, 8, and 9.
Expected change: theta absolute power should decrease from adolescence into adulthood, often strongly and sometimes even more steeply than delta depending on the study. Gasser et al. note that theta showed one of the strongest developmental decreases and that theta is gradually superseded by alpha activity during development. See row 2. Cragg et al. also report that delta and theta show the largest developmental decreases, with reductions across the scalp. See row 4.

Alpha 1 (Slow):

Alpha1 is lower alpha, and developmentally it often behaves more like the slow bands than like mature adult alpha. In your table, alpha1 is described as decreasing with age or being reduced as the brain shifts from slower alpha toward faster alpha. Gasser et al. describe alpha1 as part of a maturational sequence where alpha1 supersedes theta, and alpha2 later supersedes alpha1. See row 2. Cragg et al. report that alpha1 absolute power decreases over time. See row 4. Lüchinger et al. also report adolescent-adult differences below 10 Hz, which maps well onto alpha1. See row 8.
Expected change: alpha1 should generally decrease from adolescence into adulthood, especially if measured as absolute power. Conceptually, this reflects movement away from slower, less mature dominant rhythms and toward faster alpha organization. It is also part of the broader developmental “frequency shift” from slower to faster rhythms. See rows 2, 4, 6, and 8.

Alpha2 (Fast):

Alpha2 is the more mature upper-alpha range. It is closely connected to the adult posterior alpha rhythm and peak alpha frequency. Neurophysiologically, alpha is often linked to thalamo-cortical information exchange, cortical idling/inhibition, and coordinated rhythmic activity in posterior networks. In your sheet, alpha2 is also tied to white matter maturation and increased conduction velocity, because faster neural communication supports faster rhythmic oscillations. See rows 4 and 6.
Expected change: alpha2 is not expected to simply decrease like delta/theta. It may increase in relative power during development, especially posteriorly, even when absolute power is stable or eventually decreases. Gasser et al. describe alpha2 as increasing until around age 13 and then decreasing, producing an inverted-U pattern. See row 2. Cragg et al. report that alpha2 was the only range that did not show an age-related absolute power decrease, while relative alpha2 increased posteriorly. See row 4. For adolescents into adulthood, I would expect alpha2 to be relatively stable or show weaker decreases than slow bands, while peak alpha frequency shifts upward and posterior alpha organization becomes more adult-like. See rows 4, 6, and 8.

Beta1:

Beta1 reflects faster, lower-amplitude cortical activity and more mature cortical network dynamics. Compared with delta/theta, its developmental interpretation is less straightforward. Some rows describe beta as part of the broader shift toward faster rhythms with maturation, especially in relative power, while absolute beta may decrease slowly or show weaker age effects. See rows 2, 4, 6, and 8.
Expected change: beta1 absolute power may decrease slightly with age, but the effect is usually weaker and less robust than delta/theta. In relative terms, beta may increase because slow-wave power decreases more strongly. Gasser et al. found beta1/beta2 absolute power slowly declined, while fast bands increased in relative magnitude. See row 2. Cragg et al. found beta absolute power decreased with age, but relative beta increased. See row 4. Lüchinger et al. found higher beta frequencies did not clearly distinguish adolescents from adults. See row 8. So for your project, beta1 should be treated as a secondary/more exploratory maturational marker.

Beta2

Beta2 reflects higher-frequency, lower-amplitude activity, but it is also more vulnerable to contamination from muscle activity than lower bands. Your sheet notes caution about higher beta/gamma ranges and muscle artifacts, especially above 20–30 Hz. See row 4, and row 2 notes avoiding very high beta/gamma contamination.
Expected change: beta2 is the least clear developmental band among the six. Some studies report slow decreases in absolute beta with age, some report increases in relative fast-frequency power, and others find weak adolescent-adult differences. Soroko et al. even found beta2 increases in some groups/regions but decreases in others, suggesting strong topographic and subgroup specificity. See row 6. Lüchinger et al. found beta1/beta2 did not substantially distinguish adolescents from adults. See row 8. I would include beta2, but interpret it cautiously and emphasize that developmental effects are expected to be weaker and less consistent than delta/theta.


For topography, we will use channel information from 56 channels to create essentially 8 regions, that capture brain activity in the frontal, central, temporal, parietal and occipital regions for both hemispheres (we will ignore midline since we have no theoretical implications for it).

Left Frontal: c("Fp1","AF3","AF7","F1","F3","F5","F7","FC1","FC3","FC5")
Right Frontal: c("Fp2","AF4","AF8","F2","F4","F6","F8","FC2","FC4","FC6")
Left Central: c("C1","C3","C5","CP1","CP3","CP5") 
Right Central: c("C2","C4","C6","CP2","CP4","CP6")
Left Temporal: c("FT7","T7","TP7","FT9")
Right Temporal: c("FT8","T8","TP8","FT10")
Left Parietal: c("P1","P3","P5","P7") 
Right Parietal: c("P2","P4","P6","P8")
Left Occipital: ("PO3","PO7","O1")
Right Occipital: c("PO4","PO8","O2")

paste(Frontal, collapse = ", ")


```{r Look at the power spectra of included recordings}
# Good recording IDs + removing unwanted variables
data3L <- data2L %>%
  filter(ID %in% EEG_QC_rec3$ID) %>%
  select(ID, Sex, Age, Group, theta_good, Condition, channel, frequency, power) %>%
  arrange(ID, frequency)

# Creating frequency bands variable
data4L <- data3L %>%
  unique() %>%
  mutate(
    FB = case_when(
      frequency >= 1 & frequency <= 4 ~ "delta",
      frequency >= 4 & frequency <= 7.5 ~ "theta",
      frequency >= 8 & frequency <= 10 ~ "alpha1",
      frequency >= 10 & frequency <= 12.5 ~ "alpha2",
      frequency >= 12.5 & frequency <= 18 ~ "beta1",
      frequency >= 18 & frequency <= 34.5 ~ "beta2",
      TRUE ~ "ERROR"
    ),
    FB = factor(FB, levels = c("delta", "theta", "alpha1", "alpha2", "beta1", "beta2"))
    )

# Creating a topography_hemisphere variable (These are NON LONGER from Whitford's paper 2007)
# Creating a topography_hemisphere variable (updated 8-cluster solution)
data4L <- data4L %>%
  mutate(
    topography_hemisphere = case_when(
      # Frontal
      channel %in% c("Fp1","AF3","AF7","F1","F3","F5","F7","FC1","FC3","FC5") ~ "left_frontal", 
      channel %in% c("Fp2","AF4","AF8","F2","F4","F6","F8","FC2","FC4","FC6") ~ "right_frontal",
      # Central
      channel %in% c("C1","C3","C5","CP1","CP3","CP5") ~ "left_central",
      channel %in% c("C2","C4","C6","CP2","CP4","CP6") ~ "right_central",
      # Temporal
      channel %in% c("FT7","T7","TP7","FT9") ~ "left_temporal",
      channel %in% c("FT8","T8","TP8","FT10") ~ "right_temporal",
      # Parietal
      channel %in% c("P1","P3","P5","P7") ~ "left_parietal",
      channel %in% c("P2","P4","P6","P8") ~ "right_parietal",
      # Occipital
      channel %in% c("PO3","PO7","O1") ~ "left_occipital",
      channel %in% c("PO4","PO8","O2") ~ "right_occipital",
      
      # Handles NAs
      TRUE ~ NA_character_
    )
  )

# Creating a topography and hemisphere variable separately
data4L <- data4L %>%
  mutate(
    topography = case_when(
      grepl("frontal", topography_hemisphere) ~ "frontal",
      grepl("temporal", topography_hemisphere) ~ "temporal",
      grepl("central", topography_hemisphere) ~ "central",
      grepl("parietal", topography_hemisphere) ~ "parietal",
      grepl("occipital", topography_hemisphere) ~ "occipital"),
    hemisphere = case_when(
      grepl("left", topography_hemisphere) ~ "left",
      grepl("right", topography_hemisphere) ~ "right"),
    topography = factor(topography),
    hemisphere = factor(hemisphere)
  )
```

# Plotting the Power Spectra Density (Averaging Across Channels)

Below we are plotting the **power spectral density (PSD)** of our data. This basically means that we have power values (uV^2) for different frequencies (Hz), that's it, there is nothing more complicated that goes to this. For a source that corroborated this, please visit:

https://liquidinstruments.com/blog/what-is-power-spectral-density-and-why-is-it-important/

Anyway, the PSD is different from **absolute power**, where we take PSD and then sum them within frequency bands. Note that the correct way to calculate absolute power is by taking the sum and not the mean!

If we wanted to calculate relative power, essentially, or the proportion of power explained by that frequency, then we can sum up the power values across the range of our frequencies (0.5 - 35 Hz by 0.25 Hz, which is 139 frequencies) and then just divide each PSD by that sum. The reason why we have power values corresponding to 0.25 Hz is because our EEG segments are 4 seconds long, thus the **frequency resolution**, calculated by 1/seconds = 1/4 seconds = 0.25. 

Anyway, thinking about our full long dataset, we have power values that correspond to a range of frequencies (139). Every recording has this information for 62 channels (62 * 139 = 8,618). This is too much information to plot. So instead what we can do is start by averaging across channels within each person, so now every will have a PSD comprised of 139 values that represent the overall PSD. Anyway to accomplish this we can take our long dataset and group it by ID and frequency and obtain an average power across channels- this will be our starting point.

Next, we will take this information and this time average across participants to get a PSD that represents that information for all subjects. This information will then be embedded ontop of the plots as a black line. Now while this is great to visualize, this information is misleading. This is because the aggregate we are plotting is derived from using all channel information in our data. However, as we specified in the previous section, we are only using information from 23 channels in the final analysis, specifically by aggregating across channels of interest within topographical regions of interest. Therefore, we did the same here by averaging across channels of interest within the 4 topographical regions we care for and the plotted that along with the PSD that was further averaged across subjects.


```{r Plotting the power spectra across channels}
# Part 1: Averaging across channels
avg_acros_chans <- data4L %>%
  group_by(ID, frequency) %>%
  summarise(avg_chan_pow = mean(power),
            .groups = "drop")

# Part 2: Averaging across channels and subjects
avg_acros_chans_sub <- avg_acros_chans %>%
   group_by(frequency) %>%
   summarise(avg_chan_sub_pow = mean(avg_chan_pow),
            .groups = "drop")

# Part 3: Averaging across channels within topography
avg_acros_chans_by_top <- data4L %>%
  group_by(ID, frequency, topography) %>%
  summarise(avg_chan_pow = mean(power),
            .groups = "drop") %>%
  drop_na()

# Part 4: Averaging across channels and subjects within topography
avg_acros_chans_sub_by_top <- avg_acros_chans_by_top %>%
   group_by(frequency, topography) %>%
   summarise(avg_chan_sub_pow = mean(avg_chan_pow),
            .groups = "drop")

# Visualizing the power spectra (New) averaged across channels
avg_acros_chans %>%
  ggplot(aes(x = frequency, y = avg_chan_pow, group = ID)) +
  geom_line(color = "darkgrey", alpha = 0.7, size = 0.5) +
  labs(title = "Channel-averaged PSD per participant (grey) and\ngrand average across participants (black)",
       x = "Frequency (Hz)", y = "Power (μV²)") +
  geom_line(data = avg_acros_chans_sub, 
            aes(x = frequency, y = avg_chan_sub_pow), 
            color = "black", size = 1.2, inherit.aes = FALSE) +
  theme_clean()

# Visualizing the power spectra (New) averaged across channels within topography
avg_acros_chans_by_top %>%
  ggplot(aes(x = frequency, y = avg_chan_pow, group = ID)) +
  geom_line(color = "darkgrey", alpha = 0.7, size = 0.5) +
  facet_wrap(~topography, scales = "free") +
  geom_line(data = avg_acros_chans_sub_by_top, 
            aes(x = frequency, y = avg_chan_sub_pow), 
            color = "black", size = 1.2, inherit.aes = FALSE) +
   labs(title = "Channel-averaged PSD per participant (grey) and\ngrand average across participants (black) by topography",
       x = "Frequency (Hz)", y = "Power (μV²)") +
  theme_clean()



```


# Calculating Absolute Power (Data Reduction)

We are starting back with our long dataset and this time we will be using it to calculate **absolute power**. We used Grok's help to basically accomplish this. Okay so we established that each subject has 8,618 observations (62 * 139 = 8,618). We do not care for information from 62 channels, we are only interested in information from 23 of them, so we will start by filtering out channels we do not care for. The easiest way to do this is just to drop observations where topography has NA's, because these are channels that were not labeled for any of the 4 topography groups. 

What we have now per person is 3,197 observations (23 * 139 = 3,197). What we need to do now is simplify our data, or reduce the number of observations by averaging across channels within our frequency bands of interest. This will result in a PSD for 4 topographies for each person (4 * 139 = 556 observations).

Absolute power is defined as the **sum of power values** within frequency bands. So the next step will be to sum power values within topographical regions. We have three frequency bands of interest (slow, alpha, beta), so the 556 observations will be isolated into 4 mutually exclusive groups (topography), from those groups which represent PSD, we will create sub groups of PSD within frequency bands (4 * 3 = 12 topography x frequency band groups), and from these groups we will take the chunked PSD values and sum them together for each person! Additionally, we will multiply this sum by the frequency resolution, which is 0.25. This apparently makes the values actual absolute power, but regardless when we analyze the data we will still get the same results had we not done this. 

Therefore, each subject will now have 3 * 4 = 12 power values. With this information, we can also group by either frequency bands or frequency band * topography level combinations and calculate z-scores. In the plots below we can see the consequence of calculating z-scores failing to account for topography, the data becomes very skewed. However when including topography it looks a lot more reasonable, but still skewed. Since we are going to use models that require a normally distributed outcome, we must have to do some transformation to our data.

So in plots below, we also take the log of the outcome and then conver these values into z-scores. From the plots we see that in the log scale values are no longer that extreme, aka not that far away from the mean, which means that our analysis will probably work on our outcome (absolute power) in this scale. 



```{r Calculating absolute power}
# Filter out channels we do not care for
chan_filt <- data4L %>%
  drop_na(topography)

# Average across channels within topography AND hemisphere regions
avg_across_chan_by_top <- chan_filt %>%
  group_by(ID, frequency, topography_hemisphere, Sex, Age, theta_good) %>%
  summarise(FB = FB[1],
            mean_pow = mean(power),
            .groups = "drop") %>%
  drop_na(FB)

# Calculating absolute power
final_dat <- avg_across_chan_by_top %>%
  group_by(ID, topography_hemisphere, FB, Sex, Age, theta_good) %>%
  summarise(absolute_power = sum(mean_pow) * 0.25,
            .groups = "drop")

# Creating a topography and hemisphere variable separately
final_dat <- final_dat %>%
  mutate(
    topography = case_when(
      grepl("frontal", topography_hemisphere) ~ "frontal",
      grepl("central", topography_hemisphere) ~ "central",
      grepl("temporal", topography_hemisphere) ~ "temporal",
      grepl("parietal", topography_hemisphere) ~ "parietal",
      grepl("occipital", topography_hemisphere) ~ "occipital"),
    hemisphere = case_when(
      grepl("left", topography_hemisphere) ~ "left",
      grepl("right", topography_hemisphere) ~ "right"),
    topography = factor(topography),
    hemisphere = factor(hemisphere)
  )

# Data points per person
cat("\nData points per person, should sum to 32\n")
xtabs(~topography_hemisphere + FB, filter(final_dat, ID == "10027"))


# Plotting z-scores of Absolute Power within the Frequency Bands
norm_plot1 <- final_dat %>%
  group_by(FB) %>%
  mutate(power_z = scale(absolute_power)) %>%
  ggplot(aes(x = FB, y = power_z)) +
  geom_boxplot() +
  labs(title = "Absolute Power Z-Scores within Frequency\nBands Ignoring Topography and Hemisphere",
       x = "Frequency Bands", y = "Absolute Power (Z)") +
  theme_clean()


# Plotting z-scores of Absolute Power within the Frequency Bands by Topography_Hemisphere
norm_plot2 <- final_dat %>%
  group_by(FB, topography_hemisphere) %>%
  mutate(power_z = scale(absolute_power)) %>%
  ggplot(aes(x = FB, y = power_z)) +
  facet_wrap(~hemisphere + topography , scales = "free", nrow = 2) +
  geom_boxplot() +
  labs(title = "Absolute Power Z-Scores within Frequency\nBands Accounting for Topography and Hemisphere", 
       x = "Frequency Bands", y = "Absolute Power (Z)") +
  theme_clean()

# Create a plot from both normalized plots
norm_plot1
norm_plot2


# Plotting z-scores of Absolute Power within the Frequency Bands
norm_plot1_log <- final_dat %>%
  group_by(FB) %>%
  mutate(power_z_log = scale(log(absolute_power))) %>%
  ggplot(aes(x = FB, y = power_z_log)) +
  geom_boxplot() +
  labs(title = "Logged Absolute Power Z-Scores within\nFrequencyBands Ignoring Topography and Hemisphere",
       x = "Frequency Bands", y = "Logged Absolute Power (Z)") +
  theme_clean()


# Plotting z-scores of Absolute Power within the Frequency Bands by Topography
norm_plot2_log <- final_dat %>%
  group_by(FB, topography_hemisphere) %>%
  mutate(power_z_log = scale(log(absolute_power))) %>%
  ggplot(aes(x = FB, y = power_z_log)) +
  facet_wrap(~hemisphere + topography , scales = "free", nrow = 2) +
  geom_boxplot() +
  labs(title = "Logged Absolute Power Z-Scores within\nFrequency Bands Accounting for Topography and Hemisphere", 
       x = "Frequency Bands", y = "Logged Absolute Power (Z)") +
  theme_clean()


norm_plot1_log  
norm_plot2_log
```


# Removing Outliers

Since we are not planning on using any type of `glm()` model, I think it would make sense to remove outliers based on how we are analyzing the data. For example, our final data is going to look something like this:

absolute_power ~ Age + Topography + Spell

So what we want to know is if there are any outliers in our outcome when it is grouped like this (aka separated by factors). So what makes the most sense I think is to group the outcome for each frequency band x topography, calculate the log scores within those groups, z-transform them, and remove any values 2.58 SD from the mean. I think this should work because log transformation forces the outcome to look normal, in that case the identification of outliers (far from the mean) would make sense. Afterwards we will generate the boxplots from above to confirm we will this correctly


```{r Removing outliers}
# Grouping the data
final_dat <- final_dat %>%
  group_by(FB, topography_hemisphere) %>%
  mutate(log_absolute_power = log(absolute_power),
         z_log_absolute_power = as.vector(scale(log_absolute_power)),
         outlier = ifelse(z_log_absolute_power <= - 2.58 | z_log_absolute_power >=  2.58, 1, 0))

# Paste the number of IDs before removing outlier scores
paste0("Before removing outliers, there were ", length(unique(final_dat$ID)), " unique IDs")

# Remove the outlier
final_dat <- filter(final_dat, outlier != 1)

# Paste the number of IDs before removing outlier scores
paste0("After removing outliers, there were ", length(unique(final_dat$ID)), " unique IDs")

# Plotting z-scores of Absolute Power within the Frequency Bands
norm_plot1_log_no_out <- final_dat %>%
  group_by(FB) %>%
  mutate(power_z_log = scale(log(absolute_power))) %>%
  ggplot(aes(x = FB, y = power_z_log)) +
  geom_boxplot() +
  labs(title = "Logged Absolute Power Z-Scores within\nFrequencyBands Ignoring Topography and Hemisphere",
       x = "Frequency Bands", y = "Logged Absolute Power (Z)") +
  theme_clean()


# Plotting z-scores of Absolute Power within the Frequency Bands by Topography
norm_plot2_log_no_out <- final_dat %>%
  group_by(FB, topography_hemisphere) %>%
  mutate(power_z_log = scale(log(absolute_power))) %>%
  ggplot(aes(x = FB, y = power_z_log)) +
  facet_wrap(~hemisphere + topography , scales = "free", nrow = 2) +
  geom_boxplot() +
  labs(title = "Logged Absolute Power Z-Scores within\nFrequency Bands Accounting for Topography and Hemisphere", 
       x = "Frequency Bands", y = "Logged Absolute Power (Z)") +
  theme_clean()


norm_plot1_log_no_out
norm_plot2_log_no_out

```



# Missing Data Check (Removing Missing Data If Any- Hint There Should Not be Any!)

There are two main of analyses that will be doing in the upcoming sections. The first is to just use between-person level data to obtain fitted values for the relationships between spelling ability (theta values) and age. We are running this model specifically to identify good and poor spellers after accounting for age. This is important because if either of these data are missing, the model will drop them entirely. 

Similarly, the main analysis will be using a mixed-effects model. Now, mixed-effects models are known to handle missing data from a participant, but only if that data corresponds to the within-group (within-person) factor. Our within group factor will be topography. Therefore, if a participant is missing some levels of topography, their data will still be included in the analysis but weighted differently from the others, no problem. However, for between-group (between-person) factors, missing data is important to identify because it will force that participant's data (all of their data) to be dropped. Therefore, it is good to identify whether subjects have missing between-group factor data and if so drop them now to prevent errors when introducing fitted values back into the dataset.

Below we will only be dropping subjects that have missing spelling or age data. Technically this section is more for quality control- this should have already been done in sections long before this one.
 
```{r Missing data check}
# Current number of IDs
temp <- final_dat
cat("Our current final dataset has",length(unique(final_dat$ID)),"IDs\n\n" )

# Identifying how much data is missing
cat("After using slice_head() on the data, here is how much data is missing per variable\n")
final_dat_head <- final_dat %>%
  group_by(ID) %>%
  slice_head(n=1)
sapply(final_dat_head, function(x) sum(is.na(x)))

# Drop observations that have missing age or spelling data
final_dat <- final_dat %>%
  drop_na(Age, theta_good)

# Print out the number of IDs we will have after removing data
cat("After dropping IDs with missing Age or Spelling data, we have", length(unique(final_dat$ID)),"unique IDs")
```



# Final Demographics

From this point on, no observations will be removed from our data. Thus, here we will report our final sample size a long with descriptives of our data. Lastly, we will also show missing values for each variable. There should be no missing values in our predictors, covariates, and outcomes of interest. 

```{r Generating the final descriptives information}
# The number of unique IDs in our data
final_ID_num <- length(unique(final_dat$ID))
obs_num <- 12
total_obs <- final_ID_num * obs_num
cat("\nThere are",final_ID_num,"unique IDs in our sample and each one should have 12 observations, for a total of",total_obs,"observations\n")
cat("\nCounting the number of rows in the data, there are",nrow(final_dat),"rows\n")

# Create the age centered variable
final_dat <- mutate(final_dat, Age_min_c = Age - min(Age))

# Higher level variables (person level) 
final_dat_sliced <- final_dat %>%
  group_by(ID) %>%
  slice_head(n=1) %>%
  ungroup()

# Create a graph for the final demographics
# Save the min and max age as objects
mean_age <- round(mean(final_dat_sliced$Age),1)
sd_age <- round(sd(final_dat_sliced$Age), 1)
min_age <- min(final_dat_sliced$Age)
max_age <- max(final_dat_sliced$Age)

# Generate a histogram of our Age
dem_plot1 <- final_dat_sliced %>%
  ggplot(aes(x = Age)) +
  geom_histogram(color = "black", fill = "white") +
  labs(title = paste0("Sample age distribution\n(n=",final_ID_num,
                      "; mean = ", mean_age, "; sd = ", sd_age,";\n",
                      "min = ",min_age,"; max = ",max_age,")")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  theme_clean()

# Create a histogram of sex?
dem_plot2 <- final_dat_sliced %>%
  group_by(Sex) %>%
  summarise(count = length(Sex),
            percent = length(Sex)/nrow(final_dat_sliced)*100) %>%
  mutate(label = paste0(count, " (", round(percent,1),"%)")) %>%
  ggplot(aes(x = Sex, y = count, fill = Sex)) +
  geom_col(width = .5, color = "black") +
  geom_text(aes(label = label), vjust = -.6, size = 4.2) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.25))) +
  theme_clean() +
  labs(title = "Number and percentage of\nparticipants by sex") +
  theme(legend.position = "none")

# Create one comprehensive descriptive plot
dem_plot1 + dem_plot2

# When checking for missing data, 
cat("\nHere are values representing missing data for each variable in the final dataset\n")
sapply(final_dat, function(x) sum(is.na(x)))
```



# Create Spelling Performance Group Using Fitted Values from Quadratic to Plateau Model

[We deleted the section before showing a scatterplot between Age and Spelling Ability because who cares, we show the same thing below but with more information]

To address the concerns above, we can fit a quadratic to plateau model to model Spelling Ability as a function of Age. This will produce fitted values (aka the expected spelling score for each level of Age) that we can use to our advantage. We can take the difference of the observed value from the fitted value and then quantify this difference into quantiles. Those in the bottom 20% percentile will become the poor performance spellers while the rest will be the good performance spellers.

The advantage of this approach is that it basically categorizes poor spellers as individuals who performed poorly in comparison to individuals of a similar age. Also, the spelling ability scores were derived from a much larger sample, actually from the sample of all IDs that had spelling data. This is good because that sample was 661 subjects. Now while I am no IRT expert, I think this large of a sample size plus the fact that there were poor performers in the spelling task based on raw scores, justifies that standardized values, which represent distance from the mean performers, is still capturing true poor spellers. 


```{r Creating a spelling group using quadratic to plateau models}
# Load in the package
library(nlraa)
library(msm)

# run a quadratic to plateau regression
spell.nls <- nls(theta_good ~ SSquadp3xs(Age, a, b, jp), 
                 data = final_dat_sliced)

# Using predict2_nls() function to get SE for fitted values across level of Age
spell.nls.pred <- predict2_nls(spell.nls, newdata = NULL, interval = "confidence", level = 0.95)
final_dat_sliced$theta_good.fitted <- spell.nls.pred$Estimate

# Create the grouping variable based on subjects with less than - 1.5 SE from the model
final_dat_sliced <- final_dat_sliced %>%
  mutate(SpellDiscrepency = theta_good - theta_good.fitted,
         SpellThresh = quantile(SpellDiscrepency, 0.20), # 1-20% Percentile vs 20-100% percentile
         SpellThresh2 = quantile(SpellDiscrepency, 0.80),
         SpellGroup = ifelse(SpellDiscrepency <= SpellThresh, "Poor", "Good"),
         SpellGroup = factor(SpellGroup, levels = c("Poor", "Good")),
         SpellGroupExt = ifelse(SpellDiscrepency <= SpellThresh, "Poor",
                                ifelse(SpellDiscrepency >= SpellThresh2, "Very Good", "Average")),
         SpellGroupExt = factor(SpellGroupExt, levels = c("Poor", "Average","Very Good")))

# Add this information into the long final dataset
final_dat <- final_dat %>%
  left_join(select(final_dat_sliced, ID, SpellGroup, SpellGroupExt),
            by = "ID") %>%
  data.frame()

# View the number of participants in each group (original)
group.num <- xtabs(~SpellGroup, final_dat_sliced)
group.prop <- sapply(round(prop.table(xtabs(~SpellGroup, final_dat_sliced)) * 100,1), function(x) paste0(x,"%"))

# View the number of participants in each group (extremes)
group.num2 <- xtabs(~SpellGroupExt, final_dat_sliced)
group.prop2 <- sapply(round(prop.table(xtabs(~SpellGroupExt, final_dat_sliced)) * 100,1), function(x) paste0(x,"%"))


# Create a caption for the model
spl.coefs <- round(coef(spell.nls),2)
spll.caption <- paste0("Intercept = ", spl.coefs[1], "; slope = ", spl.coefs[2], "; join point = ", spl.coefs[3],
                       "; Good Spellers = ", group.num["Good"]," (",group.prop["Good"], ")",
                       "; Poor Spellers = ", group.num["Poor"]," (",group.prop["Poor"],")")

# Add the fitted values
final_dat_sliced %>%  
  ggplot(aes(x = Age, y = theta_good, color = SpellGroup)) +
  geom_jitter(width = .5, alpha = 0.6) +  # Added alpha for better visibility if overlapping
  geom_line(aes(y = theta_good.fitted), linewidth = 1.5, linetype = "dashed", color = "black") +
  labs(
    title = "Generating Spelling Groups from EAP Scores After Accounting for Age",
    y = "Spelling Ability (EAP Scores)",
    caption = spll.caption,
    color = "Spelling Group") +
  theme_clean() 

# Create another version for very good vs poor spellers
final_dat_sliced %>%  
  ggplot(aes(x = Age, y = theta_good, color = SpellGroupExt)) +
  geom_jitter(width = .5, alpha = 0.6) +  # Added alpha for better visibility if overlapping
  geom_line(aes(y = theta_good.fitted), linewidth = 1.5, linetype = "dashed", color = "black") +
  labs(
    title = "Generating Spelling Groups from EAP Scores After Accounting for Age\n(Extreme Groupings)",
    y = "Spelling Ability (EAP Scores)",
    caption = spll.caption,
    color = "Spelling Group") +
  theme_clean() 

```


# Age and Spelling Group Tabulation Table (Contingency Table)

Now that we have created a spelling group factor, let's see what the cell sample sizes are. We will bin people into age groups and then show within those bins, how many good and poor spellers there are. 



```{r spelling group by age contingency table}
# Create an Age group variable to see our sample size across age when turned into a factor
final_dat_sliced <- final_dat_sliced %>%
  mutate(Age_group = case_when(
    Age >= 15 & Age < 18 ~ "15-17",
    Age >= 18 & Age < 22 ~ "18-21",
    Age >= 22 & Age < 27 ~ "22-26",
    Age >= 27 & Age <= 38 ~ "27-38",
    TRUE ~ "ERROR"
  ))


# Use xtabs() to get a frequency table
xtabs_tb <- addmargins(xtabs(~Age_group + SpellGroup, final_dat_sliced), margin = 1)
xtabs_df <- as.data.frame.array(xtabs_tb)
xtabs_df$`Poor %` <- round(xtabs_df$Poor /  (xtabs_df$Good + xtabs_df$Poor),3) *100

# Some data cleaning
xtabs_df$age <- row.names(xtabs_df)
xtabs_df <- select(xtabs_df, age, Poor, Good, everything())
row.names(xtabs_df) <- NULL  

# Rename the variables
names(xtabs_df) <- c("Age Group", "Good", "Poor", "Poor %")

# Print out a nice table
xtabs_df %>%
  gt() %>%
  tab_header(
    title = "Number of Good and Poor Spellers by Age Group"
  ) %>%
  tab_options(
    table.font.size = 13,
    heading.title.font.size = 18
  )

# Redundant table using the kbl() function instead
xtabs_df %>%
  kbl(caption = "<span style='color:black; font-size:16px; display:block; text-align:center;'>Number of Good and Poor Spellers by Age Groups</span>",
      escape = FALSE) %>%
  kable_paper(full_width = F)

```



# Visualizing Data Using Topographic Maps

We will no longer be removing subjects past this point, thus we can go back to our original long data, extract the IDs that made the cut, then use that information to plot topographic maps that can show us:
a) Effects of Age across the frequency bands
b) Effects of Spelling Performance across the frequency bands
c) Age x Spelling Performance interaction across the frequency bands 

```{r Visualizing topographic maps}
# Load in package
library(eegUtils)

# Extract the long information from the IDs that will be used for the analysis
data5L <- data4L %>%
  filter(ID %in% unique(final_dat$ID))

# Introduce Spelling Label into the data
data5L <- data5L %>%
  left_join(select(final_dat_sliced, ID, SpellGroup), by = "ID")

# Create a function to plot main topographic frequency band plots
topo_plot_fun1 <- function(data, FB) {
  # Filter the data by freq band then average across electrodes
  FB_data <- data5L %>%
    filter(FB == {{FB}}) %>%
    group_by(channel) %>%
    summarise(power = mean(power, na.rm = TRUE), .groups = "drop") %>%
    rename(electrode = channel)
  
  # Introduce electrode location
  plot_FB_data <- electrode_locations(FB_data)
  
  # Plot the data
  topoplot(plot_FB_data, quantity = "power") + labs(title = paste0(FB))
 }

# Generate the plots for each topographical region
main_delta <- topo_plot_fun1(data5L, "delta")
main_theta <- topo_plot_fun1(data5L, "theta")
main_alpha1 <- topo_plot_fun1(data5L, "alpha1")
main_alpha2 <- topo_plot_fun1(data5L, "alpha2")
main_beta1  <- topo_plot_fun1(data5L, "beta1")
main_beta2  <- topo_plot_fun1(data5L, "beta2")

# Print all main plots side by side 
main_delta + main_theta + main_alpha1 # Slow waves
main_alpha2 + main_beta1 + main_beta2 # Fast waves


# Create a function to plot frequency band topographic plots by age group
topo_plot_fun2 <- function(data, FB) {
  # Filter the data by freq band then average across electrodes
  FB_data_young <- data5L %>%
    filter(FB == {{FB}},
           Age >= 15 & Age < 20) %>%
    group_by(channel) %>%
    summarise(power = mean(power, na.rm = TRUE), .groups = "drop") %>%
    rename(electrode = channel)
  
  FB_data_middle <- data5L %>%
    filter(FB == {{FB}},
           Age >= 20 & Age < 25) %>%
    group_by(channel) %>%
    summarise(power = mean(power, na.rm = TRUE), .groups = "drop") %>%
    rename(electrode = channel)
  
  FB_data_old <- data5L %>%
    filter(FB == {{FB}},
           Age >= 25 & Age < 38) %>%
    group_by(channel) %>%
    summarise(power = mean(power, na.rm = TRUE), .groups = "drop") %>%
    rename(electrode = channel)
  
  # Introduce electrode location
  FB_data_young <- electrode_locations(FB_data_young)
  FB_data_middle <- electrode_locations(FB_data_middle)
  FB_data_old <- electrode_locations(FB_data_old)
  
  # Plot the data
  plot1 <- topoplot(FB_data_young, quantity = "power") + labs(title = paste0(FB, ": 15-19"))
  plot2 <- topoplot(FB_data_middle, quantity = "power") + labs(title = paste0(FB, ": 20-24"))
  plot3 <- topoplot(FB_data_old, quantity = "power") + labs(title = paste0(FB, ": 25-38"))
  plot1 + plot2 + plot3
}

# Create topographic plots grouped by sample ages
topo_plot_fun2(data5L, "delta")
topo_plot_fun2(data5L, "theta")
topo_plot_fun2(data5L, "alpha1")
topo_plot_fun2(data5L, "alpha2")
topo_plot_fun2(data5L, "beta1")
topo_plot_fun2(data5L, "beta2")


# Create a function to plot frequency band topographic plots by Spelling Performance
topo_plot_fun3 <- function(data, FB) {
  # Filter the data by freq band then average across electrodes
  PoorSpellers <- data5L %>%
    filter(FB == {{FB}},
           SpellGroup == "Poor") %>%
    group_by(channel) %>%
    summarise(power = mean(power, na.rm = TRUE), .groups = "drop") %>%
    rename(electrode = channel)
  
  GoodSpellers <- data5L %>%
    filter(FB == {{FB}},
           SpellGroup == "Good") %>%
    group_by(channel) %>%
    summarise(power = mean(power, na.rm = TRUE), .groups = "drop") %>%
    rename(electrode = channel)
  
  # Introduce electrode location
  plot_PoorSpellers_data <- electrode_locations(PoorSpellers)
  plot_GoodSpellers_data <- electrode_locations(GoodSpellers)
  
  # Plot the data
  poor_plot <- topoplot(plot_PoorSpellers_data, quantity = "power") + labs(title = paste0(FB,": Poor"))
  good_plot <- topoplot(plot_GoodSpellers_data, quantity = "power") + labs(title = paste0(FB,": Good"))
  poor_plot + good_plot
}

# Plot the frequency bands between good and poor spellers
topo_plot_fun3(data5L, "delta")
topo_plot_fun3(data5L, "theta")
topo_plot_fun3(data5L, "alpha1")
topo_plot_fun3(data5L, "alpha2")
topo_plot_fun3(data5L, "beta1")
topo_plot_fun3(data5L, "beta2")
```



# Visualizing the Distribution of Absolute Power Within Topographies and in Relation to Age

The histograms of absolute power grouped for each topography are showing that the data is pretty skewed to the right. This is problematic because in order for us analyze our data we need to see a normal distribution to produce stable and reliable standard errors. Now this is not always the case, what really matters is the distribution of your errors, however this is a good shortcut to predict whether our errors will look normal or not.

In Whitford's paper, they also report that their distribution of absolute power for their sample was skewed also to the right. Since their analysis involved using regression, they needed their outcome distribution (absolute power) to look normal, thus they square root their data. 

Below we are doing something similar, we are plotting the distribution of absolute power in its original scale, after the values were transformed with the square root and a transformation of values using logarithm. We see below that our data looks best (normally distributed) when we use the log transformations. 

Additionally we are also showing the scatterplot between that transformation of the outcome (absolute power) and age. These descriptives give us a hint on whether or not we see that with age absolute power decreases, which theoretically is something we would expect.  

We can kinda see this with the slow wave, where with age the absolute power of the slow wave (0.5 - 7.5 Hz) decreases, however for alpha and beta we do not see this relationship.

```{r visualizing the relationship with age}
# Create a custom function for scatterplots
plot_scatr <- function(data, x, y) {
ggplot(data,aes(x = {{x}}, y = {{y}})) +
    geom_point(shape = 21,          # Creates circle with filling and border
             fill = "darkgrey",   # Inner fill color
             color = "black",     # Border color
             size = 1,            # Point size (adjust for thicker border)
             stroke = 1) +      # Border thickness (optional, default is 0.5)
  theme_clean() 
}

# Create a function to plot a frequency band by topography
plot_fb_top <- function(data, x, y, FB) {
data_filt <- data %>%
    filter(FB == !!FB) # To make it filter correctly

plot1 <- data_filt %>%
  plot_scatr({{x}}, {{y}}) +
  facet_wrap(~topography, scales = "free") +
  labs(title = paste("Frequency band:", FB))

plot2 <- data_filt %>%
  plot_hist({{y}}) +
  facet_wrap(~topography) +
  labs(title = paste("Distribution of:", FB))

# Print out both plots
plot1 + plot2
}

# Plot mean power regressed on age across topography for each FB
plot_fb_top(final_dat, Age, absolute_power, "delta")
plot_fb_top(final_dat, Age, absolute_power, "theta")
plot_fb_top(final_dat, Age, absolute_power, "alpha1")
plot_fb_top(final_dat, Age, absolute_power, "alpha2")
plot_fb_top(final_dat, Age, absolute_power, "beta1")
plot_fb_top(final_dat, Age, absolute_power, "beta2")

# Plot it but sqrt transformed
plot_fb_top(final_dat, Age, sqrt(absolute_power), "delta")
plot_fb_top(final_dat, Age, sqrt(absolute_power), "theta")
plot_fb_top(final_dat, Age, sqrt(absolute_power), "alpha1")
plot_fb_top(final_dat, Age, sqrt(absolute_power), "alpha2")
plot_fb_top(final_dat, Age, sqrt(absolute_power), "beta1")
plot_fb_top(final_dat, Age, sqrt(absolute_power), "beta2")

# Plot it but log transformed
plot_fb_top(final_dat, Age, log(absolute_power), "delta")
plot_fb_top(final_dat, Age, log(absolute_power), "theta")
plot_fb_top(final_dat, Age, log(absolute_power), "alpha1")
plot_fb_top(final_dat, Age, log(absolute_power), "alpha2")
plot_fb_top(final_dat, Age, log(absolute_power), "beta1")
plot_fb_top(final_dat, Age, log(absolute_power), "beta2")


```
# Creating Figure 1 for our paper

**Figure 1** is literally the information from above except made a little more prettier. We want to show how raw EEG power changes as a function of age. I think it makes the most sense to show the log transformed 

```{r Creating Figure 1 which is the relationship between EEG power and Age}
# Change the factor level of topography for plotting purposes
final_dat <- final_dat %>%
  mutate(topography = factor(topography, levels = c("frontal", "central","temporal", "parietal", "occipital")))

# Create some models to produce fitted values using predict
mod1 <- lm(log(absolute_power) ~ Age + I(Age^2) + topography, filter(final_dat, FB == "delta"))
mod2 <- lm(log(absolute_power) ~ Age + I(Age^2) + topography, filter(final_dat, FB == "theta"))
mod3 <- lm(log(absolute_power) ~ Age + I(Age^2) + topography, filter(final_dat, FB == "alpha1")) 
mod4 <- lm(log(absolute_power) ~ Age + I(Age^2) + topography, filter(final_dat, FB == "alpha2")) 
mod5 <- lm(log(absolute_power) ~ Age + I(Age^2) + topography, filter(final_dat, FB == "beta1"))
mod6 <- lm(log(absolute_power) ~ Age + I(Age^2) + topography, filter(final_dat, FB == "beta2"))

# Create fitted values
newdata <- expand_grid(
  Age = seq(14, 35, length.out = 100),
  topography = levels(final_dat$topography)) %>%
  mutate(topography = factor(topography, levels = levels(final_dat$topography)))


# Create a data frame of fitted values
fittedvals <- data.frame(
  Age = newdata$Age,
  topography = factor(newdata$topography, levels = levels(final_dat$topography)),
  fit1 = predict(mod1, newdata = newdata),
  fit2 = predict(mod2, newdata = newdata),
  fit3 = predict(mod3, newdata = newdata),
  fit4 = predict(mod4, newdata = newdata),
  fit5 = predict(mod5, newdata = newdata),
  fit6 = predict(mod6, newdata = newdata))


# Save the plots
p1 <- final_dat %>%
  filter(FB == "delta") %>%
  ggplot(aes(x = Age, y = log(absolute_power))) +
  geom_point(size = 1, alpha = .5) +
  facet_wrap(~topography, scales = "free") +
  geom_line(data = fittedvals, aes(x = Age, y = fit1), linewidth = 1, color = "red") +
  labs(title = "Delta (0.5 - 3.5 Hz)", y = "Log of Absolute Power\n", x = NULL) +
  theme_clean()

# Save the plots
p2 <- final_dat %>%
  filter(FB == "theta") %>%
  ggplot(aes(x = Age, y = log(absolute_power))) +
  geom_point(size = 1, alpha = .5) +
  facet_wrap(~topography, scales = "free") +
  geom_line(data = fittedvals, aes(x = Age, y = fit2), linewidth = 1, color = "red") +
  labs(title = "Theta (4 - 7.5 Hz)", y = "Log of Absolute Power\n", x = NULL) +
  theme_clean()

# Save the plots
p3 <- final_dat %>%
  filter(FB == "alpha1") %>%
  ggplot(aes(x = Age, y = log(absolute_power))) +
  geom_point(size = 1, alpha = .5) +
  facet_wrap(~topography, scales = "free") +
  geom_line(data = fittedvals, aes(x = Age, y = fit3), linewidth = 1, color = "red") +
  labs(title = "Alpha1 (8 - 10 Hz)", x = "Age (Years)", y = NULL) +
  theme_clean()


# Save the plots
p4 <- final_dat %>%
  filter(FB == "alpha2") %>%
  ggplot(aes(x = Age, y = log(absolute_power))) +
  geom_point(size = 1, alpha = .5) +
  facet_wrap(~topography, scales = "free") +
  geom_line(data = fittedvals, aes(x = Age, y = fit4), linewidth = 1, color = "red") +
  labs(title = "Alpha2 (10 - 12.5 Hz)",  x = NULL, y = NULL) +
  theme_clean()

# Save the plots
p5 <- final_dat %>%
  filter(FB == "beta1") %>%
  ggplot(aes(x = Age, y = log(absolute_power))) +
  geom_point(size = 1, alpha = .5) +
  facet_wrap(~topography, scales = "free") +
  geom_line(data = fittedvals, aes(x = Age, y = fit5), linewidth = 1, color = "red") +
  labs(title = "Beta1 (12.5 - 18 Hz)",  x = NULL, y = NULL) +
  theme_clean()

# Save the plots
p6 <- final_dat %>%
  filter(FB == "beta2") %>%
  ggplot(aes(x = Age, y = log(absolute_power))) +
  geom_point(size = 1, alpha = .5) +
  facet_wrap(~topography, scales = "free") +
  geom_line(data = fittedvals, aes(x = Age, y = fit6), linewidth = 1, color = "red") +
  labs(title = "Beta2 (18 - 34.5 Hz)",  x = NULL, y = NULL) +
  theme_clean()


# Plot all the graphs together
p1 + p2 + p3 
p4 + p5 + p6
```



### Visualize the Relationship Between Spelling and EEG power

Visually it looks like spelling ability does not seem to be predictive at all of rsEEG absolute power. This is preparing us for the fact that we are very likely not going to see any differences between good and poor spellers- their EEG activity looks the same.

```{r visualize the relationship between spelling ability and EEG power}
# Create a custom function that plots absolute_power regressed onto Age by spelling group
plot_age_pwr_grp <- function(data, x, y, FB) {
  data_filt <- filter(data, FB == !!FB)
  data_filt %>%
    ggplot(aes(x = {{x}}, y = {{y}}, color = SpellGroup)) +
    geom_jitter(alpha = .5, size = .7) +
    facet_wrap(~topography, scales = "free") +
    labs(title = paste0("Frequency Band: ", FB)) +
    theme_clean()
}

# Creating a custom function that turns Age into a factor with 5 levels
# This is a really bad function but w/e
plot_age_Q_pwr_grp <- function(data, x, y, FB) {
  data %>%
    mutate(Age_Q = cut({{ x }}, 
                       breaks = quantile({{ x }}, probs = seq(0, 1, by = 0.2), na.rm = TRUE),
                       include.lowest = TRUE,
                       labels = c("Q1", "Q2", "Q3", "Q4", "Q5"))) %>%
    filter(FB == !!FB) %>%
    ggplot(aes(x = Age_Q, y = {{ y }}, fill = SpellGroup)) +
    facet_wrap(~topography, scales= "free") +
    geom_boxplot() +
    labs(title = paste0("Frequency Band: ", FB),
         x = "Age Quantiles") +
    theme_clean()
}

# Plot EEG power regressed by Age and Spelling Group for each FB
plot_age_pwr_grp(final_dat, Age, log(absolute_power), "delta") + geom_smooth(method = "lm", se = FALSE)
plot_age_pwr_grp(final_dat, Age, log(absolute_power), "theta") + geom_smooth(method = "lm", se = FALSE)
plot_age_pwr_grp(final_dat, Age, log(absolute_power), "alpha1") + geom_smooth(method = "lm", se = FALSE)
plot_age_pwr_grp(final_dat, Age, log(absolute_power), "alpha2") + geom_smooth(method = "lm", se = FALSE)
plot_age_pwr_grp(final_dat, Age, log(absolute_power), "beta1") + geom_smooth(method = "lm", se = FALSE)
plot_age_pwr_grp(final_dat, Age, log(absolute_power), "beta2") + geom_smooth(method = "lm", se = FALSE)

# Show the differences in EEG power by group
plot_age_Q_pwr_grp(final_dat, Age, log(absolute_power), "delta")
plot_age_Q_pwr_grp(final_dat, Age, log(absolute_power), "theta")
plot_age_Q_pwr_grp(final_dat, Age, log(absolute_power), "alpha1")
plot_age_Q_pwr_grp(final_dat, Age, log(absolute_power), "alpha2")
plot_age_Q_pwr_grp(final_dat, Age, log(absolute_power), "beta1")
plot_age_Q_pwr_grp(final_dat, Age, log(absolute_power), "beta2")
```


### Simulate three potential outcomes for our data

```{r Simulating different outcomes we could find}
# a = intercept
# b = slope
# c = quadratic term (curvy bit)
# jp = join point = break point = critical concentration

qp <- function(x, a, b, jp) {
    c <- -0.5 * b / jp
    if_else(condition = x < jp,
            true  = a + (b * x) + (c * x * x),
            false = a + (b * jp) + (c * jp * jp))
}

# Create some fitted values
x <- 12:30
y1 <- qp(x, 28+16, -2-1.2, 25)
y2 <- qp(x, 36+16, -2-1.2, 25)
y3 <- qp(x, 36+20, -2-1.57, 29)
y4 <- qp(x, 28+18.5+3, -2-1.25, 26.5)
y5 <- qp(x, 28+18.5+4, -2-1.2, 27)

# Create a data frame
p_out <- data.frame(x = x, 
                    Good = y1,
                    Poor = y2)

# Create another dataframe
p_out2 <- data.frame(x = x, 
                    Good = y1,
                    Poor = y3)


# Create another data frame
p_out3 <- data.frame(x = x, 
                     Good = y4,
                     Poor = y5)

# Convert to long format
p_out_long <- p_out %>%
  pivot_longer(cols = c(Good, Poor),
               names_to = "group",
               values_to = "y")

p_out_long2 <- p_out2 %>%
  pivot_longer(cols =  c(Good, Poor),
               names_to = "group",
               values_to = "y")

p_out_long3 <- p_out3 %>%
  pivot_longer(cols =  c(Good, Poor),
               names_to = "group",
               values_to = "y")

# Plot the lines
h1 <- p_out_long2 %>%
  mutate(group = factor(group, levels = c("Poor", "Good"))) %>%
  ggplot(aes(x = x, y = y, color = group)) +
  geom_vline(xintercept = 24, linetype="dashed", color = "black") +
  geom_vline(xintercept = 28, linetype="dashed", color = "black") +
  geom_hline(yintercept = 4, linetype="dashed", color = "black") +
  geom_line(linewidth = 1) +
  labs(title = "A: Maturational Lag", x= NULL, y = "Absolute Power", color = NULL) +
  coord_cartesian(
    xlim = c(11, 30),     
    ylim = c(0, 22)       
  ) +
  theme_clean()

h2 <- p_out_long %>%
  mutate(group = factor(group, levels = c("Poor", "Good"))) %>%
  ggplot(aes(x = x, y = y, color = group)) +
  geom_vline(xintercept = 24, linetype="dashed", color = "black") +
  geom_hline(yintercept = 12, linetype="dashed", color = "black") +
  geom_hline(yintercept = 4, linetype="dashed", color = "black") +
  geom_line(linewidth = 1) +
  labs(title = "B: Maturational Deficit", x= "Age (Years)", y = NULL, color = "Spelling Groups") +
  coord_cartesian(
    xlim = c(11, 30),     
    ylim = c(0, 22)       
  ) +
  theme_clean()


h3 <- p_out_long3 %>%
  mutate(group = factor(group, levels = c("Poor", "Good"))) %>%
  ggplot(aes(x = x, y = y, color = group)) +
  geom_vline(xintercept = 26, linetype="dashed", color = "black") +
  geom_hline(yintercept = 7.5, linetype="dashed", color = "black") +
  geom_hline(yintercept = 6.5, linetype="dashed", color = "black") +
  geom_line(linewidth = 1) +
  labs(title = "C: Comparable Development", x= NULL, y = NULL, color = NULL) +
  coord_cartesian(
    xlim = c(11, 30),     
    ylim = c(0, 22)       
  ) +
  theme_clean()

# Print all side by side
h1 + h2 + h3
```


### Fitting Best Quadratic to Plateau Model to Delta Frequency Band

Below we will be fitting the final model to answer our research question:

- Are there differences between good and poor spelling adults and does it show evidence for a maturational lag or a neurobiological deficit.
- Is the relationship above moderated by topographic regions? hemisphere? or both?

We will start by fitting the spelling factor first, then adding additional factors plus their interactions. Then we will do model comparison to identify which model best explains the data. 


```{r Fitting QTP on Delta data}
# Convert sex into a factor
final_dat <- final_dat %>%
  mutate(Sex = factor(Sex))

# Extracting datasets of interest
delta_dat <- filter(final_dat, FB == "delta")


# Create an outcome with the log transformed version of power for data within each frequency band
delta_dat <- mutate(delta_dat, log_absolute_delta_power = log(absolute_power))


##
###### Intercept = a
##

# Main effects only
delta_a_m1 <-  nlme(log_absolute_delta_power ~ SSquadp3xs(Age_min_c, a, b, xs),
                        data = delta_dat,                               
                        fixed = list(a ~ Sex + hemisphere + topography + SpellGroup, 
                                     b ~ SpellGroup, xs ~ SpellGroup),
                        random = a ~ 1 | ID,
                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + 1 - 1),
                                   b = -4, 0, 
                                   xs = 28, 0),
                        control = nlmeControl(msMaxIter = 2000))


# Spelling with Interactions for topography
delta_a_m2 <-  nlme(log_absolute_delta_power ~ SSquadp3xs(Age_min_c, a, b, xs),
                        data = delta_dat,                               
                        fixed = list(a ~ Sex + hemisphere + topography + SpellGroup +
                                       SpellGroup:hemisphere + SpellGroup:topography, 
                                     b ~ SpellGroup, xs ~ SpellGroup),
                        random = a ~ 1 | ID,
                        start = c(3, rep(0, 12),
                                   b = -4, 0, 
                                   xs = 28, 0),
                        control = nlmeControl(msMaxIter = 2000))

# Spelling with Interactions for topography and hemisphere
delta_a_m3 <-  nlme(log_absolute_delta_power ~ SSquadp3xs(Age_min_c, a, b, xs),
                        data = delta_dat,                               
                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
                                     b ~ SpellGroup, xs ~ SpellGroup),
                        random = a ~ 1 | ID,
                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
                                   b = -4, 0, 
                                   xs = 28, 0),
                        control = nlmeControl(msMaxIter = 2000))

# Model comparison
anova(delta_a_m1, delta_a_m2, delta_a_m3)


##
###### Slope = b
##


# Spelling Interaction with topography
delta_b_m1 <-  nlme(log_absolute_delta_power ~ SSquadp3xs(Age_min_c, a, b, xs),
                        data = delta_dat,                               
                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
                                     b ~ hemisphere + topography + SpellGroup + 
                                       SpellGroup:hemisphere + SpellGroup:topography, 
                                     xs ~ SpellGroup),
                        random = a ~ 1 | ID,
                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
                                   b = -4, rep(0, 11), 
                                   xs = 28, 0),
                        control = nlmeControl(msMaxIter = 2000))

# Spelling Interaction with topography and hemisphere
delta_b_m2 <-  nlme(log_absolute_delta_power ~ SSquadp3xs(Age_min_c, a, b, xs),
                        data = delta_dat,                               
                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
                                     b ~ topography * hemisphere * SpellGroup, xs ~ SpellGroup),
                        random = a ~ 1 | ID,
                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
                                   b = -4, rep(0, 4 + 1 + 1 + (1*4) + (1*4) + (1*1)+ (1*1*4) + 1 - 1), 
                                   xs = 28, 0),
                        control = nlmeControl(msMaxIter = 2000))

# Model comparison
anova(delta_a_m3, delta_b_m1, delta_b_m2)

##
###### Join Point = C
##

# Spelling Interaction with topography
delta_c_m1 <-  nlme(log_absolute_delta_power ~ SSquadp3xs(Age_min_c, a, b, xs),
                        data = delta_dat,                               
                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
                                     b ~ topography * SpellGroup, xs ~ topography * SpellGroup),
                        random = a ~ 1 | ID,
                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
                                   b = -4, rep(0, 4 + 1 + (1*4) + 1 - 1), 
                                   xs = 28, rep(0, 4 + 1 + (1*4) + 1 - 1)),
                        control = nlmeControl(msMaxIter = 2000))

# Spelling Interaction with topography and hemisphere
delta_c_m2 <-  nlme(log_absolute_delta_power ~ SSquadp3xs(Age_min_c, a, b, xs),
                        data = delta_dat,                               
                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
                                     b ~ topography * SpellGroup, xs ~ topography * SpellGroup * hemisphere),
                        random = a ~ 1 | ID,
                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
                                   b = -4, rep(0, 4 + 1 + (1*4) + 1 - 1), 
                                   xs = 28,  rep(0, 4 + 1 + 1 + (1*4) + (1*4) + (1*1)+ (1*1*4) + 1 - 1)),
                        control = nlmeControl(msMaxIter = 2000))

# Model comparison
anova(delta_b_m1, delta_c_m1, delta_c_m2)

# Best fitting model
paste0("Thus the best fitting mode is this (type III SS):")
anova(delta_b_m1, type = "marginal")

emmeans(delta_b_m1, ~ SpellGroup | topography, param = "a") %>% pairs()
emmeans(delta_b_m1, ~ topography | SpellGroup , param = "a") %>% pairs()
```


### Fitting Best Quadratic to Plateau Model to Theta Frequency Band


```{r Fitting QTP on Delta data}
# Extracting datasets of interest
theta_dat <- filter(final_dat, FB == "theta")


# Create an outcome with the log transformed version of power for data within each frequency band
theta_dat <- mutate(theta_dat, log_absolute_theta_power = log(absolute_power))


##
###### Intercept = a
##

# Spelling with Interactions for topography and hemisphere
theta_a_m3 <-  nlme(log_absolute_theta_power ~ SSquadp3xs(Age_min_c, a, b, xs),
                        data = theta_dat,                               
                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
                                     b ~ SpellGroup, xs ~ SpellGroup),
                        random = a ~ 1 | ID,
                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
                                   b = -4, 0, 
                                   xs = 28, 0),
                        control = nlmeControl(msMaxIter = 2000))

# Model comparison
anova(theta_a_m3, type = "marginal")


##
###### Slope = b (Causes Singularity ERROR)
##


# Spelling Interaction with topography
#theta_b_m1 <-  nlme(log_absolute_theta_power ~ SSquadp3xs(Age_min_c, a, b, xs),
#                        data = theta_dat,                               
#                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
#                                     b ~ topography * SpellGroup, xs ~ SpellGroup),
#                        random = a ~ 1 | ID,
#                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
#                                   b = -4, rep(0, 4 + 1 + (1*4) + 1 - 1), 
#                                   xs = 28, 0),
#                        control = nlmeControl(msMaxIter = 2000))

# Spelling Interaction with topography and hemisphere
#theta_b_m2 <-  nlme(log_absolute_theta_power ~ SSquadp3xs(Age_min_c, a, b, xs),
#                        data = theta_dat,                               
#                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
#                                     b ~ topography * hemisphere * SpellGroup, xs ~ SpellGroup),
#                        random = a ~ 1 | ID,
#                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
#                                   b = -4, rep(0, 4 + 1 + 1 + (1*4) + (1*4) + (1*1)+ (1*1*4) + 1 - 1), 
#                                   xs = 28, 0),
#                        control = nlmeControl(msMaxIter = 2000))

# Model comparison
# anova(theta_a_m3, theta_b_m1, theta_b_m2)

##
###### Join Point = C (Causes step halving factor reduced below minumum in PNLS step)
##

# Spelling Interaction with topography
#theta_c_m1 <-  nlme(log_absolute_theta_power ~ SSquadp3xs(Age_min_c, a, b, xs),
#                        data = theta_dat,                               
#                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
#                                     b ~ SpellGroup, xs ~ topography * SpellGroup),
#                        random = a ~ 1 | ID,
#                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
#                                   b = -4, 0, 
#                                   xs = 28, rep(0, 4 + 1 + (1*4) + 1 - 1)),
#                        control = nlmeControl(msMaxIter = 5000))



# Best fitting model
paste0("Thus the best fitting mode by default is this one that converged (type III SS):")
anova(theta_a_m3, type = "marginal")
```



### Fitting Best Quadratic to Plateau Model to Alpha1 Frequency Band


```{r Fitting QTP on Delta data}
# Extracting datasets of interest
alpha1_dat <- filter(final_dat, FB == "alpha1")

# Create an outcome with the log transformed version of power for data within each frequency band
alpha1_dat <- mutate(alpha1_dat, log_absolute_alpha1_power = log(absolute_power))


##
###### Intercept = a
##

# Spelling with Interactions for topography and hemisphere
alpha1_a_m3 <-  nlme(log_absolute_alpha1_power ~ SSquadp3xs(Age_min_c, a, b, xs),
                        data = alpha1_dat,                               
                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
                                     b ~ SpellGroup, xs ~ SpellGroup),
                        random = a ~ 1 | ID,
                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
                                   b = -4, 0, 
                                   xs = 28, 0),
                        control = nlmeControl(msMaxIter = 2000))

# Model comparison
anova(alpha1_a_m3, type = "marginal")


##
###### Slope = b (Causes Singularity ERROR)
##


# Spelling Interaction with topography
#alpha1_b_m1 <-  nlme(log_absolute_alpha1_power ~ SSquadp3xs(Age_min_c, a, b, xs),
#                        data = alpha1_dat,                               
#                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
#                                     b ~ topography * SpellGroup, xs ~ SpellGroup),
#                        random = a ~ 1 | ID,
#                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
#                                   b = -4, rep(0, 4 + 1 + (1*4) + 1 - 1), 
#                                   xs = 28, 0),
#                        control = nlmeControl(msMaxIter = 2000))

# Spelling Interaction with topography and hemisphere
#alpha1_b_m2 <-  nlme(log_absolute_alpha1_power ~ SSquadp3xs(Age_min_c, a, b, xs),
#                        data = alpha1_dat,                               
#                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
#                                     b ~ topography * hemisphere * SpellGroup, xs ~ SpellGroup),
#                        random = a ~ 1 | ID,
#                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
#                                   b = -4, rep(0, 4 + 1 + 1 + (1*4) + (1*4) + (1*1)+ (1*1*4) + 1 - 1), 
#                                   xs = 28, 0),
#                        control = nlmeControl(msMaxIter = 2000))

# Model comparison
# anova(alpha1_a_m3, alpha1_b_m1, alpha1_b_m2)

##
###### Join Point = C (Singularity in backsolve at level 0, block 1)
##

# Spelling Interaction with topography
#alpha1_c_m1 <-  nlme(log_absolute_alpha1_power ~ SSquadp3xs(Age_min_c, a, b, xs),
#                        data = alpha1_dat,                               
#                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
#                                     b ~ SpellGroup, xs ~ topography * SpellGroup),
#                        random = a ~ 1 | ID,
#                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
#                                   b = -4, 0, 
#                                   xs = 28, rep(0, 4 + 1 + (1*4) + 1 - 1)),
#                        control = nlmeControl(msMaxIter = 2000))


# Model comparison
# anova(alpha1_b_m1, alpha1_c_m1)

# Best fitting model
paste0("Thus the best fitting mode by default is this one that converged (type III SS):")
anova(alpha1_a_m3, type = "marginal")
```


### Fitting Best Quadratic to Plateau Model to Alpha2 Frequency Band


```{r Fitting QTP on Delta data}
# Extracting datasets of interest
alpha2_dat <- filter(final_dat, FB == "alpha2")

# Create an outcome with the log transformed version of power for data within each frequency band
alpha2_dat <- mutate(alpha2_dat, log_absolute_alpha2_power = log(absolute_power))


##
###### Intercept = a
##

# Spelling with Interactions for topography and hemisphere
alpha2_a_m3 <-  nlme(log_absolute_alpha2_power ~ SSquadp3xs(Age_min_c, a, b, xs),
                        data = alpha2_dat,                               
                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
                                     b ~ SpellGroup, xs ~ SpellGroup),
                        random = a ~ 1 | ID,
                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
                                   b = -4, 0, 
                                   xs = 28, 0),
                        control = nlmeControl(msMaxIter = 2000))

# Model comparison
anova(alpha2_a_m3, type = "marginal")


##
###### Slope = b (step halving factor reduced below minimum in PNLS step)
##


# Spelling Interaction with topography
#alpha2_b_m1 <-  nlme(log_absolute_alpha2_power ~ SSquadp3xs(Age_min_c, a, b, xs),
#                        data = alpha2_dat,                               
#                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
#                                     b ~ topography * SpellGroup, xs ~ SpellGroup),
#                        random = a ~ 1 | ID,
#                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
#                                   b = -4, rep(0, 4 + 1 + (1*4) + 1 - 1), 
#                                   xs = 28, 0),
#                        control = nlmeControl(msMaxIter = 2000))

# Spelling Interaction with topography and hemisphere
#alpha1_b_m2 <-  nlme(log_absolute_alpha1_power ~ SSquadp3xs(Age_min_c, a, b, xs),
#                        data = alpha1_dat,                               
#                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
#                                     b ~ topography * hemisphere * SpellGroup, xs ~ SpellGroup),
#                        random = a ~ 1 | ID,
#                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
#                                   b = -4, rep(0, 4 + 1 + 1 + (1*4) + (1*4) + (1*1)+ (1*1*4) + 1 - 1), 
#                                   xs = 28, 0),
#                        control = nlmeControl(msMaxIter = 2000))

# Model comparison
# anova(alpha1_a_m3, alpha1_b_m1, alpha1_b_m2)

##
###### Join Point = C (Causes step halving factor reduced below minumum in PNLS step)
##

# Spelling Interaction with topography
#alpha2_c_m1 <-  nlme(log_absolute_alpha2_power ~ SSquadp3xs(Age_min_c, a, b, xs),
#                        data = alpha2_dat,                               
#                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
#                                     b ~ SpellGroup, xs ~ topography * SpellGroup),
#                        random = a ~ 1 | ID,
#                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
#                                   b = -4, 0, 
#                                   xs = 28, rep(0, 4 + 1 + (1*4) + 1 - 1)),
#                        control = nlmeControl(msMaxIter = 2000))


# Model comparison
# anova(alpha1_b_m1, alpha1_c_m1)

# Best fitting model
paste0("Thus the best fitting mode by default is this one that converged (type III SS):")
anova(alpha2_a_m3, type = "marginal")
```


### Fitting Best Quadratic to Plateau Model to Beta1 Frequency Band


```{r Fitting QTP on Delta data}
# Extracting datasets of interest
beta1_dat <- filter(final_dat, FB == "beta1")

# Create an outcome with the log transformed version of power for data within each frequency band
beta1_dat <- mutate(beta1_dat, log_absolute_beta1_power = log(absolute_power))


##
###### Intercept = a
##

# Spelling with Interactions for topography and hemisphere
beta1_a_m3 <-  nlme(log_absolute_beta1_power ~ SSquadp3xs(Age_min_c, a, b, xs),
                        data = beta1_dat,                               
                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
                                     b ~ SpellGroup, xs ~ SpellGroup),
                        random = a ~ 1 | ID,
                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
                                   b = -4, 0, 
                                   xs = 28, 0),
                        control = nlmeControl(msMaxIter = 2000))

##
###### Slope = b (step halving factor reduced below minimum in PNLS step)
##


# Spelling Interaction with topography
beta1_b_m1 <-  nlme(log_absolute_beta1_power ~ SSquadp3xs(Age_min_c, a, b, xs),
                        data = beta1_dat,                               
                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
                                     b ~ topography * SpellGroup, xs ~ SpellGroup),
                        random = a ~ 1 | ID,
                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
                                   b = -4, rep(0, 4 + 1 + (1*4) + 1 - 1), 
                                   xs = 28, 0),
                        control = nlmeControl(msMaxIter = 2000))


# Spelling Interaction with topography and hemisphere
beta1_b_m2 <-  nlme(log_absolute_beta1_power ~ SSquadp3xs(Age_min_c, a, b, xs),
                        data = beta1_dat,                               
                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
                                     b ~ topography * hemisphere * SpellGroup, xs ~ SpellGroup),
                        random = a ~ 1 | ID,
                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
                                   b = -4, rep(0, 4 + 1 + 1 + (1*4) + (1*4) + (1*1)+ (1*1*4) + 1 - 1), 
                                   xs = 28, 0),
                        control = nlmeControl(msMaxIter = 2000))

# Model comparison
anova(beta1_a_m3, beta1_b_m1, beta1_b_m2)

##
###### Join Point = C (Singularity in backsolve at level 0, block 1)
##

# Spelling Interaction with topography
#alpha2_c_m1 <-  nlme(log_absolute_beta1_power ~ SSquadp3xs(Age_min_c, a, b, xs),
#                        data = beta1_dat,                               
#                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
#                                     b ~ topography * SpellGroup, xs ~ topography * SpellGroup),
#                        random = a ~ 1 | ID,
#                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
#                                   b = -4, rep(0, 4 + 1 + (1*4) + 1 - 1), 
#                                   xs = 28, rep(0, 4 + 1 + (1*4) + 1 - 1)),
#                        control = nlmeControl(msMaxIter = 2000))



# Best fitting model
paste0("Thus the best fitting mode by default is this one (type III SS):")
anova(beta1_b_m1, type = "marginal")

```


### Fitting Best Quadratic to Plateau Model to Beta2 Frequency Band


Can't even run a single model

```{r Fitting QTP on Delta data}
# Extracting datasets of interest
beta2_dat <- filter(final_dat, FB == "beta2")

# Create an outcome with the log transformed version of power for data within each frequency band
beta2_dat <- mutate(beta2_dat, log_absolute_beta2_power = log(absolute_power))


##
###### Intercept = a (Singularity in backsolve at level 0, block 1)
##

# Main effects only
#beta2_a_m1 <-  nlme(log_absolute_beta2_power ~ SSquadp3xs(Age_min_c, a, b, xs),
#                        data = beta2_dat,                               
#                        fixed = list(a ~ Sex + hemisphere + topography + SpellGroup, 
#                                     b ~ SpellGroup, xs ~ SpellGroup),
#                        random = a ~ 1 | ID,
#                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + 1 - 1),
#                                   b = -4, 0, 
#                                   xs = 28, 0),
#                        control = nlmeControl(msMaxIter = 2000))

# Spelling with Interactions for topography
#beta2_a_m2 <-  nlme(log_absolute_beta2_power ~ SSquadp3xs(Age_min_c, a, b, xs),
#                        data = beta2_dat,                               
#                        fixed = list(a ~ Sex + hemisphere + topography * SpellGroup, 
#                                     b ~ SpellGroup, xs ~ SpellGroup),
#                        random = a ~ 1 | ID,
#                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (4*1) + 1 - 1),
#                                   b = -4, 0, 
#                                   xs = 28, 0),
#                        control = nlmeControl(msMaxIter = 2000))

# Spelling with Interactions for topography and hemisphere
#beta2_a_m3 <-  nlme(log_absolute_beta2_power ~ SSquadp3xs(Age_min_c, a, b, xs),
#                        data = beta2_dat,                               
#                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
#                                     b ~ SpellGroup, xs ~ SpellGroup),
#                        random = a ~ 1 | ID,
#                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
#                                   b = -4, 0, 
#                                   xs = 28, 0),
#                        control = nlmeControl(msMaxIter = 2000))

##
###### Slope = b (step halving factor reduced below minimum in PNLS step)
##


# Spelling Interaction with topography
#beta1_b_m1 <-  nlme(log_absolute_beta1_power ~ SSquadp3xs(Age_min_c, a, b, xs),
#                        data = beta1_dat,                               
#                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
#                                     b ~ topography * SpellGroup, xs ~ SpellGroup),
#                        random = a ~ 1 | ID,
#                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
#                                   b = -4, rep(0, 4 + 1 + (1*4) + 1 - 1), 
#                                  xs = 28, 0),
#                        control = nlmeControl(msMaxIter = 2000))


# Spelling Interaction with topography and hemisphere
#beta1_b_m2 <-  nlme(log_absolute_beta1_power ~ SSquadp3xs(Age_min_c, a, b, xs),
##                        data = beta1_dat,                               
#                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
#                                     b ~ topography * hemisphere * SpellGroup, xs ~ SpellGroup),
#                        random = a ~ 1 | ID,
#                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
#                                   b = -4, rep(0, 4 + 1 + 1 + (1*4) + (1*4) + (1*1)+ (1*1*4) + 1 - 1), 
#                                   xs = 28, 0),
#                        control = nlmeControl(msMaxIter = 2000))

# Model comparison
# anova(beta1_a_m3, beta1_b_m1, beta1_b_m2)

##
###### Join Point = C (Singularity in backsolve at level 0, block 1)
##

# Spelling Interaction with topography
#alpha2_c_m1 <-  nlme(log_absolute_beta1_power ~ SSquadp3xs(Age_min_c, a, b, xs),
#                        data = beta1_dat,                               
#                        fixed = list(a ~ Sex + hemisphere * topography * SpellGroup, 
#                                     b ~ topography * SpellGroup, xs ~ topography * SpellGroup),
#                        random = a ~ 1 | ID,
#                        start = c(a = 3, rep(0, 1 + 1 + 4 + 1 + (1*4) + (1*4) + (1*1) +  (1*1*4) + 1 - 1),
#                                   b = -4, rep(0, 4 + 1 + (1*4) + 1 - 1), 
#                                   xs = 28, rep(0, 4 + 1 + (1*4) + 1 - 1)),
#                        control = nlmeControl(msMaxIter = 2000))



# Best fitting model
#paste0("Thus the best fitting mode by default is this one (type III SS):")
#anova(beta1_b_m1, type = "marginal")

```



# Visualizing Model Results 


```{r Visualizing Model Results}
###
####### Delta
###

# Show the spellgroup and topography interaction for the intercept
emmip(delta_b_m1, SpellGroup ~ topography, param = "a", CIs = TRUE) +
  labs(title = "Spelling Group Differences on the Intercept") +
  theme_clean()

# Show the spellgroup and topography interaction for the slopes
emmip(delta_b_m1, SpellGroup ~ topography, param = "b", CIs = TRUE) +
  labs(title = "Spelling Group Differences on the Instantaneous Slope") +
  theme_clean()


###
####### Theta
###

# No significant two-way interactions



###
######## Alpha1
###


```








### Printing a Table for the Model Output

```{r Printing a table for the model ouput}
# Extract the fixed effects from the model.
slow_fix <- tidy(final_slow_mod) %>%
  mutate_if(is.numeric, ~round(., 3)) %>%
  select(-effect)

##
#### Correcting for multiple comparison
##

# Part1: get the index for topographic rows
topo_rows <- which(grepl("topography", slow_fix$term))

# Part2: split indices by parameter types
topo_pvals_a <- slow_fix$p.value[topo_rows[1:3]]
topo_pvals_b <- slow_fix$p.value[topo_rows[4:6]]

# Part 3: adjust the p-values using Holm-Bonferroni (stronger than Bonferroni)
topo_pvals_a_adj <- p.adjust(topo_pvals_a, method = "holm")
topo_pvals_b_adj <- p.adjust(topo_pvals_b, method = "holm") 

# IMPORTANT: Nothing really changed because 2 of the 3 p-values were already really small

# Test that it worked 
#p.adjust(c(0.02, 0.3, 0.01), method = "holm") %>% round(3)
#p.adjust(c(0.001, 0.19, 0.00001), method = "holm") %>% round(3)

# Introduced corrected-pvalues into the data
slow_fix$p.value[topo_rows[1:3]] <- topo_pvals_a_adj
slow_fix$p.value[topo_rows[4:6]] <- topo_pvals_b_adj


# Make the estimates prettier
slow_fix <- slow_fix %>%
    mutate(term = case_when(
        grepl("ecfrontal", term)          ~ "Frontal (Effect)",
        grepl("ecoccipital", term)        ~ "Occipital (Effect)",
        grepl("ectemporal", term)         ~ "Temporal (Effect)",
        grepl("a.Spell|b.Spell", term)          ~ "Poor Spellers (Effect)",
        grepl("a.(Intercept)", term, fixed = TRUE)  ~ "Good Spellers (Topography Grand Mean)",
        grepl("b.(Intercept)", term, fixed = TRUE)  ~ "Good Spellers (Topography Grand Mean)",
        grepl("xs.(Intercept)", term, fixed = TRUE) ~ "Good Spellers",
        grepl("xs.SpellGroup", term) ~ "Poor Spellers (Effect)",
    ))

# Change the names of the variables
names(slow_fix) <- c("Variable", "Estimate", "SE", "DF", "t", "p")

# Create a function for significance
p_function <- function(p) {
  ifelse(p < .001, "<.001***",
         ifelse(p < .01, paste0(as.character(p),"**"),
                ifelse(p <.05, paste0(as.character(p),"*"), 
                       as.character(p))))
}

slow_fix$p <- p_function(slow_fix$p)

# Uncentered Title
slow_fix %>%
  kbl(caption = "<span style='color:black; font-size:16px;'>Slow Wave: Quadratic to Plateau Model Output</span>",
      escape = FALSE) %>%
  kable_paper(full_width = F) %>%
  pack_rows("Intercept (a)", 1, 5) %>%
  pack_rows("Instantaneous Slope (b)", 6, 10) %>%
  pack_rows("Join Point (xs)", 11, 12)

# Centered Title
slow_fix %>%
  kbl(caption = "<span style='color:black; font-size:16px; display:block; text-align:center;'>Slow Wave: Quadratic to Plateau Model Output</span>",
      escape = FALSE) %>%
  kable_paper(full_width = F) %>%
  pack_rows("Intercept", 1, 5) %>%
  pack_rows("Instantaneous Slope", 6, 10) %>%
  pack_rows("Join Point", 11, 12) %>%
  footnote("Topography effects display adjusted p-values using the Holm-Bonferroni method within each parameter section.")
  
```




### Results of 'Slow' absolute power as a function of age with Topography Between Spelling Group (Quadratic to Plateau Model)

The best fitting model was the one that still had slopes estimated for k-1 topography levels, so that is the one that we will be interpreting below. However, since this is our final model for the analysis, we will be interpreting the model not using the p-values directly from the output table, but the ones that we obtain using follow-up tests. This is so we can add adjustments to correct for multiple comparison. Additionally this will allow us to generate estimates for all levels of topography rather than just k-1 levels. Honestly, using **marginal effects** below, all previous models can essentially be ignored. Their reasoning for specification was to compare the results from these model back to them, to see if there are any weird discrepancies (ex: having a positive slope rather than negative or something crazy like that). 

Unfortunately, the omnibus test that we would typically use, aka the one from the Anova() function does not work here, so we will be ignoring that.

Okay so there is a lot of information down below, we will just make sense of the follow-up tests rather than the original model summary output. However, what is good about the model output is that it shows 1408 observations and 352 groups, which aligns with our dataset.

**Parameter a: The Intercept**

- Starting with **topography**, the estimated marginal means are showing the mean log absolute power for each level of topography when centered age is equal to 0 (age is 15 years). We see that all of them except for parietal are 2 log absolute power or larger. We can then do pairwise comparisons to test whether log absolute power differences between levels of topography. Looking at the p-values, we see that there are significant differences between frontal and parietal, temporal and occipital, temporal and parietal, and occipital and parietal. Essentially, from this we can see that in the parietal region, log absolute power tends to be lower overall compared to the other regions after adjusting for multiple comparisons (tukey)
- Next with **spelling group**, we can see the estimates between poor (2.03) and good (2.10) spellers. Instantly we can see that the values are very close to each other, not only that but **unexpectedly**, there is more log absolute power for good spellers than poor spellers. Next we investigate whether these means when centered age is 0 (age is 15 years) are significantly different from each other, and unsurprisingly they are not (p = 0.5844).

**Parameter b: The Slope**

- **Note**: The following slope estimates where all calculated from age centered being equal to 0 (age is 15 years). To do this, we created a reference grid that specified age centered being equal to 0, and then using that grid to obtain estimated slopes. This is important because failing to do this step would have produced either a) derivatives (slopes) averaged across all data points which include all age data points or b) an estimated slope at the sample mean of age centered. Those would have been **main effects**, but we don't really care for that since what we really want is the slope that corresponds with age centered being at 0.

- Starting with **topography**, we get the **simple slopes** of all levels of topography, where they are all negative and are all significant, indicating that with age log absolute power is decreasing, as we would expect. Next we use pairwise comparisons to see if the rate of change differs by topographical region. After controlling for multiple comparison using the Benjamini-Hochberg (BH) procedure, we found no differences in slope comparisons.
- Next with **spelling group**, we see that the simple slopes for poor spellers is -0.0691 and good spellers is -0.0766, which look similar, with the slope being slightly stronger for good spellers than poor spellers. Not surprisingly, the slope differences between the two were not significant.


**Parameter xs: The Join Point**

- We only modeled the join point using the **spell group** factor. We can see that the age for log absolute power plateau in poor spellers is 21.8 + 15 = 36.8 and in good spellers is 17.9 + 15 = 32.9. Again these numbers do not really make sense and the standard errors are large, especially for the poor spelling group with was 17.10 unlike the good spellers which was 5.52. How large these standard errors are really undermine the model entirely I would say. Regardles there was no significant difference in the join point between the two groups. 


```{r Results of the age with topography between spelling groups quadratic to plateau model}
# Print out the summary of the best fitting model
summary(final_slow_mod)

# Follow up tests on the intercepts (estimated marginal means)
(emm_top_a <- emmeans(final_slow_mod, pairwise ~ topography_ec, param = "a", infer = TRUE, adjust = "tukey"))
(emm_grp_a <- emmeans(final_slow_mod, pairwise ~ SpellGroup, param = "a", infer = TRUE, adjust = "tukey"))

# Create a dataset to get the slopes when Age is equal to 0
age_0_dat <- datagrid(Age_min_c = 0,
                      topography_ec = levels(slow_dat$topography_ec),
                      SpellGroup = levels(slow_dat$SpellGroup),
                      model = final_slow_mod)

# Follow up tests on the slopes
avg_slopes(final_slow_mod, variables = "Age_min_c", by = "topography_ec", newdata = age_0_dat)
avg_slopes(final_slow_mod, variables = "Age_min_c", by = "topography_ec", newdata = age_0_dat,  hypothesis = ~ pairwise) %>%
  hypotheses(multcomp = "holm")
avg_slopes(final_slow_mod, variables = "Age_min_c", by = "SpellGroup", newdata = age_0_dat)
avg_slopes(final_slow_mod, variables = "Age_min_c", by = "SpellGroup", newdata = age_0_dat,  hypothesis = ~ pairwise)

# Follow up tests on the 
(emm_grp_xs <- emmeans(final_slow_mod, pairwise ~ SpellGroup, param = "xs", infer = TRUE, adjust = "tukey"))
```
Just like before we can obtain fitted values of the model to create a scatterplot showing the relationships. We do this by using the `expand.grid()` function, where we can make different level combinations of all three predictors of interest: age centered, topography, and spell group. With this we can basically generate 100 observations for each level combination of the factors and confirm that this was correct using `xtabs()`.



```{r Visualizing the fitted values of the final model}
# Create a grid that has the predictor values we want for corresponding fitted values
newdat_final <- expand.grid(
  Age_min_c = seq(min(slow_dat$Age_min_c), max(slow_dat$Age_min_c), length.out = 100),
  topography_ec = levels(slow_dat$topography_ec),
  SpellGroup = levels(slow_dat$SpellGroup)
  
)

# Inspect the reference grid created
xtabs(~ SpellGroup + topography_ec , newdat_final)

# Add the predicted values back into the grid + adjust the age and relevel topography
newdat_final <- newdat_final %>%
  mutate(log_absolute_power = predict(final_slow_mod, newdata = newdat_final, level = 0),
         Age = Age_min_c + 15,
         topography = factor(topography_ec,
                             levels = c("frontal", "temporal", "parietal", "occipital")))

# Creating a pretty scatter plot
slow_dat %>%
  mutate(SpellGroup = factor(SpellGroup, levels = c("Poor", "Good"))) %>%
  ggplot(aes(x = Age, y = log_absolute_slow_power, color = SpellGroup, shape = SpellGroup)) +
  geom_point(size = 1, alpha = .7) +
  geom_line(data = newdat_final, aes(x = Age, y = log_absolute_power), size = 1) + 
  facet_wrap(~topography, scales = "free") +
  labs(title = "Fitted Values of Quadratic to Plateau Model Showcasing\n Log of Absolute Power Regressed onto Age by Topography\nBetween Spelling Groups for Slow Frequencies",
       y = "log of absolute power\n(slow frequencies)",
       color = "Spelling Groups",
       shape = "Spelling Groups") +
  theme_clean()

```


### Quadratic to Plateau Model Residuals Investigation 


```{r Investigate the model fit of the quadratic to plateau model}
# Extract residuals and fitted values from the Quadratic to Plateau Model
resid_norm     <- residuals(final_slow_mod, type = "normalized")
resid_pearson  <- residuals(final_slow_mod, type = "pearson")
resid_response <- residuals(final_slow_mod, type = "response")
fitted_vals    <- fitted(final_slow_mod)

# Introduce the vectors into the 'slow_dat' dataset
diag_df <- slow_dat %>%
  mutate(
    resid_norm     = resid_norm,
    resid_pearson  = resid_pearson,
    resid_response = resid_response,
    fitted         = fitted_vals
  )

# 1. QQ Plot
p1 <- ggplot(diag_df, aes(sample = resid_norm)) +
  stat_qq() +
  stat_qq_line(color = "firebrick", linewidth = 1.8) +
  labs(title = "QQ-plot of Standardized Residuals",
       x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_clean()

# 2. Histogram with N(0,1) density overlay
p2 <- ggplot(diag_df, aes(x = resid_norm)) +
  geom_histogram(aes(y = after_stat(density)), bins = 25,
                 fill = "grey80", color = "white") +
  stat_function(fun = dnorm, color = "red", linewidth = 1.8) +
  labs(title = "Standardized Residuals Density Plot",
       x = "Standardized Residual", y = "Density") +
  theme_clean()

# 3. Residuals vs Fitted (standardized)
p3 <- ggplot(diag_df, aes(x = fitted, y = resid_norm)) +
  geom_point(alpha = 0.4) +
  geom_hline(yintercept = 0, color = "firebrick", linewidth = 1) +
  labs(title = "Residuals vs Fitted (Standardized)",
       x = "Fitted Values", y = "Standardized Residuals") +
  theme_clean()

# 4. Response residuals vs Age by SpellGroup
p4 <- ggplot(diag_df, aes(x = Age, y = resid_response)) +
  geom_point(alpha = 0.4) +
  geom_hline(yintercept = 0, color = "firebrick", linewidth = 1) +
  facet_wrap(~ SpellGroup) +
  labs(title = "Response Residuals vs Age by Spelling Group",
       x = "Age", y = "Response Residuals") +
  theme_clean()


# Combined diagnostic panel
# Combined diagnostic panel
cs_p <- (p1 | p2) / p3 / p4 +
  plot_annotation(
    title = "A: Original Model Residuals",
    theme = theme(
      plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
    )
  )

cs_p

```


### Modeling the Error Structure


```{r Modeling the error structure}
# Our model already kinda has compound symmetric errors
# Do not run the following code, it produces the same output at the cost of an additional parameter
# final_slow_mod_cs <- update(final_slow_mod, correlation = corCompSymm(form = ~1 | ID))

# Produce the conditional residual correlations between topography
cond_resid <- final_slow_mod$residuals[,2]
cond_resid_matrix <- matrix(cond_resid, nrow = 352, ncol = 4, byrow = TRUE)
(top_res_cov <- round(cov(cond_resid_matrix), 3))
(top_res_cor <- round(cor(cond_resid_matrix), 3))

# Average pairwise correlations between topographies residuals
mean(top_res_cov[upper.tri(top_res_cov)])
mean(top_res_cor[upper.tri(top_res_cor)])

# Update the model with heterogenous variance
final_slow_mod_hcs <- update(final_slow_mod,
                          weights = varIdent(form = ~1 | topography_ec))

# Creating a model with correlated residuals but homogenous variances
empty_lme_cor_res <- update(final_slow_mod,
                           correlation = corSymm(form = ~1 | ID))

# Update to Unstructured Errors
final_slow_mod_un <- update(final_slow_mod,
                            weights = varIdent(form = ~1 | topography_ec),
                            correlation = corSymm(form = ~1 | ID))

# Model comparison on the modeleted error structures
anova(final_slow_mod, final_slow_mod_hcs, empty_lme_cor_res, final_slow_mod_un) # Unstructured wins

# View the residuals of this model - we can see the reduction in variances!
p_res_cs <- residuals(final_slow_mod, type = "normalize")# Now make the matrix and look at cov/cor
std_matrix_cs <- matrix(p_res_cs, nrow = 352 , ncol = 4, byrow = TRUE)
round(cov(std_matrix_cs), 3)

p_res_hcs <- residuals(final_slow_mod_un, type = "normalize")# Now make the matrix and look at cov/cor
std_matrix_hcs <- matrix(p_res_hcs, nrow = 352 , ncol = 4, byrow = TRUE)
round(cov(std_matrix_hcs), 3)

```

### Investigating the residuals after modeling the errors as Unstructured

### Quadratic to Plateau Model Residuals Investigation 


```{r Investigate the model fit of the quadratic to plateau model Unstructured Errors}
# Extract residuals and fitted values from the Quadratic to Plateau Model
resid_norm     <- residuals(final_slow_mod_un, type = "normalized")
resid_pearson  <- residuals(final_slow_mod_un, type = "pearson")
resid_response <- residuals(final_slow_mod_un, type = "response")
fitted_vals    <- fitted(final_slow_mod_un)

# Introduce the vectors into the 'slow_dat' dataset
diag_df2 <- slow_dat %>%
  mutate(
    resid_norm     = resid_norm,
    resid_pearson  = resid_pearson,
    resid_response = resid_response,
    fitted         = fitted_vals
  )

# 1. QQ Plot
p1 <- ggplot(diag_df2, aes(sample = resid_norm)) +
  stat_qq() +
  stat_qq_line(color = "firebrick", linewidth = 1.8) +
  labs(title = "QQ-plot of Standardized Residuals",
       x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_clean()

# 2. Histogram with N(0,1) density overlay
p2 <- ggplot(diag_df2, aes(x = resid_norm)) +
  geom_histogram(aes(y = after_stat(density)), bins = 25,
                 fill = "grey80", color = "white") +
  stat_function(fun = dnorm, color = "red", linewidth = 1.8) +
  labs(title = "Standardized Residuals Density Plot",
       x = "Standardized Residual", y = "Density") +
  theme_clean()

# 3. Residuals vs Fitted (standardized)
p3 <- ggplot(diag_df2, aes(x = fitted, y = resid_norm)) +
  geom_point(alpha = 0.4) +
  geom_hline(yintercept = 0, color = "firebrick", linewidth = 1) +
  labs(title = "Residuals vs Fitted (Standardized)",
       x = "Fitted Values", y = "Standardized Residuals") +
  theme_clean()

# 4. Pearson residuals vs Age by SpellGroup
p4 <- ggplot(diag_df2, aes(x = Age, y = resid_norm)) +
  geom_point(alpha = 0.4) +
  geom_hline(yintercept = 0, color = "firebrick", linewidth = 1) +
  facet_wrap(~ SpellGroup) +
  labs(title = "Normalized Residuals vs Age by Spelling Group",
       x = "Age", y = "Normalized Residuals") +
  theme_clean()


# Combined diagnostic panel
un_p <- (p1 | p2) / p3 / p4 +
  plot_annotation(
    title = "B: Unstructured Model Residuals",
    theme = theme(
      plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
    )
  )

un_p
```



### Investigating differences in absolute between Spelling Groups in 'Alpha' and 'Beta' frequencies

Each model was fit using **maximum likelihood**, which makes model comparison comparable. We explicitly state maximum likelihood below since the default is restricted maximum likelihood. Regardless of approach to estimation, with technically REML being less biased, at a large enough sample size (which we have), the estimates between the models are essentially the same.

```{r Investigating differneces in absolute power for alpha and beta}
# Filter the data by frequency
alpha_dat <- final_dat %>%
  filter(FB == "alpha") %>%
  mutate(log_absolute_alpha_power = log(absolute_power))

beta_dat <- final_dat %>%
  filter(FB == "beta") %>%
  mutate(log_absolute_beta_power = log(absolute_power))

# Specify an alpha model with all variables of interest
a_mod1 <- lme(fixed = log_absolute_alpha_power ~ Age_min_c + topography_ec + SpellGroup,
              random = ~ 1|ID,
              data = alpha_dat, 
              method = "ML") 

# Specify another alpha model with a quadratic term
a_mod1_quad <- update(a_mod1, . ~. + I(Age_min_c^2))

# Model comparison to see if the quadratic produced better fit
anova(a_mod1, a_mod1_quad)

# Introduce alpha models with interactions because why not
a_mod2 <- update(a_mod1, . ~. + Age_min_c * SpellGroup)
a_mod3 <- update(a_mod1, . ~. + Age_min_c * topography_ec)
a_mod4 <- update(a_mod1, . ~. + Age_min_c * topography_ec * SpellGroup)

# Model comparison to see if any interaction led to better model fit
anova(a_mod1, a_mod2, a_mod3, a_mod4)


# Specify a beta model with all variables of interest
b_mod1 <- lme(fixed = log_absolute_beta_power ~ Age_min_c + topography_ec + SpellGroup,
              random = ~ 1|ID,
              data = beta_dat,
              method = "ML") 

# Specify another alpha model with a quadratic term
b_mod1_quad <- update(b_mod1, . ~. + I(Age_min_c^2))

# Model comparison to see if the quadratic produced better fit
anova(b_mod1, b_mod1_quad)

# Introduce beta models with interactions because why not
b_mod2 <- update(b_mod1, . ~. + Age_min_c * SpellGroup)
b_mod3 <- update(b_mod1, . ~. + Age_min_c * topography_ec)
b_mod4 <- update(b_mod1, . ~. + Age_min_c * topography_ec * SpellGroup)

# Model comparison to see if any interaction led to better model fit
anova(b_mod1, b_mod2, b_mod3, b_mod4)

```
### Printing out the best fitting models for higher frequencies


```{r Printing out the best fitting models for higher frequencies}
# Create a table with combined fixed effects from both models
alpha_beta_fix <- rbind(

tidy(a_mod1) %>%
  drop_na(p.value) %>%
  select(-group, -effect),

tidy(b_mod1) %>%
  drop_na(p.value) %>%
  select(-group, -effect)

) %>%
  mutate_if(is.numeric, ~round(., 3))

##
#### Correcting for multiple comparison
##

# Part1: get the index for topographic rows
topo_rows <- which(grepl("topography", alpha_beta_fix$term))

# Part2: split indices by parameter types
topo_pvals_a <- alpha_beta_fix$p.value[topo_rows[1:3]]
topo_pvals_b <- alpha_beta_fix$p.value[topo_rows[4:6]]

# Part 3: adjust the p-values using Holm-Bonferroni (stronger than Bonferroni)
topo_pvals_a_adj <- p.adjust(topo_pvals_a, method = "holm")
topo_pvals_b_adj <- p.adjust(topo_pvals_b, method = "holm") 

# IMPORTANT: Nothing really changed because 2 of the 3 p-values were already really small

# Test that it worked 
#p.adjust(c(0.02, 0.3, 0.01), method = "holm") %>% round(3)
#p.adjust(c(0.001, 0.19, 0.00001), method = "holm") %>% round(3)

# Introduced corrected-pvalues into the data
alpha_beta_fix$p.value[topo_rows[1:3]] <- topo_pvals_a_adj
alpha_beta_fix$p.value[topo_rows[4:6]] <- topo_pvals_b_adj

# Make the estimates prettier
alpha_beta_fix <- alpha_beta_fix %>%
    mutate(term = case_when(
        grepl("(Intercept)", term, fixed = TRUE) ~ "Intercept",
        grepl("Age_min_c", term) ~ "Age (centered at 14)",
        grepl("ecfrontal", term)          ~ "Frontal",
        grepl("ecoccipital", term)        ~ "Occipital",
        grepl("ectemporal", term)         ~ "Temporal",
        grepl("SpellGroupPoor", term)          ~ "Poor Spellers"
    ))


# Change the names of the variables
names(alpha_beta_fix) <- c("Variable", "Estimate", "SE", "DF", "t", "p")

# Make p-values prettier
alpha_beta_fix$p <- p_function(alpha_beta_fix$p)

# Centered Title
alpha_beta_fix %>%
  kbl(caption = "<span style='color:black; font-size:16px; display:block; text-align:center;'>Alpha and Beta Waves: Linear Model Output</span>",
      escape = FALSE) %>%
  kable_paper(full_width = F) %>%
  pack_rows("Alpha Model", 1, 6) %>%
  pack_rows("Beta Model", 7, 12) %>%
  footnote("Topography effects display adjusted p-values using the Holm-Bonferroni method within each model.")

```

### Inspecting the Model Residuals of the High Frequency Models


```{r Inspecting Model Residuals}
resid_norm     <- residuals(a_mod1, type = "normalized")
resid_pearson  <- residuals(a_mod1, type = "pearson")
resid_response <- residuals(a_mod1, type = "response")
fitted_vals    <- fitted(a_mod1)

# Introduce the vectors into the 'slow_dat' dataset
diag_df2 <- alpha_dat %>%
  mutate(
    resid_norm     = resid_norm,
    resid_pearson  = resid_pearson,
    resid_response = resid_response,
    fitted         = fitted_vals
  )

# 1. QQ Plot
p1 <- ggplot(diag_df2, aes(sample = resid_norm)) +
  stat_qq() +
  stat_qq_line(color = "firebrick", linewidth = 1.8) +
  labs(title = "QQ-plot of Standardized Residuals",
       x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_clean()

p2 <- ggplot(diag_df2, aes(x = resid_norm)) +
  geom_histogram(aes(y = after_stat(density)), bins = 25,
                 fill = "grey80", color = "white") +
  stat_function(fun = dnorm, color = "red", linewidth = 1.8) +
  labs(title = "Standardized Residuals Density Plot",
       x = "Standardized Residual", y = "Density") +
  theme_clean()

# 3. Residuals vs Fitted (standardized)
p3 <- ggplot(diag_df2, aes(x = fitted, y = resid_norm)) +
  geom_point(alpha = 0.4) +
  geom_hline(yintercept = 0, color = "firebrick", linewidth = 1) +
  labs(title = "Residuals vs Fitted (Standardized)",
       x = "Fitted Values", y = "Standardized Residuals") +
  theme_clean()

# Print out the alpha compound symmetric error diagnostic plots
cs_alpha <- p1 + p2 + p3


# Obtain different residuals and fitted values for beta model
resid_norm     <- residuals(b_mod1, type = "normalized")
resid_pearson  <- residuals(b_mod1, type = "pearson")
resid_response <- residuals(b_mod1, type = "response")
fitted_vals    <- fitted(b_mod1)

# Introduce the vectors into the 'slow_dat' dataset
diag_df2 <- beta_dat %>%
  mutate(
    resid_norm     = resid_norm,
    resid_pearson  = resid_pearson,
    resid_response = resid_response,
    fitted         = fitted_vals
  )

# 1. QQ Plot
p1 <- ggplot(diag_df2, aes(sample = resid_norm)) +
  stat_qq() +
  stat_qq_line(color = "firebrick", linewidth = 1.8) +
  labs(title = "QQ-plot of Standardized Residuals",
       x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_clean()

p2 <- ggplot(diag_df2, aes(x = resid_norm)) +
  geom_histogram(aes(y = after_stat(density)), bins = 25,
                 fill = "grey80", color = "white") +
  stat_function(fun = dnorm, color = "red", linewidth = 1.8) +
  labs(title = "Standardized Residuals Density Plot",
       x = "Standardized Residual", y = "Density") +
  theme_clean()

# 3. Residuals vs Fitted (standardized)
p3 <- ggplot(diag_df2, aes(x = fitted, y = resid_norm)) +
  geom_point(alpha = 0.4) +
  geom_hline(yintercept = 0, color = "firebrick", linewidth = 1) +
  labs(title = "Residuals vs Fitted (Standardized)",
       x = "Fitted Values", y = "Standardized Residuals") +
  theme_clean()

# Print out the alpha compound symmetric error diagnostic plots
cs_beta <- p1 + p2 + p3
```


### Introducing Error Modeling for Both Models

```{r Introducing error modeling}
###
##### ALPHA
###

# Update the model with heterogenous variance
a_mod1_hcs <- update(a_mod1,
                     weights = varIdent(form = ~1 | topography_ec))

# Creating a model with correlated residuals but homogenous variances
a_mod1_cor_res <- update(a_mod1,
                        correlation = corSymm(form = ~1 | ID))

# Update to Unstructured Errors (Increased iterations to prevent convergence failure)
a_mod1_un <- update(a_mod1,
                   weights = varIdent(form = ~1 | topography_ec),
                   correlation = corSymm(form = ~1 | ID),
                   control = nlmeControl(msMaxIter = 2000))

# Model comparison on the modeled error structures
anova(a_mod1, a_mod1_hcs, a_mod1_cor_res, a_mod1_un) # Unstructured wins

###
##### BETA
###


# Update the model with heterogenous variance
b_mod1_hcs <- update(b_mod1,
                     weights = varIdent(form = ~1 | topography_ec))

# Creating a model with correlated residuals but homogenous variances
b_mod1_cor_res <- update(b_mod1,
                        correlation = corSymm(form = ~1 | ID))

# Update to Unstructured Errors
b_mod1_un <- update(b_mod1,
                    weights = varIdent(form = ~1 | topography_ec),
                    correlation = corSymm(form = ~1 | ID),
                    control = nlmeControl(msMaxIter = 2000))

# Model comparison on the modeled error structures
anova(b_mod1, b_mod1_hcs, b_mod1_cor_res, b_mod1_un) # Unstructured wins
```