---
title: "Analyzing rsEEG and FB Power across Topography Part 2"
author: "Leandro Ledesma"
date: "2024-10-11"
output: html_document
---
### Universal block code settings

```{r setup}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(comment = NULL)
knitr::opts_chunk$set(warning = FALSE)

```



```{r loading in the packages, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(readxl)
library(kableExtra)
library(stringdist)
library(broom) # Converts regression outputs into dataframes using the tidy() function
library(psych)
library(corrtable) # Potential correlation matrix code
library(quantreg) # rq (quantile regreeion)
library(MASS, exclude = "select") # This package is loaded with QuantPsyc, must exclude "select" or you wont be able to use it. 
library(QuantPsyc) # Can use the lm.beta function to calculate the standardized betas
library(car) # To calculate VIF
library(purrr) # Allows for the usage of the map function
library(interactions)
```



```{r load in all variables, warning= FALSE}
# Set the working directory
setwd("~/MegaGrant")

# Load in the ARFA performance measures
ARFA <- read.csv("Processed/BehavioralData/ARFA.Spelling.Errors.Scored.csv")

# Add Spelling ERror Variable
ARFA <- select(ARFA, ID, SpellingError = Total_SpellingError)

# Load in the CFIT data
CFIT <- read.csv("Processed/BehavioralData/CFIT.scores.csv")

# Data cleaning
CFIT <- select(CFIT, ID, CFIT = Raw.Scores)

# Load in the rsEEG datasets
eyes.closed <- read_excel("Processed/EEGData/EEG Eyes Open and Closed FB Power (Whitford, 2007).xlsx")

# load in demographic information
demo <- read_excel("MegaGrant_TBL.xlsx")

# data cleaning
demo <- demo %>%
  select(ID, 
         Sex, 
         Age, 
         Group) %>%
  mutate(ID = as.numeric(ID),
         Age = as.numeric(Age)) 
```



```{r combine the datasets}
# Combine the the datasets by using the left_join function
data <- demo %>%
  full_join(CFIT, by = "ID") %>%
  full_join(ARFA, by = "ID") %>%
  full_join(eyes.closed, by = "ID")
  
```



### Index subjects by those that met the requirements

- No self-report injury
- No Epilepsy
- Removing missing data
- No one below 'IQ' of 70 (more than 2 SD below the mean)
- Extremely poor spellers (more than 3 SD below the mean)
- Issues with EEG recordings
- Did not survive EEG cleaning


```{r keep non rejects subjects}
# Set the working directory
setwd("~/MegaGrant")

# Load in the final sample data
Final_Sample_Size <- read_excel("Processed/Final.Sample.Size.xlsx")

# Index the data by the IDs in the final sample size
data2 <- data %>%
  filter(ID %in% Final_Sample_Size$ID)

# Change ID into a factor
data2$ID <- factor(data2$ID)

```


### Check for missing data

```{r check for missing data}
colSums(do.call(cbind,lapply(data2, function(x) is.na(x))))
```


### Descriptives

```{r some general descriptives}
# Use the describe function for these variables
Age_info <- round(data.frame(describe(data2$Age)),2)
CFIT_info <- round(data.frame(describe(data2$CFIT)),2)
Spell_info <- round(data.frame(describe(data2$SpellingError)),2)

# Rename the row
row.names(Age_info) <- "Age"
row.names(CFIT_info) <- "CFIT"
row.names(Spell_info) <- "SpellingEr"

# Binde them into one
descriptives_bound <- rbind(Age_info, CFIT_info) %>% rbind(Spell_info)

# Print out the table
descriptives_bound %>%
  kbl() %>% 
   kable_classic_2(full_width = F)


# Print out the frequency of groups
table_descr <- data.frame(cbind(table(data2$Sex, data2$Group)))

# Change the row names
row.names(table_descr) <- c("Female", "Male")

# Print the table
table_descr %>%
  kbl() %>% 
   kable_classic_2(full_width = F)
```

### Graph the distributions of numeric variables (before log)

```{r extract all continuous variables, out.width = "25%"}
# Create an index for numeric variables
num.indx <- which(lapply(data2, function(x) class(x)) %in% c("numeric", "integer"))

# Extract all index variables
num_df <- data2[,num.indx]

# Create a for loop to graph each of these variables
for(ii in 1:length(num_df)) {
  
  # Current variable
  var <- num_df[,ii]
  
  # Save the name of the variable
  name <- names(var)
  
  # Change it's name
  names(var) <- c("x")
  
  # Plot the histogram
  curplot <- var %>%
    ggplot(aes(x=x)) +
    geom_histogram(bins = 20,
                   color = "black",
                   fill = "white",
                   size = 1.5) +
    theme_minimal() +
    labs(title = paste(name,"Histogram"),
         x = NULL) +
    theme(plot.title = element_text(size = 30,
                                  hjust = 0.5),
        axis.title = element_text(size = 30, face = "bold"),
        axis.text = element_text(size = 30),
        plot.caption = element_text(size = 8,
                                    hjust = 0))
  
  # Plot it
  plot(curplot)
  
}
```

### Graph the distributions of numeric variables (after log)

```{r extract all continuous variables log, out.width = "25%"}
# Convert all 0's to .1 or smt
num_df <- tibble(data.frame(do.call(cbind,lapply(num_df, function(x) ifelse(x == 0, .1, x)))))


# transform all the variables to log
num_dflog <- log(num_df)

# Change the names
names(num_dflog) <- paste0(names(num_dflog),"_log")

# Create a for loop to graph each of these variables
for(ii in 1:length(num_dflog)) {
  
  # Current variable
  var <- num_dflog[,ii]
  
  # Save the name of the variable
  name <- names(var)
  
  # Change it's name
  names(var) <- c("x")
  
  # Plot the histogram
  curplot <- var %>%
    ggplot(aes(x=x)) +
    geom_histogram(bins = 20,
                   color = "black",
                   fill = "white",
                   size = 1.5) +
    theme_minimal() +
    labs(title = paste(name,"Histogram (Log)"),
         x = NULL) +
    theme(plot.title = element_text(size = 28,
                                  hjust = 0.5),
        axis.title = element_text(size = 28, face = "bold"),
        axis.text = element_text(size = 28),
        plot.caption = element_text(size = 8,
                                    hjust = 0))
  
  # Plot it
  plot(curplot)
  
}
```

### Multicollinearity? (Correlation Matrix)

```{r run a correlation matrix}
# Select the variables to correlate
cor_matrix <- select(num_dflog, Age_log, CFIT_log, SpellingError_log)

# Run the correlation matrix
data.frame(round(cor(cor_matrix),2)) %>%
  kbl() %>% 
   kable_classic_2(full_width = F)

```


### Converting data to Long format

Data in R and most statistical software must (or mostly) be in long format to be compatible with statistical functions. 

```{r creating the four variable datasets}
# Identify the variables they will all share
Front <- data2 %>% select(ID, Sex, Group)
covLog <- num_dflog %>% select(Age_log, CFIT_log, SpellingError_log)
delta <-  num_dflog %>% select(contains("delta"))
theta <- num_dflog %>% select(contains("theta"))
alpha <- num_dflog %>% select(contains("alpha"))
beta <- num_dflog %>% select(contains("beta"))

# Transform some variables into factors
Front$Sex <- factor(Front$Sex)
Front$Group <- factor(Front$Group)


# Merge these final datasets together
Delta <- cbind(Front, covLog,  delta)
Theta <- cbind(Front, covLog, theta)
Alpha <- cbind(Front, covLog, alpha)
Beta <- cbind(Front, covLog,  beta)
  
# Transform these data into long format
Delta.long <- Delta %>% 
  pivot_longer(cols = contains("delta"), names_to = "Topography", values_to = "Power") %>%
  mutate(Topography = unlist(lapply(strsplit(Topography, "_"), function(x) x[1])),
         Topography = factor(Topography))

Theta.long <- Theta %>%
  pivot_longer(cols = contains("theta"), names_to = "Topography", values_to = "Power") %>%
  mutate(Topography = unlist(lapply(strsplit(Topography, "_"), function(x) x[1])),
         Topography = factor(Topography))

Alpha.long <- Alpha %>%
  pivot_longer(cols = contains("alpha"), names_to = "Topography", values_to = "Power") %>%
  mutate(Topography = unlist(lapply(strsplit(Topography, "_"), function(x) x[1])),
         Topography = factor(Topography))

Beta.long <- Beta %>%
  pivot_longer(cols = contains("beta"), names_to = "Topography", values_to = "Power") %>%
  mutate(Topography = unlist(lapply(strsplit(Topography, "_"), function(x) x[1])),
         Topography = factor(Topography))
```



### Analysis Plan (Simple- no interactions)

Simplifying the analysis by investigating the relationship between spelling and power at 4 frequency bands (delta, theta, alpha, beta) while controlling for age, and topography. Since there are 4 outcome variables then we will be interpreting 4 models. 

We will be using dummy coding for right now, thus the intercepts of the model will represent the power of the frontal region for a female raised by her biological family. 

- Type I sums of squares were used for calculating the p values

```{r creating our first analyses}
# We will be running four regressions
contrasts(Alpha.long$Topography)
contrasts(Alpha.long$Sex)
contrasts(Alpha.long$Group)

# Run the first four models
model1a <- lm(Power ~ Age_log + Sex + CFIT_log + Group + Topography + SpellingError_log, data = Delta.long) 
model2a <- lm(Power ~  Age_log + Sex + CFIT_log + Group + Topography + SpellingError_log, data = Theta.long) 
model3a <- lm(Power ~  Age_log + Sex + CFIT_log + Group + Topography + SpellingError_log, data = Alpha.long) 
model4a <- lm(Power ~  Age_log + Sex + CFIT_log + Group + Topography + SpellingError_log, data = Beta.long) 
  
# Print out the model summaries
summary(model1a)
summary(model2a)
summary(model3a)
summary(model4a)
```
### Analysis Plan (2 way-interactions)

Is spelling error predictive of absolute power in some ages but not others? Thus, here we regress absolute power for four frequency bands (delta, theta, alpha, beta) on the same covariates as above with an age x spelling error interaction. 

- Type III sums of squares were used for calculating the p values
- Gives us more power when examining interactions
- Order of the covariates should not matter

```{r second models using a two way interaction}
# Run the first four models
model1b <- lm(Power ~ Sex + CFIT_log + Group + Topography + Age_log*SpellingError_log, data = Delta.long) 
model2b <- lm(Power ~ Sex + CFIT_log + Group + Topography + Age_log*SpellingError_log, data = Theta.long) 
model3b <- lm(Power ~ Sex + CFIT_log + Group + Topography + Age_log*SpellingError_log, data = Alpha.long) 
model4b <- lm(Power ~ Sex + CFIT_log + Group + Topography + Age_log*SpellingError_log, data = Beta.long) 
  

# run the model with type 3 SS
Anova(model1b, type = "III")
Anova(model2b, type = "III")
Anova(model3b, type = "III")
Anova(model4b, type = "III")
```


### Analysis Plan (3 way-interactions)

This is ambitious, what if the interaction of spelling error and age in predicting absolute power differed depending on the topographical location? Thus this overly ambitious three-way interaction was tested to see if that is the case. Since we have ~450 subjects, we may have enough power to test if this is true- not sure though. 

- Type III sums of squares were used for calculating the p values

```{r third model using a three way interaction}
# Run the first four models
model1c <- lm(Power ~ Sex + CFIT_log + Group + Topography*Age_log*SpellingError_log, data = Delta.long) 
model2c <- lm(Power ~ Sex + CFIT_log + Group + Topography*Age_log*SpellingError_log, data = Theta.long) 
model3c <- lm(Power ~ Sex + CFIT_log + Group + Topography*Age_log*SpellingError_log, data = Alpha.long) 
model4c <- lm(Power ~ Sex + CFIT_log + Group + Topography*Age_log*SpellingError_log, data = Beta.long) 
  

# run the model with type 3 SS
Anova(model1c, type = "III")
Anova(model2c, type = "III")
Anova(model3c, type = "III")
Anova(model4c, type = "III")
```

### Two-Way Interaction model seems to be the best and most informative.

We have decided to proceed the remaining of the analysis by focusing solely on the models that used a two-way interaction. 

### Visualizing the interactions

The graphs show the interaction for:

- Plot1: Delta (p = ns)
- Plot2: Theta (p = ns)
- Plot3: Alpha (p = ns)
- Plot4: Beta (p < .001)

```{r visualizing the interaction, out.width = "45%"}
# Plot the interaction effect of delta 
interact_plot(model = model1b, pred = SpellingError_log, modx = Age_log)
interact_plot(model = model2b, pred = SpellingError_log, modx = Age_log)
interact_plot(model = model3b, pred = SpellingError_log, modx = Age_log)
interact_plot(model = model4b, pred = SpellingError_log, modx = Age_log)

```

### Model Assumptions

```{r plot the residuals of the models, out.width="45%"}
plot(model1b, which = 1:2)
plot(model2b, which = 1:2)
plot(model3b, which = 1:2)
plot(model4b, which = 1:2)

```
